### 2017-01-22 homework/week-1/main.tex
Adding week 1 homework which closes #1 and closes #2

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath} 
\usepackage{amsthm}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\setlength\parindent{0pt}

\newtheorem*{definition}{Definition}
\newtheorem*{theorem}{Theorem}

\begin{document}

\section*{Construct a graph}
Construct a simple graph, \(G = (V, E)\), such that:

\begin{align}
|V| &\geq 8 \\
\chi(G) &> \chi^{*}(G)
\end{align}

Let \(\chi\) refer to the coloring generated by using the greedy algorithm with degree sequencing heuristic. Let \(\chi^{*}\) be the optimal coloring.

\subsection*{Example 1}
I constructed a simple graph, \(G = (V, E)\), such that \(\Delta(G) = 3\) and \(\delta(G) = 1\). I've let \(|V| = 8\) for this example.

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{graph-1.png}
\caption{Uncolored original graph \(G\)}
\end{figure}

By applying the greedy algorithm plus degree sequencing heuristic, we get the following coloring. The numbers indicate the ordering of vertices before applying the greedy algorithm. Any vertex of the same degree got assigned arbitrarily. This results in \(\chi(G) = 3\).

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{graph-2.png}
\caption{Coloring from applying the greedy plus degree sequencing heuristic}
\end{figure}

We can see that this is a bipartite graph and thus is \emph{2-colorable}. This means \(\chi^{*}(G) = 2\). Thus, this is an example graph that satisfies conditions \((1)\) and \((2)\). The optimal coloring is shown below.

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{graph-3.png}
\caption{Optimal coloring of graph \(G\)}
\end{figure}

\newpage

\section*{Upper Bounds}
Examine the upper bounds of the greedy heuristic and the greedy plus degree sequencing heuristic. \newline

To do so, let's create another example that satisfies conditions \((1)\) and \((2)\) from above. After doing some research into bipartite graphs, I learned that \emph{crown graphs} are excellent at showing how bad greedy heuristics can be.

\begin{definition}[Crown Graph]
A crown graph is a complete bipartite graph where the edges of a perfect matching have been removed.
\end{definition}

\subsection*{Example 2}

I constructed a simple crown graph, \(H = (V, E)\). I've let \(|V| = 10\) for this example.

\begin{figure}[H]
\centering
\includegraphics[scale=0.40]{graph-4.png}
\caption{Uncolored original graph \(H\)}
\end{figure}

We can see that \(\Delta(G) = 4\). We can also see \(d(v) = 4\) for all \(v \in V\). Thus, in both heuristics, the greedy algorithm would pick an order arbitrarily. We can show using this crown graph the worst-case scenario of these heuristics.

\begin{figure}[H]
\centering
\includegraphics[scale=0.40]{graph-5.png}
\caption{Worst-case coloring of \(H\) using either greedy heuristic}
\end{figure}

\newpage

In Figure 5, using either greedy heuristic with this ordering, we get \(\chi(H) = 5\). This gives us \(\frac{|V|}{2}\) colors. This is the worst-case for a crown graph, but this graph also demonstrates the upper bounds for these greedy algorithms. \newline

\subsection*{In General}

Let \(P = (V, E)\) be a simple, complete graph. \newline

We can see that \(\chi(H) = \Delta(H) + 1\). This is very easy to see in a \emph{complete} graph, where all vertices are connected to every other vertex. This means all vertices have degree \(|V| - 1\). Thus, every time we color a node, a new color is needed. And since we have \(\Delta(P) = |V| - 1\) and \(|V|\) vertices, we will need \(\Delta(P) + 1\) colors. This is proved in \emph{Brooks' Theorem}.

\begin{theorem}[Brooks' Theorem]
For any connected undirected graph \(G\) with maximum degree \(\Delta\), the chromatic number of \(G\) is at most \(\Delta\) unless \(G\) is a complete graph or an odd cycle, in which case the chromatic number is \(\Delta + 1\).
\end{theorem}

Overall, the greedy and the greedy plus degree sequencing heuristics can produce some very bad results. In graph \(H\), at the worst case, the greedy heuristic produces \(\chi(H) = 5\) when \(\chi^{*}(H) = 2\) as it is bipartite. This is shown below.

\begin{figure}[H]
\centering
\includegraphics[scale=0.40]{graph-6.png}
\caption{Optimal coloring of \(H\)}
\end{figure}

\end{document}

### 2017-01-24 homework/week-1/main.tex
Adding references and some changes to homework 1

@@
 \usepackage{amsthm}
-\usepackage[margin=1in]{geometry}
+\usepackage[left=1in,right=1in,top=0.9in,bottom=0.9in]{geometry}
 \usepackage{graphicx}
@@
 
+\section*{Heuristics}
+In this report, I will use two vertex coloring heuristics as defined in class:
+\begin{itemize}
+\item \textbf{Heuristic A}: Greedy algorithm
+\item \textbf{Heuristic B}: Greedy algorithm plus degree sequencing
+\end{itemize}
+
+Heuristic B is also known as the Welsh-Powell algorithm, which is defined in \cite{welsh}.
+
 \section*{Construct a graph}
@@
 
-Let \(\chi\) refer to the coloring generated by using the greedy algorithm with degree sequencing heuristic. Let \(\chi^{*}\) be the optimal coloring.
+Let \(\chi\) refer to the coloring generated by \emph{heuristic B}. Let \(\chi^{*}\) be the optimal coloring.
 
@@
 \centering
-\includegraphics[scale=0.5]{graph-1.png}
+\includegraphics[scale=0.6]{graph-1.png}
 \caption{Uncolored original graph \(G\)}
@@
 
-By applying the greedy algorithm plus degree sequencing heuristic, we get the following coloring.
-The numbers indicate the ordering of vertices before applying the greedy algorithm.
+By applying \emph{heuristic B}, we get the following coloring.
+The numbers indicate the ordering of vertices before applying the heuristic.
 Any vertex of the same degree got assigned arbitrarily.
@@
 \centering
-\includegraphics[scale=0.5]{graph-2.png}
-\caption{Coloring from applying the greedy plus degree sequencing heuristic}
+\includegraphics[scale=0.6]{graph-2.png}
+\caption{Coloring from applying heuristic B}
 \end{figure}
 
-We can see that this is a bipartite graph and thus is \emph{2-colorable}.
+\begin{definition}[Bipartite graph]
+A bipartite graph is one whose vertex set can be partitioned into two subsets X and Y such that each edge has one end in X and one end in Y
+\end{definition}
+
+We can see that this is a bipartite graph, defined above by \cite{bondymurty}.
+Thus, \(G\) is \emph{2-colorable}.
 This means \(\chi^{*}(G) = 2\).
-Thus, this is an example graph that satisfies conditions \((1)\) and \((2)\).
+\(G\) is an example graph that satisfies conditions \((1)\) and \((2)\).
 The optimal coloring is shown below.
@@
 \centering
-\includegraphics[scale=0.5]{graph-3.png}
+\includegraphics[scale=0.6]{graph-3.png}
 \caption{Optimal coloring of graph \(G\)}
 \end{figure}
-
-\newpage
 
@@
 To do so, let's create another example that satisfies conditions \((1)\) and \((2)\) from above.
-After doing some research into bipartite graphs, I learned that \emph{crown graphs} are excellent at showing how bad greedy heuristics can be.
+After doing some research into bipartite graphs, I learned that \emph{crown graphs} are excellent at showing how bad greedy heuristics can be, as shown and defined in \cite{kordecki}.
 
 \begin{definition}[Crown Graph]
-A crown graph is a complete bipartite graph where the edges of a perfect matching have been removed.
+A crown graph \(C_n = (V, E)\) is an undirected graph with two sets of vertices where \(V = V_1 \cup V_2\) with an edge from \(v_{1,i}\) to \(v_{2,j}\) whenever \(i \neq j\).
+A crown graph can also be viewed as a complete bipartite graph where the edges of a perfect matching have been removed.
 \end{definition}
@@
 \centering
-\includegraphics[scale=0.40]{graph-4.png}
+\includegraphics[scale=0.38]{graph-4.png}
 \caption{Uncolored original graph \(H\)}
@@
 \centering
-\includegraphics[scale=0.40]{graph-5.png}
+\includegraphics[scale=0.38]{graph-5.png}
 \caption{Worst-case coloring of \(H\) using either greedy heuristic}
@@
 
-\newpage
-
-In Figure 5, using either greedy heuristic with this ordering, we get \(\chi(H) = 5\).
+In Figure 5, using either heuristic with this ordering, we get \(\chi(H) = 5\).
 This gives us \(\frac{|V|}{2}\) colors.
-This is the worst-case for a crown graph, but this graph also demonstrates the upper bounds for these greedy algorithms.
+This is the worst-case for a crown graph as shown in \cite{johnson}, but this graph also demonstrates the upper bounds for these heuristics.
 \newline
@@
 And since we have \(\Delta(P) = |V| - 1\) and \(|V|\) vertices, we will need \(\Delta(P) + 1\) colors.
-This is proved in \emph{Brooks' Theorem}.
+This is proved in \emph{Brooks' Theorem}, defined below from \cite{bondymurty}.
 
@@
 
-Overall, the greedy and the greedy plus degree sequencing heuristics can produce some very bad results.
-In graph \(H\), at the worst case, the greedy heuristic produces \(\chi(H) = 5\) when \(\chi^{*}(H) = 2\) as it is bipartite.
+Overall, heuristic A and heuristic B can produce some very undesirable results.
+In graph \(H\), at the worst case, these heuristics produce \(\chi(H) = 5\) when \(\chi^{*}(H) = 2\) as it is bipartite.
 This is shown below.
@@
 \centering
-\includegraphics[scale=0.40]{graph-6.png}
+\includegraphics[scale=0.38]{graph-6.png}
 \caption{Optimal coloring of \(H\)}
@@
 
+\begin{thebibliography}{9}
+\bibitem{bondymurty}
+Bondy, J.A. and U.S.R. Murty [1976],
+\emph{Graph Theory with Applications},
+American Elsevier Publishing, New York, NY.
+
+\bibitem{johnson}
+Johnson, D.
+S.
+[1974], \emph{Worst-case behavior of graph coloring algorithms}, Proc.
+5th Southeastern Conf.
+on Combinatorics, Graph Theory, and Computing, Utilitas Mathematicae, Winnipeg, pp.
+513–527.
+
+\bibitem{kordecki}
+Kordecki, W. and A. Łyczkowska-Hanćkowiak [2016], \emph{Greedy online colouring with buffering}, arXiv preprint arXiv:1601.00252.
+
+\bibitem{welsh}
+Welsh, D. J. and Powell, M. B. [1967], \emph{An upper bound for the chromatic number of a graph and its application to timetabling problems}, The Computer Journal, 10(1), 85-86.
+
+\end{thebibliography}
+
 \end{document}

### 2017-01-25 homework/week-1/main.tex
Making edits to week 1 homework

@@
 \begin{definition}[Crown Graph]
-A crown graph \(C_n = (V, E)\) is an undirected graph with two sets of vertices where \(V = V_1 \cup V_2\) with an edge from \(v_{1,i}\) to \(v_{2,j}\) whenever \(i \neq j\).
+A crown graph \(CR_n = (V, E)\) is an undirected graph with two sets of vertices where \(V = V_1 \cup V_2\) with an edge from \(v_i \in V_1\) to \(v_{j} \in V_2\) whenever \(i \neq j\).
 A crown graph can also be viewed as a complete bipartite graph where the edges of a perfect matching have been removed.
@@
 And since we have \(\Delta(P) = |V| - 1\) and \(|V|\) vertices, we will need \(\Delta(P) + 1\) colors.
-This is proved in \emph{Brooks' Theorem}, defined below from \cite{bondymurty}.
+This is stated in \emph{Brooks' Theorem}.
 
@@
 
+The proof of \emph{Brook's Theorem} can be found in \cite{lovasz}.
 Overall, heuristic A and heuristic B can produce some very undesirable results.
@@
 
+\bibitem{lovasz}
+Lovász, L. [1975], \emph{Three short proofs in graph theory}, Journal of Combinatorial Theory, Series B, 19(3), 269-271.
+
 \bibitem{welsh}

### 2017-01-30 work/work.tex
Refactoring homework to report

--- 2017-01-25 homework/week-1/main.tex
@@
 \usepackage{float}
+\usepackage{titling}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+\setlength{\droptitle}{-1.5cm}
 \setlength\parindent{0pt}
 
+\makeatletter
+\def\BState{\State\hskip-\ALG@thistlm}
+\makeatother
+
+\title{Homework Report}
+\date{Last modified: \today}
+\author{Shawn Seymour\\ University of Minnesota Morris}
+
 \newtheorem*{definition}{Definition}
+\newtheorem{prop}{Proposition}
 \newtheorem*{theorem}{Theorem}
@@
 
+\maketitle
+
 \section*{Heuristics}
-In this report, I will use two vertex coloring heuristics as defined in class:
-\begin{itemize}
-\item \textbf{Heuristic A}: Greedy algorithm
-\item \textbf{Heuristic B}: Greedy algorithm plus degree sequencing
-\end{itemize}
+In this report, I will discuss multiple vertex coloring heuristics. I will define them as follows:
 
-Heuristic B is also known as the Welsh-Powell algorithm, which is defined in \cite{welsh}.
+\subsection*{Heuristic A}
+
+Heuristic A is the greedy algorithm. The input is a simple graph \(G = (V, E)\).
+
+\begin{algorithm}
+\caption{Greedy algorithm}
+\begin{algorithmic}[1]
+\State Label each vertex in $V$, i.e. $v_1, v_2, \dots, v_n$
+\For{each $v \in V$}
+\State Assign a color $p_i$ to $v_i$ using the smallest available $p_i$
+\EndFor
+\end{algorithmic}
+\end{algorithm}
+
+\subsection*{Heuristic B}
+
+Heuristic B is the greedy algorithm with degree sequencing.
+It orders the vertices according to the decreasing value of their degree.
+This is also known as the Welsh-Powell algorithm, which is defined in \cite{welsh}.
+
+\begin{algorithm}
+\caption{Welsh-Powell algorithm}
+\begin{algorithmic}[1]
+\State Label each vertex in $V$, i.e. $v_1, v_2, \dots, v_n$, such that $d_G(v_1) \geq d_G(v_2) \geq \dots \geq d_G(v_n)$
+\ForAll{$v \in V$}
+\State Assign a color $p_i$ to $v_i$ using the smallest available $p_i$
+\EndFor
+\end{algorithmic}
+\end{algorithm}
 
@@
 \centering
-\includegraphics[scale=0.6]{graph-1.png}
+\includegraphics[scale=0.6]{images/graph-1.png}
 \caption{Uncolored original graph \(G\)}
@@
 \centering
-\includegraphics[scale=0.6]{graph-2.png}
+\includegraphics[scale=0.6]{images/graph-2.png}
 \caption{Coloring from applying heuristic B}
@@
 \centering
-\includegraphics[scale=0.6]{graph-3.png}
+\includegraphics[scale=0.6]{images/graph-3.png}
 \caption{Optimal coloring of graph \(G\)}
@@
 \centering
-\includegraphics[scale=0.38]{graph-4.png}
+\includegraphics[scale=0.38]{images/graph-4.png}
 \caption{Uncolored original graph \(H\)}
@@
 \centering
-\includegraphics[scale=0.38]{graph-5.png}
+\includegraphics[scale=0.38]{images/graph-5.png}
 \caption{Worst-case coloring of \(H\) using either greedy heuristic}
@@
 \centering
-\includegraphics[scale=0.38]{graph-6.png}
+\includegraphics[scale=0.38]{images/graph-6.png}
 \caption{Optimal coloring of \(H\)}
 \end{figure}
+
+\newpage
 

### 2017-01-30 work/work.tex
Starting to move towards proposition structure

@@
 
-\section*{Construct a graph}
-Construct a simple graph, \(G = (V, E)\), such that:
+\section*{Findings}
+
+\begin{prop}
+Heuristic A and Heuristic B do not always produce an optimal solution, i.e. they do not always produce a minimum coloring of a graph.
+\end{prop}
+
+I will show this for \emph{heuristic B}.
+It is trivial to show the same for \emph{heuristic A} as \emph{heuristic B} includes the same steps as \emph{heuristic A}.
+I will construct a simple graph, \(G = (V, E)\), such that:
 
@@
 
-Let \(\chi\) refer to the coloring generated by \emph{heuristic B}. Let \(\chi^{*}\) be the optimal coloring.
+Let \(\chi\) refer to the coloring of \(G\) generated by \emph{heuristic B}. Let \(\chi^{*}\) be the optimal coloring of \(G\).
 
@@
 
-\section*{Upper Bounds}
-Examine the upper bounds of the greedy heuristic and the greedy plus degree sequencing heuristic. \newline
+\begin{prop}
+Heuristic A and heuristic B have an upper bound of \(\Delta + 1\).
+\end{prop}
 
-To do so, let's create another example that satisfies conditions \((1)\) and \((2)\) from above.
+To show this, I will create another example that satisfies conditions \((1)\) and \((2)\) from above.
 After doing some research into bipartite graphs, I learned that \emph{crown graphs} are excellent at showing how bad greedy heuristics can be, as shown and defined in \cite{kordecki}.
@@
 We can see that \(\Delta(G) = 4\).
-We can also see \(d(v) = 4\) for all \(v \in V\).
-Thus, in both heuristics, the greedy algorithm would pick an order arbitrarily.
+We can also see \(d_G(v) = 4\) for all \(v \in V\).
+Thus, in both heuristics A and B, the greedy algorithm would pick an order arbitrarily.
 We can show using this crown graph the worst-case scenario of these heuristics.
@@
 \includegraphics[scale=0.38]{images/graph-5.png}
-\caption{Worst-case coloring of \(H\) using either greedy heuristic}
+\caption{Worst-case coloring of \(H\) using either heuristic A or B}
 \end{figure}
@@
 
-\subsection*{In General}
-
-Let \(P = (V, E)\) be a simple, complete graph. \newline
+\textbf{In General}: Let's take a look at how this looks in general. Let \(P = (V, E)\) be a simple, complete graph. \newline
 
@@
 This is stated in \emph{Brooks' Theorem}.
+
+\newpage
 
@@
 \end{figure}
-
-\newpage
 

### 2017-01-30 work/work.tex
Adding heuristic C and example on how it is not optimal

@@
 \section*{Heuristics}
-In this report, I will discuss multiple vertex coloring heuristics. I will define them as follows:
+In this report, I will discuss multiple vertex coloring heuristics. The input for all heuristics defined here is a simple graph \(G = (V, E)\). I will define them as follows:
 
@@
 
-Heuristic A is the greedy algorithm. The input is a simple graph \(G = (V, E)\).
+Heuristic A is the greedy algorithm.
 
@@
 \end{algorithm}
+
+\subsection*{Heuristic C}
+
+Heuristic C colors a graph by finding maximal independent sets of \(G\).
+
+\begin{algorithm}
+\caption{Coloring via maximal independent set algorithm}
+\begin{algorithmic}[1]
+\While{$G$ is non-empty}
+\State Find a maximal independent set of $G$, i.e. $S_i$
+\State Color all vertices in $S_i$ with color $p$, where $p$ is the smallest color available
+\State Let $G \leftarrow G[V \setminus S_i ] $
+\EndWhile
+\end{algorithmic}
+\end{algorithm}
+
+\newpage
 
@@
 \begin{prop}
+Heuristic C also does not always produce an optimal solution, i.e. it does not always produce a minimum coloring of a graph.
+\end{prop}
+
+\subsection*{Example 2}
+To show this, I will create another example that satisfies conditions \((1)\) and \((2)\) from above. I've constructed a graph \(K = (V,E)\) for this example.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.38]{images/mis-1.png}
+\caption{Uncolored original graph \(K\)}
+\end{figure}
+
+By applying heuristic C and finding maximal independent sets, we see that \(K\) is \emph{5-colorable}.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.38]{images/mis-2.png}
+\caption{Coloring of \(K\) after applying heuristic C}
+\end{figure}
+
+This is a non-optimal solution.
+We can see that the subgraph created by vertices \(\{5, 6, 7, 8\}\) is complete and thus we need at least 4 colors.
+This graph is \emph{4-colorable} however, meaning \(\chi^{*}(K) = 4\).
+Thus, we can see that heuristic C does not always give us an optimal solution.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.38]{images/mis-3.png}
+\caption{Optimal coloring of \(K\)}
+\end{figure}
+
+\begin{prop}
 Heuristic A and heuristic B have an upper bound of \(\Delta + 1\).
@@
 
-\subsection*{Example 2}
+\subsection*{Example 3}
 
@@
 This is stated in \emph{Brooks' Theorem}.
-
-\newpage
 

### 2017-01-31 code/src/com/devshawn/coloring/ColoringException.java
Initial design of vertex coloring implementation

@@ -1,0 +1,12 @@
+package com.devshawn.coloring;
+
+public class ColoringException extends RuntimeException {
+
+    public static String INVALID_MATRIX = "Your adjacency matrix is invalid.";
+    public static String SELF_LOOPS = "Your graph is not a simple graph. It contains self-loops.";
+
+    public ColoringException (String message) {
+        super(message);
+    }
+
+}

### 2017-01-31 code/src/com/devshawn/coloring/ColoringHeuristic.java

@@ -1,0 +1,19 @@
+package com.devshawn.coloring;
+
+import java.awt.*;
+
+public enum ColoringHeuristic {
+    GREEDY ("greedy"),
+    WELSH_POWELL ("welsh-powell"),
+    MAXIMAL_INDEPENDENT_SET ("maximal independent set");
+
+    private String name;
+
+    ColoringHeuristic(String name) {
+        this.name = name;
+    }
+
+    public String getName() {
+        return this.name;
+    }
+}

### 2017-01-31 code/src/com/devshawn/coloring/ColoringResult.java

@@ -1,0 +1,65 @@
+package com.devshawn.coloring;
+
+public class ColoringResult {
+    private int[] result;
+    private int highestColor;
+    private ColoringHeuristic heuristicUsed;
+    private int[] colors;
+
+    public ColoringResult(int[] result, ColoringHeuristic heuristicUsed) {
+
+        // Find how many colors we used
+        for(int i = 0; i < result.length; i++) {
+            int highestColor = -1;
+
+            if(result[i] > highestColor) {
+                highestColor = result[i];
+            }
+
+            // Add 1 as colors start with 0
+            this.highestColor = highestColor + 1;
+        }
+
+        // Count our color usage
+        this.colors = new int[highestColor];
+        for(int i = 0; i < result.length; i++) {
+            colors[result[i]]++;
+        }
+
+        this.result = result;
+        this.heuristicUsed = heuristicUsed;
+    }
+
+    public int getColoringNumber() {
+        return highestColor;
+    }
+
+    public int[] getColoring() {
+        return this.result;
+    }
+
+    public String getSummary() {
+        String string = "The graph was colored with the " + heuristicUsed.getName() + " heuristic.";
+        string += "\r\n\tColors used: " + highestColor;
+
+        string += "\r\n\t-----------------------------------";
+
+        string += "\r\n\tColor usage:";
+        for(int i = 0; i < colors.length; i++) {
+            string += "\r\n\tColor " + i + " => " + colors[i];
+        }
+
+        string += "\r\n\t-----------------------------------";
+
+        string += "\r\n\tVertex to color mapping:";
+        for(int i = 0; i < result.length; i++) {
+            string += "\r\n\tVertex " + i + " => color " + result[i];
+        }
+
+        return string;
+    }
+
+    public void printSummary() {
+        System.out.println(getSummary());
+    }
+}

### 2017-01-31 src/main/java/com/devshawn/coloring/Coloring.java
Setting up gradle for building and testing

@@ -1,0 +1,114 @@
+package com.devshawn.coloring;
+
+import java.util.Arrays;
+import java.util.HashMap;
+import java.util.Map;
+
+public class Coloring {
+
+    private int[][] graph;
+    private int vertexCount;
+    private int edgeCount;
+
+    // Apply a given heuristic to the graph
+    public ColoringResult applyHeuristic(ColoringHeuristic heuristic) {
+        switch(heuristic) {
+            case GREEDY:
+                return applyGreedyHeuristic();
+            case MAXIMAL_INDEPENDENT_SET:
+                break;
+            case WELSH_POWELL:
+                break;
+        }
+
+        return null;
+    }
+
+    // Implemented heuristics
+    private ColoringResult applyGreedyHeuristic() {
+
+        // Make result array of vertices, assign no color to them
+        int[] result = new int[graph.length];
+        for(int i = 0; i < graph.length; i++) result[i] = -1;
+
+        // Make temporary available array to see what colors are available
+        boolean available[] = new boolean[graph.length];
+        for(int i = 0; i < graph.length; i++) available[i] = true;
+
+        // Color the graph!
+        for(int i = 0; i < graph.length; i++) {
+
+            // Check neighbors for colors
+            for(int j = 0; j < graph[i].length; j++) {
+                if(graph[i][j] == 1 && result[j] != -1) {
+                    available[result[j]] = false;
+                }
+            }
+
+            // Find smallest available color
+            int color = -1;
+            for(int j = 0; j < available.length; j++) {
+                if(available[j]) {
+                    color = j;
+                    break;
+                }
+            }
+
+            // Set this vertex to that color
+            result[i] = color;
+
+            // Reset our available colors
+            for(int j = 0; j < graph.length; j++) available[j] = true;
+        }
+
+        // Generate a ColorResult
+        return new ColoringResult(result, ColoringHeuristic.GREEDY);
+    }
+
+    // Simple functions to get information about the graph
+
+    public int[][] getGraph() {
+        return graph;
+    }
+
+    public void setGraph(int[][] graph) {
+        if(graph.length != 0 && graph.length != graph[0].length) {
+            throw new ColoringException(ColoringException.INVALID_MATRIX);
+        }
+
+        for(int i = 0; i < graph.length; i++) {
+            if(graph[i][i] != 0) {
[…]

### 2017-01-31 src/main/java/com/devshawn/coloring/Main.java

@@ -1,0 +1,19 @@
+package com.devshawn.coloring;
+
+public class Main {
+
+    // For testing without writing JUnit tests
+
+    public static void main(String[] args) {
+        int[][] graph = {
+                {0, 1, 1},
+                {1, 0, 1},
+                {1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
+        result.printSummary();
+    }
+}

### 2017-01-31 src/test/java/com/devshawn/coloring/VertexColoringSpec.java

@@ -1,0 +1,114 @@
+package com.devshawn.coloring;
+
+import org.junit.Rule;
+import org.junit.Test;
+import org.junit.rules.ExpectedException;
+
+import static org.junit.Assert.assertArrayEquals;
+import static org.junit.Assert.assertEquals;
+import static org.junit.Assert.assertTrue;
+
+public class VertexColoringSpec {
+
+    @Rule
+    public ExpectedException thrown = ExpectedException.none();
+
+    // Tests for setting up
+
+    @Test
+    public void validGraphSetup() {
+        int[][] graph = {
+                {0, 1, 1},
+                {1, 0, 1},
+                {1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+        assertTrue(graph == coloring.getGraph());
+        assertEquals(3, coloring.getVertexCount());
+        assertEquals(3, coloring.getEdgeCount());
+
+    }
+
+    @Test
+    public void invalidGraphSet() throws ColoringException {
+        thrown.expect(ColoringException.class);
+        thrown.expectMessage(ColoringException.INVALID_MATRIX);
+
+        int[][] graph = {
+                {1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+    }
+
+    @Test
+    public void nonSimpleGraph() throws ColoringException {
+        thrown.expect(ColoringException.class);
+        thrown.expectMessage(ColoringException.SELF_LOOPS);
+
+        int[][] graph = {
+                {0, 1, 1},
+                {1, 1, 1},
+                {1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+    }
+
+    // Test greedy heuristic with some known greedy colorings
+
+    @Test
+    public void smallGreedyCompleteGraph() {
+
+        // a complete 3-vertex graph
+        int[][] graph = {
+                {0, 1, 1},
+                {1, 0, 1},
+                {1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
+
+        assertEquals(3, result.getColoringNumber());
+        int[] expectedColoring = {0, 1, 2};
[…]

### 2017-01-31 code/src/com/devshawn/coloring/Coloring.java
Removing old code directory

@@ -1,114 +1,0 @@
-package com.devshawn.coloring;
-
-import java.util.Arrays;
-import java.util.HashMap;
-import java.util.Map;
-
-public class Coloring {
-
-    private int[][] graph;
-    private int vertexCount;
-    private int edgeCount;
-
-    // Apply a given heuristic to the graph
-    public ColoringResult applyHeuristic(ColoringHeuristic heuristic) {
-        switch(heuristic) {
-            case GREEDY:
-                return applyGreedyHeuristic();
-            case MAXIMAL_INDEPENDENT_SET:
-                break;
-            case WELSH_POWELL:
-                break;
-        }
-
-        return null;
-    }
-
-    // Implemented heuristics
-    private ColoringResult applyGreedyHeuristic() {
-
-        // Make result array of vertices, assign no color to them
-        int[] result = new int[graph.length];
-        for(int i = 0; i < graph.length; i++) result[i] = -1;
-
-        // Make temporary available array to see what colors are available
-        boolean available[] = new boolean[graph.length];
-        for(int i = 0; i < graph.length; i++) available[i] = true;
-
-        // Color the graph!
-        for(int i = 0; i < graph.length; i++) {
-
-            // Check neighbors for colors
-            for(int j = 0; j < graph[i].length; j++) {
-                if(graph[i][j] == 1 && result[j] != -1) {
-                    available[result[j]] = false;
-                }
-            }
-
-            // Find smallest available color
-            int color = -1;
-            for(int j = 0; j < available.length; j++) {
-                if(available[j]) {
-                    color = j;
-                    break;
-                }
-            }
-
-            // Set this vertex to that color
-            result[i] = color;
-
-            // Reset our available colors
-            for(int j = 0; j < graph.length; j++) available[j] = true;
-        }
-
-        // Generate a ColorResult
-        return new ColoringResult(result, ColoringHeuristic.GREEDY);
-    }
-
-    // Simple functions to get information about the graph
-
-    public int[][] getGraph() {
-        return graph;
-    }
-
-    public void setGraph(int[][] graph) {
-        if(graph.length != 0 && graph.length != graph[0].length) {
-            throw new ColoringException(ColoringException.INVALID_MATRIX);
-        }
-
-        for(int i = 0; i < graph.length; i++) {
-            if(graph[i][i] != 0) {
[…]

### 2017-01-31 code/src/com/devshawn/coloring/ColoringException.java

@@ -1,12 +1,0 @@
-package com.devshawn.coloring;
-
-public class ColoringException extends RuntimeException {
-
-    public static String INVALID_MATRIX = "Your adjacency matrix is invalid.";
-    public static String SELF_LOOPS = "Your graph is not a simple graph. It contains self-loops.";
-
-    public ColoringException (String message) {
-        super(message);
-    }
-
-}

### 2017-01-31 code/src/com/devshawn/coloring/ColoringHeuristic.java

@@ -1,19 +1,0 @@
-package com.devshawn.coloring;
-
-import java.awt.*;
-
-public enum ColoringHeuristic {
-    GREEDY ("greedy"),
-    WELSH_POWELL ("welsh-powell"),
-    MAXIMAL_INDEPENDENT_SET ("maximal independent set");
-
-    private String name;
-
-    ColoringHeuristic(String name) {
-        this.name = name;
-    }
-
-    public String getName() {
-        return this.name;
-    }
-}

### 2017-01-31 code/src/com/devshawn/coloring/ColoringResult.java

@@ -1,65 +1,0 @@
-package com.devshawn.coloring;
-
-public class ColoringResult {
-    private int[] result;
-    private int highestColor;
-    private ColoringHeuristic heuristicUsed;
-    private int[] colors;
-
-    public ColoringResult(int[] result, ColoringHeuristic heuristicUsed) {
-
-        // Find how many colors we used
-        for(int i = 0; i < result.length; i++) {
-            int highestColor = -1;
-
-            if(result[i] > highestColor) {
-                highestColor = result[i];
-            }
-
-            // Add 1 as colors start with 0
-            this.highestColor = highestColor + 1;
-        }
-
-        // Count our color usage
-        this.colors = new int[highestColor];
-        for(int i = 0; i < result.length; i++) {
-            colors[result[i]]++;
-        }
-
-        this.result = result;
-        this.heuristicUsed = heuristicUsed;
-    }
-
-    public int getColoringNumber() {
-        return highestColor;
-    }
-
-    public int[] getColoring() {
-        return this.result;
-    }
-
-    public String getSummary() {
-        String string = "The graph was colored with the " + heuristicUsed.getName() + " heuristic.";
-        string += "\r\n\tColors used: " + highestColor;
-
-        string += "\r\n\t-----------------------------------";
-
-        string += "\r\n\tColor usage:";
-        for(int i = 0; i < colors.length; i++) {
-            string += "\r\n\tColor " + i + " => " + colors[i];
-        }
-
-        string += "\r\n\t-----------------------------------";
-
-        string += "\r\n\tVertex to color mapping:";
-        for(int i = 0; i < result.length; i++) {
-            string += "\r\n\tVertex " + i + " => color " + result[i];
-        }
-
-        return string;
-    }
-
-    public void printSummary() {
-        System.out.println(getSummary());
-    }
-}

### 2017-01-31 code/src/com/devshawn/coloring/Main.java

@@ -1,19 +1,0 @@
-package com.devshawn.coloring;
-
-public class Main {
-
-    // For testing without writing JUnit tests
-
-    public static void main(String[] args) {
-        int[][] graph = {
-                {0, 1, 1},
-                {1, 0, 1},
-                {1, 1, 0}
-        };
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
-        result.printSummary();
-    }
-}

### 2017-01-31 code/test/com/devshawn/coloring/VertexColoringSpec.java

@@ -1,114 +1,0 @@
-package com.devshawn.coloring;
-
-import org.junit.Rule;
-import org.junit.Test;
-import org.junit.rules.ExpectedException;
-
-import static org.junit.Assert.assertArrayEquals;
-import static org.junit.Assert.assertEquals;
-import static org.junit.Assert.assertTrue;
-
-public class VertexColoringSpec {
-
-    @Rule
-    public ExpectedException thrown = ExpectedException.none();
-
-    // Tests for setting up
-
-    @Test
-    public void validGraphSetup() {
-        int[][] graph = {
-                {0, 1, 1},
-                {1, 0, 1},
-                {1, 1, 0}
-        };
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        assertTrue(graph == coloring.getGraph());
-        assertEquals(3, coloring.getVertexCount());
-        assertEquals(3, coloring.getEdgeCount());
-
-    }
-
-    @Test
-    public void invalidGraphSet() throws ColoringException {
-        thrown.expect(ColoringException.class);
-        thrown.expectMessage(ColoringException.INVALID_MATRIX);
-
-        int[][] graph = {
-                {1, 1, 0}
-        };
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-    }
-
-    @Test
-    public void nonSimpleGraph() throws ColoringException {
-        thrown.expect(ColoringException.class);
-        thrown.expectMessage(ColoringException.SELF_LOOPS);
-
-        int[][] graph = {
-                {0, 1, 1},
-                {1, 1, 1},
-                {1, 1, 0}
-        };
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-    }
-
-    // Test greedy heuristic with some known greedy colorings
-
-    @Test
-    public void smallGreedyCompleteGraph() {
-
-        // a complete 3-vertex graph
-        int[][] graph = {
-                {0, 1, 1},
-                {1, 0, 1},
-                {1, 1, 0}
-        };
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
-
-        assertEquals(3, result.getColoringNumber());
-        int[] expectedColoring = {0, 1, 2};
[…]

### 2017-01-31 src/main/java/com/devshawn/coloring/Coloring.java
Adding more testing to bring up coverage

@@ -12,16 +12,19 @@ public class Coloring {
 
     // Apply a given heuristic to the graph
     public ColoringResult applyHeuristic(ColoringHeuristic heuristic) {
+        ColoringResult result = new ColoringResult();
+
         switch(heuristic) {
             case GREEDY:
-                return applyGreedyHeuristic();
-            case MAXIMAL_INDEPENDENT_SET:
+                result = applyGreedyHeuristic();
                 break;
             case WELSH_POWELL:
                 break;
+            case MAXIMAL_INDEPENDENT_SET:
+                break;
         }
 
-        return null;
+        return result;
     }
 
     // Implemented heuristics

### 2017-01-31 src/main/java/com/devshawn/coloring/ColoringResult.java

@@ -6,6 +6,10 @@ public class ColoringResult {
     private ColoringHeuristic heuristicUsed;
     private int[] colors;
 
+    public ColoringResult() {
+
+    }
+
     public ColoringResult(int[] result, ColoringHeuristic heuristicUsed) {
 
         // Find how many colors we used

### 2017-01-31 src/test/java/com/devshawn/coloring/ColoringHeuristicSpec.java

@@ -1,0 +1,14 @@
+package com.devshawn.coloring;
+
+import org.junit.Test;
+
+import static org.junit.Assert.assertEquals;
+
+public class ColoringHeuristicSpec {
+
+    @Test
+    public void getName() {
+        ColoringHeuristic heuristic = ColoringHeuristic.GREEDY;
+        assertEquals("greedy", heuristic.getName());
+    }
+}

### 2017-01-31 src/test/java/com/devshawn/coloring/ColoringResultSpec.java

@@ -1,0 +1,39 @@
+package com.devshawn.coloring;
+
+import org.junit.Test;
+
+import static org.junit.Assert.assertEquals;
+import static org.junit.Assert.assertTrue;
+
+public class ColoringResultSpec {
+
+    @Test
+    public void getSummary() {
+        int[] results = {0};
+        ColoringResult result = new ColoringResult(results, ColoringHeuristic.GREEDY);
+
+        String expectedResult = "The graph was colored with the greedy heuristic.";
+        expectedResult += "\r\n\tColors used: 1";
+        expectedResult += "\r\n\t-----------------------------------";
+        expectedResult += "\r\n\tColor usage:";
+        expectedResult += "\r\n\tColor 0 => 1";
+        expectedResult += "\r\n\t-----------------------------------";
+        expectedResult += "\r\n\tVertex to color mapping:";
+        expectedResult += "\r\n\tVertex 0 => color 0";
+
+        assertEquals(expectedResult, result.getSummary());
+    }
+
+    @Test
+    public void printSummary() {
+        // test for coverage on this function
+        int[] results = {0};
+        ColoringResult result = new ColoringResult(results, ColoringHeuristic.GREEDY);
+
+        result.printSummary();
+        assertTrue(true);
+    }
+
+
+}
+

### 2017-01-31 src/test/java/com/devshawn/coloring/ColoringSpec.java

@@ -8,7 +8,7 @@ import static org.junit.Assert.assertArrayEquals;
 import static org.junit.Assert.assertEquals;
 import static org.junit.Assert.assertTrue;
 
-public class VertexColoringSpec {
+public class ColoringSpec {
 
     @Rule
     public ExpectedException thrown = ExpectedException.none();

### 2017-01-31 work/work.tex
Adding a couple more propositions

@@
 
+\newpage
+
+\begin{prop}
+Heuristic A and Heuristic B have a running time of $O(|V| + |E|)$.
+\end{prop}
+
+It is possible to implement heuristic A and heuristic B efficiently and achieve a linear running time \cite{kubale}.
+By using an adjacency list representation of a graph \(G = (V, E)\), the greedy algorithm will look once at each vertex and each edge twice.
+This is because for each vertex, all of its neighbors are checked.
+This leads to the algorithm crossing each edge twuce.
+As this is still linear, we know \(O(|V| + 2|E|) = O(|V| + |E|)\).
+
+\begin{prop}
+Heuristic A and Heuristic B produce a feasible, proper coloring of input graph \(G = (V, E)\).
+\end{prop}
+
+Let \(G = (V, E)\) be a simple graph.
+Assume that two vertices, say \(v_1, v_2 \in V\), are connected by an edge \(e_1 \in E\).
+Assume that \(v_1, v_2\) are colored the same color, say \(p_1\).
+We know that in both heuristics, we assign the smallest color \(p_i\) that is not being used by any of the neighboring vertices connected by an edge \(e_i\).
+WLOG, the heuristic would color \(v_1\) color \(p_1\).
+The next iteration, while vertex \(v_2\) is selected, the heuristic would see color \(p_1\) is assigned to a neighboring vertex \((v_1)\), and assign the next available color \(p_2\).
+Thus, we've reached a contradiction.
+
 \begin{thebibliography}{9}
@@
 Bondy, J.A. and U.S.R. Murty [1976],
-\emph{Graph Theory with Applications},
+\emph{Graph theory with applications},
 American Elsevier Publishing, New York, NY.
@@
 
+\bibitem{kubale}
+Kubale, Marek [2004], \emph{Graph colorings}, Vol. 352, American Mathematical Soc.
+
 \bibitem{lovasz}

### 2017-01-31 work/update.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+latexmk -pdf work.tex

### 2017-01-31 work/work.tex
Adding propositions to homework report

@@
 
+\subsection*{Complexity of Vertex Coloring}
+\begin{prop}
+Vertex coloring is an NP-hard problem.
+\end{prop}
+
+Coloring a graph is computationally complex.
+It is NP-hard to compute the chromatic number of a given graph \cite{garey}.
+It is NP-complete to decide if a given graph has a \emph{k-coloring} for a given \(k\).
+This is because it is easy to verify if a coloring of the given graph is a \emph{k-coloring} in polynomial time.
+This can be done by iterating through the list of edges and checking that each attached vertex has a different color.\newline
+
 \begin{thebibliography}{9}
@@
 513–527.
+
+\bibitem{garey}
+Garey, M. R., Johnson, D. S., \& Stockmeyer, L. [1976], \emph{Some simplified NP-complete graph problems}, Theoretical computer science, 1(3), 237-267.
 

### 2017-01-31 src/main/java/com/devshawn/coloring/Coloring.java
Working on graph generator

@@ -1,9 +1,5 @@
 package com.devshawn.coloring;
 
-import java.util.Arrays;
-import java.util.HashMap;
-import java.util.Map;
-
 public class Coloring {
 
     private int[][] graph;

### 2017-01-31 src/main/java/com/devshawn/coloring/ColoringResult.java

@@ -1,5 +1,7 @@
 package com.devshawn.coloring;
 
+import java.util.Arrays;
+
 public class ColoringResult {
     private int[] result;
     private int highestColor;
@@ -13,19 +15,19 @@ public class ColoringResult {
     public ColoringResult(int[] result, ColoringHeuristic heuristicUsed) {
 
         // Find how many colors we used
-        for(int i = 0; i < result.length; i++) {
-            int highestColor = -1;
+        int highestColor = -1;
 
+        for(int i = 0; i < result.length; i++) {
             if(result[i] > highestColor) {
                 highestColor = result[i];
             }
-
-            // Add 1 as colors start with 0
-            this.highestColor = highestColor + 1;
         }
 
+        // Add 1 as colors start with 0
+        this.highestColor = highestColor + 1;
+
         // Count our color usage
-        this.colors = new int[highestColor];
+        this.colors = new int[this.highestColor];
         for(int i = 0; i < result.length; i++) {
             colors[result[i]]++;
         }

### 2017-01-31 src/main/java/com/devshawn/coloring/GraphGenerator.java

@@ -1,0 +1,37 @@
+package com.devshawn.coloring;
+
+public class GraphGenerator {
+
+    // simple graph with a given number of vertices
+    // an edge between any two vertices with given probability
+    // this is referred to as the Erdos-Renyi model
+    public static int[][] simple(int vertices, double probability) {
+        if(probability < 0 || probability > 1) {
+            throw new ColoringException("Probability for generating a graph must be between 0 and 1.");
+        }
+
+        int[][] graph = new int[vertices][vertices];
+
+        for(int i = 0; i < vertices; i++) {
+            for(int j = 0; j < i; j++) {
+                int number = (Math.random() < probability) ? 1 : 0;
+                graph[i][j] = number;
+                graph[j][i] = number;
+            }
+        }
+
+        return graph;
+    }
+
+    public static boolean isSymmetric(int[][] graph) {
+        for(int i = 0; i < graph.length; i++) {
+            for(int j = 0; j < i; j++) {
+                if(graph[i][j] != graph[j][i]) {
+                    return false;
+                }
+            }
+        }
+
+        return true;
+    }
+}

### 2017-01-31 src/main/java/com/devshawn/coloring/Main.java

@@ -1,15 +1,93 @@
 package com.devshawn.coloring;
 
+import java.util.Arrays;
+import java.util.Scanner;
+
 public class Main {
+
+    private static Scanner sc = new Scanner(System.in);
 
     // For testing without writing JUnit tests
 
     public static void main(String[] args) {
-        int[][] graph = {
-                {0, 1, 1},
-                {1, 0, 1},
-                {1, 1, 0}
-        };
+
+        while(true) {
+            printOptions();
+            pickOption();
+        }
+
+    }
+
+    private static void printOptions() {
+        System.out.println("Welcome to the vertex coloring library! Please pick an option below by typing its number:");
+        System.out.println("\r\n\t1. Enter your own graph");
+        System.out.println("\r\n\t2. Generate a graph with a specified number of vertices and number of edges");
+        System.out.println("\r\n\t3. Generate a graph with a specified number of vertices and percentage of edges\r\n");
+    }
+
+    private static void pickOption() {
+        int option = sc.nextInt();
+        sc.nextLine();
+
+
+        switch(option) {
+            case 1:
+                enterGraph();
+                break;
+            case 2:
+                System.out.println("Oops! Not implemented yet!");
+                break;
+            case 3:
+                generatePercentageGraph();
+                break;
+            default:
+                System.out.println("Please choose an option.");
+                break;
+        }
+    }
+
+    private static void generatePercentageGraph() {
+        System.out.println("Please enter the number of vertices you want:");
+
+        int vertices = sc.nextInt();
+        sc.nextLine();
+
+        System.out.println("Please enter the percentage at which an edge should be placed (i.e. 30):");
+        double percentage = sc.nextDouble();
+        percentage = percentage / (double) 100;
+        sc.nextLine();
+
+        int[][] graph = GraphGenerator.simple(vertices, percentage);
+
+        System.out.println(GraphGenerator.isSymmetric(graph));
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
+        result.printSummary();
+    }
+
+    private static void enterGraph() {
+        System.out.println("Please enter the number of vertices your graph has: ");
+
+        int vertices = sc.nextInt();
+        sc.nextLine();
+
+        System.out.println("Please enter the adjacency matrix of your graph. Example for |V| = 3: ");
[…]

### 2017-01-31 src/test/java/com/devshawn/coloring/GraphGeneratorSpec.java

@@ -1,0 +1,16 @@
+package com.devshawn.coloring;
+
+import org.junit.Test;
+
+import java.util.Arrays;
+
+import static org.junit.Assert.assertEquals;
+
+public class GraphGeneratorSpec {
+
+    @Test
+    public void simple() {
+        int[][] graph = GraphGenerator.simple(1000, 0.3);
+        assertEquals(true, GraphGenerator.isSymmetric(graph));
+    }
+}

### 2017-02-01 src/main/java/com/devshawn/coloring/ColoringException.java
Adding more testing for current main function

@@ -4,6 +4,7 @@ public class ColoringException extends RuntimeException {
 
     public static String INVALID_MATRIX = "Your adjacency matrix is invalid.";
     public static String SELF_LOOPS = "Your graph is not a simple graph. It contains self-loops.";
+    public static String INVALID_PROBABILITY = "Probability for generating a graph must be between 0 and 1.";
 
     public ColoringException (String message) {
         super(message);

### 2017-02-01 src/main/java/com/devshawn/coloring/ColoringResult.java

@@ -1,5 +1,6 @@
 package com.devshawn.coloring;
 
+import java.io.PrintWriter;
 import java.util.Arrays;
 
 public class ColoringResult {
@@ -65,7 +66,8 @@ public class ColoringResult {
         return string;
     }
 
-    public void printSummary() {
-        System.out.println(getSummary());
+    public void printSummary(PrintWriter output) {
+        output.println(getSummary());
+        output.flush();
     }
 }

### 2017-02-01 src/main/java/com/devshawn/coloring/GraphGenerator.java

@@ -7,7 +7,7 @@ public class GraphGenerator {
     // this is referred to as the Erdos-Renyi model
     public static int[][] simple(int vertices, double probability) {
         if(probability < 0 || probability > 1) {
-            throw new ColoringException("Probability for generating a graph must be between 0 and 1.");
+            throw new ColoringException(ColoringException.INVALID_PROBABILITY);
         }
 
         int[][] graph = new int[vertices][vertices];

### 2017-02-01 src/main/java/com/devshawn/coloring/Main.java

@@ -1,15 +1,23 @@
 package com.devshawn.coloring;
 
+import java.io.PrintStream;
+import java.io.PrintWriter;
+import java.io.StringWriter;
 import java.util.Arrays;
 import java.util.Scanner;
 
 public class Main {
 
-    private static Scanner sc = new Scanner(System.in);
+    static Scanner sc;
+    static PrintWriter output;
+    static StringWriter outputString;
 
     // For testing without writing JUnit tests
 
     public static void main(String[] args) {
+
+        sc = new Scanner(System.in);
+        output = new PrintWriter(System.out);
 
         while(true) {
             printOptions();
@@ -18,11 +26,24 @@ public class Main {
 
     }
 
+    static void main(Scanner scanner, StringWriter myOutput) {
+        sc = scanner;
+        outputString = myOutput;
+        output = new PrintWriter(outputString);
+
+        while(true) {
+            printOptions();
+            pickOption();
+        }
+    }
+
     private static void printOptions() {
-        System.out.println("Welcome to the vertex coloring library! Please pick an option below by typing its number:");
-        System.out.println("\r\n\t1. Enter your own graph");
-        System.out.println("\r\n\t2. Generate a graph with a specified number of vertices and number of edges");
-        System.out.println("\r\n\t3. Generate a graph with a specified number of vertices and percentage of edges\r\n");
+        output.println("Welcome to the vertex coloring library! Please pick an option below by typing its number:");
+        output.println("\r\n\t1. Enter your own graph");
+        output.println("\r\n\t2. Generate a graph with a specified number of vertices and percentage of edges");
+        output.println("\r\n\t3. Exit program\r\n");
+        output.flush();
+
     }
 
     private static void pickOption() {
@@ -35,47 +56,54 @@ public class Main {
                 enterGraph();
                 break;
             case 2:
-                System.out.println("Oops! Not implemented yet!");
-                break;
-            case 3:
                 generatePercentageGraph();
                 break;
+            case 3:
+                System.exit(0);
+                break;
             default:
-                System.out.println("Please choose an option.");
+                output.println("Please choose an option.");
+                output.flush();
                 break;
         }
     }
 
     private static void generatePercentageGraph() {
-        System.out.println("Please enter the number of vertices you want:");
+        output.println("Please enter the number of vertices you want:");
+        output.flush();
 
         int vertices = sc.nextInt();
[…]

### 2017-02-01 src/test/java/com/devshawn/coloring/ColoringResultSpec.java

@@ -2,7 +2,12 @@ package com.devshawn.coloring;
 
 import org.junit.Test;
 
+import java.io.PrintWriter;
+import java.io.StringWriter;
+
+import static org.hamcrest.CoreMatchers.containsString;
 import static org.junit.Assert.assertEquals;
+import static org.junit.Assert.assertThat;
 import static org.junit.Assert.assertTrue;
 
 public class ColoringResultSpec {
@@ -26,13 +31,14 @@ public class ColoringResultSpec {
 
     @Test
     public void printSummary() {
-        // test for coverage on this function
         int[] results = {0};
         ColoringResult result = new ColoringResult(results, ColoringHeuristic.GREEDY);
-
-        result.printSummary();
-        assertTrue(true);
+        StringWriter writer = new StringWriter();
+        result.printSummary(new PrintWriter(writer));
+        assertThat(writer.toString(), containsString("Colors used: 1"));
     }
+
+
 
 
 }

### 2017-02-01 src/test/java/com/devshawn/coloring/GraphGeneratorSpec.java

@@ -1,6 +1,8 @@
 package com.devshawn.coloring;
 
+import org.junit.Rule;
 import org.junit.Test;
+import org.junit.rules.ExpectedException;
 
 import java.util.Arrays;
 
@@ -8,9 +10,40 @@ import static org.junit.Assert.assertEquals;
 
 public class GraphGeneratorSpec {
 
+    @Rule
+    public ExpectedException thrown = ExpectedException.none();
+
     @Test
     public void simple() {
         int[][] graph = GraphGenerator.simple(1000, 0.3);
+        assertEquals(true, GraphGenerator.isSymmetric(graph));
+    }
+
+    @Test
+    public void badProbability() throws ColoringException {
+        thrown.expect(ColoringException.class);
+        thrown.expectMessage(ColoringException.INVALID_PROBABILITY);
+
+        int[][] graph = GraphGenerator.simple(4, 2);
+    }
+
+    @Test
+    public void isNotSymmetric() {
+        int[][] graph = {
+                {1, 1},
+                {0, 1}
+        };
+
+        assertEquals(false, GraphGenerator.isSymmetric(graph));
+    }
+
+    @Test
+    public void isSymmetric() {
+        int[][] graph = {
+                {0, 1},
+                {1, 0}
+        };
+
         assertEquals(true, GraphGenerator.isSymmetric(graph));
     }
 }

### 2017-02-01 src/test/java/com/devshawn/coloring/MainSpec.java

@@ -1,0 +1,77 @@
+package com.devshawn.coloring;
+
+import org.junit.Rule;
+import org.junit.Test;
+import org.junit.contrib.java.lang.system.Assertion;
+import org.junit.contrib.java.lang.system.ExpectedSystemExit;
+
+import java.io.*;
+import java.util.Scanner;
+
+import static org.hamcrest.CoreMatchers.containsString;
+import static org.junit.Assert.assertEquals;
+import static org.junit.Assert.assertThat;
+
+public class MainSpec {
+
+    @Rule
+    public final ExpectedSystemExit exit = ExpectedSystemExit.none();
+
+    @Test
+    public void testExit() {
+        exit.expectSystemExitWithStatus(0);
+
+        String input = "3\n";
+        StringWriter output = new StringWriter();
+        Main.main(new Scanner(input), output);
+    }
+
+    @Test
+    public void generateGraph() {
+        exit.expectSystemExitWithStatus(0);
+        exit.checkAssertionAfterwards(() -> assertThat(Main.outputString.toString(), containsString("Colors used: 1")));
+
+        String input = "2\n"
+                + "1\n" // 1 vertex
+                + "0\n\n"
+                + "3\n";
+
+        StringWriter output = new StringWriter();
+        Main.main(new Scanner(input), output);
+    }
+
+    @Test
+    public void enterGraph() {
+        exit.expectSystemExitWithStatus(0);
+        exit.checkAssertionAfterwards(() -> {
+            assertThat(Main.outputString.toString(), containsString("Colors used: 2"));
+            assertThat(Main.outputString.toString(), containsString("Color 0 => 1"));
+            assertThat(Main.outputString.toString(), containsString("Color 1 => 1"));
+            assertThat(Main.outputString.toString(), containsString("Vertex 0 => color 0"));
+            assertThat(Main.outputString.toString(), containsString("Vertex 1 => color 1"));
+        });
+
+        String input = "1\n"
+                + "2\n"
+                + "01\n"
+                + "10\n\n"
+                + "3\n";
+
+        StringWriter output = new StringWriter();
+        Main.main(new Scanner(input), output);
+    }
+
+    @Test
+    public void invalidOption() {
+        exit.expectSystemExitWithStatus(0);
+        exit.checkAssertionAfterwards(() -> {
+            assertThat(Main.outputString.toString(), containsString("Please choose an option."));
+        });
+
+        String input = "4\n"
+                + "3\n";
+
[…]

### 2017-02-01 src/main/java/com/devshawn/coloring/Coloring.java
Changing results to show color number at the end

@@ -1,5 +1,7 @@
 package com.devshawn.coloring;
 
+import java.util.Arrays;
+
 public class Coloring {
 
     private int[][] graph;

### 2017-02-01 src/main/java/com/devshawn/coloring/ColoringResult.java

@@ -47,9 +47,6 @@ public class ColoringResult {
 
     public String getSummary() {
         String string = "The graph was colored with the " + heuristicUsed.getName() + " heuristic.";
-        string += "\r\n\tColors used: " + highestColor;
-
-        string += "\r\n\t-----------------------------------";
 
         string += "\r\n\tColor usage:";
         for(int i = 0; i < colors.length; i++) {
@@ -62,6 +59,10 @@ public class ColoringResult {
         for(int i = 0; i < result.length; i++) {
             string += "\r\n\tVertex " + i + " => color " + result[i];
         }
+
+        string += "\r\n\t-----------------------------------";
+
+        string += "\r\n\tColors used: " + highestColor;
 
         return string;
     }

### 2017-02-01 src/main/java/com/devshawn/coloring/Main.java

@@ -60,7 +60,6 @@ public class Main {
                 break;
             case 3:
                 System.exit(0);
-                break;
             default:
                 output.println("Please choose an option.");
                 output.flush();
@@ -84,11 +83,12 @@ public class Main {
 
         int[][] graph = GraphGenerator.simple(vertices, percentage);
 
-        output.println(GraphGenerator.isSymmetric(graph));
-        output.flush();
-
         Coloring coloring = new Coloring();
         coloring.setGraph(graph);
+
+        output.println("Edges generated: " + coloring.getEdgeCount());
+        output.flush();
+
         ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
         result.printSummary(output);
         sc.nextLine();

### 2017-02-01 src/test/java/com/devshawn/coloring/ColoringResultSpec.java

@@ -18,13 +18,13 @@ public class ColoringResultSpec {
         ColoringResult result = new ColoringResult(results, ColoringHeuristic.GREEDY);
 
         String expectedResult = "The graph was colored with the greedy heuristic.";
-        expectedResult += "\r\n\tColors used: 1";
-        expectedResult += "\r\n\t-----------------------------------";
         expectedResult += "\r\n\tColor usage:";
         expectedResult += "\r\n\tColor 0 => 1";
         expectedResult += "\r\n\t-----------------------------------";
         expectedResult += "\r\n\tVertex to color mapping:";
         expectedResult += "\r\n\tVertex 0 => color 0";
+        expectedResult += "\r\n\t-----------------------------------";
+        expectedResult += "\r\n\tColors used: 1";
 
         assertEquals(expectedResult, result.getSummary());
     }

### 2017-02-01 src/main/java/com/devshawn/coloring/Coloring.java
Implementing welsh powell heuristic

@@ -1,6 +1,8 @@
 package com.devshawn.coloring;
 
 import java.util.Arrays;
+import java.util.Collections;
+import java.util.List;
 
 public class Coloring {
 
@@ -17,6 +19,7 @@ public class Coloring {
                 result = applyGreedyHeuristic();
                 break;
             case WELSH_POWELL:
+                result = applyWelshPowell();
                 break;
             case MAXIMAL_INDEPENDENT_SET:
                 break;
@@ -26,6 +29,8 @@ public class Coloring {
     }
 
     // Implemented heuristics
+
+    // Greedy heuristic
     private ColoringResult applyGreedyHeuristic() {
 
         // Make result array of vertices, assign no color to them
@@ -64,6 +69,66 @@ public class Coloring {
 
         // Generate a ColorResult
         return new ColoringResult(result, ColoringHeuristic.GREEDY);
+    }
+
+    // Greedy plus degree sequencing heuristic (Welsh-Powell algorithm)
+    private ColoringResult applyWelshPowell() {
+
+        // Make result array of vertices, assign no color to them
+        int[] result = new int[graph.length];
+        for(int i = 0; i < graph.length; i++) result[i] = -1;
+
+        // Calculate degree of each vertex
+        int[] degrees = new int[graph.length];
+        Integer[] orderedVertices = new Integer[graph.length];
+
+        for(int i = 0; i < graph.length; i++) {
+            orderedVertices[i] = i;
+            for(int j = 0; j < graph.length; j++) {
+                if(graph[i][j] == 1) {
+                    degrees[i]++;
+                }
+            }
+        }
+
+        // Sort our vertices by decreasing degree
+        List<Integer> indexList = Arrays.asList(orderedVertices);
+        Collections.sort(indexList, (left, right) -> (degrees[indexList.indexOf(right)] - degrees[indexList.indexOf(left)]));
+        orderedVertices = indexList.toArray(new Integer[graph.length]);
+        System.out.println(Arrays.toString(orderedVertices));
+
+        // Make temporary available array to see what colors are available
+        boolean available[] = new boolean[graph.length];
+        for(int i = 0; i < graph.length; i++) available[i] = true;
+
+        // Color the graph!
+        for(int i = 0; i < graph.length; i++) {
+
+            // Check neighbors for colors
+            for(int j = 0; j < graph[orderedVertices[i]].length; j++) {
+                if(graph[orderedVertices[i]][j] == 1 && result[j] != -1) {
+                    available[result[j]] = false;
+                }
+            }
+
[…]

### 2017-02-01 src/test/java/com/devshawn/coloring/ColoringSpec.java

@@ -59,56 +59,4 @@ public class ColoringSpec {
         coloring.setGraph(graph);
     }
 
-    // Test greedy heuristic with some known greedy colorings
-
-    @Test
-    public void smallGreedyCompleteGraph() {
-
-        // a complete 3-vertex graph
-        int[][] graph = {
-                {0, 1, 1},
-                {1, 0, 1},
-                {1, 1, 0}
-        };
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
-
-        assertEquals(3, result.getColoringNumber());
-        int[] expectedColoring = {0, 1, 2};
-        assertArrayEquals(expectedColoring, result.getColoring());
-    }
-
-    @Test
-    public void smallGreedyGraph() {
-
-        // a 5-vertex graph
-
-        /*
-         * (0) ----- (2)
-         *  | \     / |
-         *  |   (4)   |
-         *  | /     \ |
-         * (1) ----- (3)
-         *
-         */
-
-        int[][] graph = {
-                {0, 1, 1, 0, 1},
-                {1, 0, 0, 1, 1},
-                {1, 0, 0, 1, 1},
-                {0, 1, 1, 0, 1},
-                {1, 1, 1, 1, 0}
-        };
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
-
-        assertEquals(3, result.getColoringNumber());
-        int[] expectedColoring = {0, 1, 1, 0, 2};
-        assertArrayEquals(expectedColoring, result.getColoring());
-    }
-
 }

### 2017-02-01 src/test/java/com/devshawn/coloring/GreedySpec.java

@@ -1,0 +1,60 @@
+package com.devshawn.coloring;
+
+import org.junit.Test;
+
+import static org.junit.Assert.assertArrayEquals;
+import static org.junit.Assert.assertEquals;
+
+public class GreedySpec {
+
+    @Test
+    public void smallGreedyCompleteGraph() {
+
+        // a complete 3-vertex graph
+        int[][] graph = {
+                {0, 1, 1},
+                {1, 0, 1},
+                {1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
+
+        assertEquals(3, result.getColoringNumber());
+        int[] expectedColoring = {0, 1, 2};
+        assertArrayEquals(expectedColoring, result.getColoring());
+    }
+
+    @Test
+    public void smallGreedyGraph() {
+
+        // a 5-vertex graph
+
+        /*
+         * (0) ----- (2)
+         *  | \     / |
+         *  |   (4)   |
+         *  | /     \ |
+         * (1) ----- (3)
+         *
+         */
+
+        int[][] graph = {
+                {0, 1, 1, 0, 1},
+                {1, 0, 0, 1, 1},
+                {1, 0, 0, 1, 1},
+                {0, 1, 1, 0, 1},
+                {1, 1, 1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
+
+        assertEquals(3, result.getColoringNumber());
+        int[] expectedColoring = {0, 1, 1, 0, 2};
+        assertArrayEquals(expectedColoring, result.getColoring());
+    }
+
+}

### 2017-02-01 src/test/java/com/devshawn/coloring/WelshPowellSpec.java

@@ -1,0 +1,72 @@
+package com.devshawn.coloring;
+
+import org.junit.Test;
+
+import static org.junit.Assert.assertArrayEquals;
+import static org.junit.Assert.assertEquals;
+import static org.junit.Assert.assertNotEquals;
+
+public class WelshPowellSpec {
+
+    @Test
+    public void testWelshPowell() {
+
+        // a 5-vertex graph
+
+        /*
+         * (0) ----- (2)
+         *  | \     / |
+         *  |   (4)   |
+         *  | /     \ |
+         * (1) ----- (3)
+         *
+         */
+
+        int[][] graph = {
+                {0, 1, 1, 0, 1},
+                {1, 0, 0, 1, 1},
+                {1, 0, 0, 1, 1},
+                {0, 1, 1, 0, 1},
+                {1, 1, 1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.WELSH_POWELL);
+
+        assertEquals(3, result.getColoringNumber());
+        int[] expectedColoring = {1, 2, 2, 1, 0};
+        assertArrayEquals(expectedColoring, result.getColoring());
+    }
+
+    @Test
+    public void producesDifferentThanGreedy() {
+
+        // a 5-vertex graph
+
+        /*
+         * (0) ----- (2)
+         *  | \     / |
+         *  |   (4)   |
+         *  | /     \ |
+         * (1) ----- (3)
+         *
+         */
+
+        int[][] graph = {
+                {0, 1, 1, 0, 1},
+                {1, 0, 0, 1, 1},
+                {1, 0, 0, 1, 1},
+                {0, 1, 1, 0, 1},
+                {1, 1, 1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+        ColoringResult greedyResult = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
+        ColoringResult welshPowellResult = coloring.applyHeuristic(ColoringHeuristic.WELSH_POWELL);
+        for(int i = 0; i < greedyResult.getColoring().length; i++) {
+            assertNotEquals(greedyResult.getColoring()[i], welshPowellResult.getColoring()[i]);
+        }
+    }
+}

### 2017-02-01 src/main/java/com/devshawn/coloring/Coloring.java
Starting to implement MIS

@@ -1,8 +1,6 @@
 package com.devshawn.coloring;
 
-import java.util.Arrays;
-import java.util.Collections;
-import java.util.List;
+import java.util.*;
 
 public class Coloring {
 
@@ -22,6 +20,7 @@ public class Coloring {
                 result = applyWelshPowell();
                 break;
             case MAXIMAL_INDEPENDENT_SET:
+                result = applyMaximalIndependentSet();
                 break;
         }
 
@@ -129,6 +128,58 @@ public class Coloring {
 
 
         return new ColoringResult(result, ColoringHeuristic.WELSH_POWELL);
+    }
+
+    private ColoringResult applyMaximalIndependentSet() {
+
+        // Make result array of vertices, assign no color to them
+        int[] result = new int[graph.length];
+        for(int i = 0; i < graph.length; i++) result[i] = -1;
+
+        // Copy our graph so we can manipulate it
+        int[][] newGraph = new int[graph.length][graph.length];
+
+        for(int i = 0; i < graph.length; i++) {
+            for(int j = 0; j <graph.length; j++) {
+                newGraph[i][j] = graph[i][j];
+            }
+        }
+
+        Set<Integer> vertices = new HashSet<>();
+        for(int i = 0; i < graph.length; i++) {
+            vertices.add(i);
+        }
+
+        // Color using MIS heuristic
+        int color = 0;
+        while(!vertices.isEmpty()) {
+            Set<Integer> tempVertices = new HashSet<>(vertices);
+            Set<Integer> set = new HashSet<>();
+
+            // Find a maximal independent set
+            while(!tempVertices.isEmpty()) {
+                Integer i = tempVertices.iterator().next();
+                set.add(i);
+                vertices.remove(i);
+                tempVertices.remove(i);
+
+                for(int k = 0; k < graph.length; k++) {
+                    if(graph[i][k] == 1) {
+                        tempVertices.remove(k);
+
+                    }
+                }
+            }
+
+            // Color the MIS with the next color
+            for(int i : set) {
+                result[i] = color;
+            }
+
+            color++;
+        }
+
+        return new ColoringResult(result, ColoringHeuristic.MAXIMAL_INDEPENDENT_SET);
     }
 
     // Simple functions to get information about the graph

### 2017-02-01 src/test/java/com/devshawn/coloring/MaximalIndependentSetSpec.java

@@ -1,0 +1,40 @@
+package com.devshawn.coloring;
+
+import org.junit.Test;
+
+import static org.junit.Assert.*;
+
+public class MaximalIndependentSetSpec {
+
+    @Test
+    public void testMIS() {
+
+        // a 5-vertex graph
+
+        /*
+         * (0) ----- (2)
+         *  | \     / |
+         *  |   (4)   |
+         *  | /     \ |
+         * (1) ----- (3)
+         *
+         */
+
+        int[][] graph = {
+                {0, 1, 1, 0, 1},
+                {1, 0, 0, 1, 1},
+                {1, 0, 0, 1, 1},
+                {0, 1, 1, 0, 1},
+                {1, 1, 1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.MAXIMAL_INDEPENDENT_SET);
+
+        assertEquals(3, result.getColoringNumber());
+        int[] expectedColoring = {0, 1, 1, 0, 2};
+        assertArrayEquals(expectedColoring, result.getColoring());
+    }
+
+}

### 2017-02-06 work/work.tex
Starting complexity write up

--- 2017-01-31 work/work.tex
@@
 
-\newtheorem*{definition}{Definition}
 \newtheorem{prop}{Proposition}
 \newtheorem*{theorem}{Theorem}
+
+\theoremstyle{definition}
+\newtheorem*{definition}{Definition}
 
@@
 
+\section*{Vertex Coloring}
+The vertex coloring problem (VCP) is one of the most common and most researched problems in graph theory.
+Vertex coloring is an assignment of colors, or labels, to each vertex in a graph \(G = (V, E)\) such that no edge connects two vertices of the same color.
+The vertex coloring problem aims to find the chromatic number of a graph, denoted \(\chi(G)\), the minimum number of colors needed to properly color a graph as defined above.
+The \emph{k-Colorability} problem asks where a graph can be colored with \(k\) colors.
+
+\begin{prop}
+Vertex coloring is a NP-hard problem.
+\end{prop}
+
+Coloring a graph is computationally complex. It is NP-hard to compute the chromatic number of a given graph \cite{garey}.
+
+\begin{prop}
+The k-Colorability problem, i.e. can a given graph be colored in k-colors, is a NP-complete problem.
+\end{prop}
+
+To prove this, we need to reduce a known NP-complete problem to our coloring problem in polynomial time.
+By reducing a known NP-complete problem to k-Colorability, we can conclude that k-Colorability is NP-complete as all NP-complete problems are redicible to all other NP-complete problems \cite{gareynp}.
+We will reduce the well-known 3-SAT problem first to the 3-Colorability problem, and then from there reduce to the k-Colorability problem.
+
+\subsection*{3-Satisfiability (3-SAT)}
+The 3-SAT problem is an NP-complete problem that is a reduction of the original NP-complete problem, satisfiability (SAT).
+The 3-SAT problem is defined and proven to be NP-complete in \cite{gareynp}.
+3-SAT consists of a conjunction of a number of clauses.
+A clause is a disjunction of a given number of propositions or their negations.
+We will assume all of our satisfiability problems are in conjunctive normal form (CNF).
+3-SAT asks whether the variables can be assigned \emph{true} or \emph{false} in such a way that the whole expression evaluates to \emph{true}.
+
+\begin{definition}
+SAT is a conjunction of $m$ clauses, i.e.
+$C_1 \wedge C_2 \wedge \dots \wedge C_m$.
+Each clause is a disjunction of at most $n$ literals, i.e.
+$L_1 \vee L_2 \vee \dots \vee L_n$.
+Each literal can be a variable or negative of a variable, i.e.
+$x_1, \neg x_1, x_2, \neg x_2, \dots , x_k, \neg x_k$ where $k$ is the number of distinct literals.
+Each literal has a value of $true$ or $false$.
+3-SAT is where each clause has exactly 3 literals \((n = 3)\).
+\end{definition}
+
+\subsection*{Reducing 3-SAT to 3-Colorability}
+3-Colorability is NP-complete and was proven by \cite{moret}. To show this reduction, we need to:
+
+\begin{enumerate}
+\item Given an instance of 3-SAT, construct an instance of 3-Colorability
+\item Show that the graph is 3-colorable, then the given 3-SAT problem is satisfied
+\end{enumerate}
+
+To show how this reduction can be done, I'll create an example 3-SAT problem and reduce it to 3-Colorability. Let's take the following 3-SAT expression:
+
+\begin{align*}
+\left( x_1 \vee \neg x_2 \vee \neg x_3 \right) \wedge \left( \neg x_1 \vee x_2 \vee x_3 \right) \wedge \left( x_1 \vee x_2 \vee x_3 \right)
+\end{align*}
+
 \section*{Heuristics}
-In this report, I will discuss multiple vertex coloring heuristics. The input for all heuristics defined here is a simple graph \(G = (V, E)\). I will define them as follows:
+As shown above, solving the VCP is very computationally heavy.
+There are heuristics that give a good, but not necessarily optimal, coloring of a given graph in polynomial time.
+I will be analyzing and comparing a few of the popular heuristics for solving the VCP.
+The input for all heuristics defined here is a simple graph \(G = (V, E)\).
+I will define them as follows:
 
@@
 
+\bibitem{garey}
+Garey, M. R., Johnson, D. S., and Stockmeyer, L. [1976], \emph{Some simplified NP-complete graph problems}, Theoretical computer science, 1(3), 237-267.
+
+\bibitem{gareynp}
+Garey, M.R. and Johnson, D.S. [2002], \emph{Computers and intractability} (Vol. 29), New York: wh freeman.
+
 \bibitem{kordecki}
@@
 
+\bibitem{moret}
+Moret, B.M. [1998], \emph{The theory of computation}, Addison-Wesley, Reading, Mass.
+
+\bibitem{sharma}
+Sharma, P. C. and Chaudhari, N.S. [2012], \emph{A new reduction from 3-SAT to graph k-colorability for frequency assignment problem}, Int. J. Comp. Applic, 23-27.
+
 \bibitem{welsh}

### 2017-02-07 work/work.tex
Starting 3sat to 3color example

@@
 
-To show how this reduction can be done, I'll create an example 3-SAT problem and reduce it to 3-Colorability. Let's take the following 3-SAT expression:
+To show how this reduction can be done, I'll create an example 3-SAT problem and reduce it to 3-Colorability.
+This reduction will create a graph \(G = (V, E)\) having \((2n + 3m + 1)\) vertices and \((3n+6m)\) edges \cite{moret}.
+The process for creating the graph, described below, was taken from \cite{sharma}.\newline
+
+Let's take the following 3-SAT expression:
 
 \begin{align*}
-\left( x_1 \vee \neg x_2 \vee \neg x_3 \right) \wedge \left( \neg x_1 \vee x_2 \vee x_3 \right) \wedge \left( x_1 \vee x_2 \vee x_3 \right)
+E = \left( x_1 \vee \neg x_2 \vee \neg x_3 \right) \wedge \left( \neg x_1 \vee x_2 \vee x_3 \right) \wedge \left( x_1 \vee x_2 \vee x_3 \right)
 \end{align*}
+
+\subsubsection*{Creating the graph}
+\begin{enumerate}
+\item Create a \emph{triangle} for each variable, i.e.
+\(x_i\).
+Each has a common vertex \(B\), called the base vertex.
+This base vertex will be colored a color, say \(p_1\), so that the other two vertices correspond to the variable and it's negation.
+\item Create a \emph{triangle} for each clause in the formula. 
+\item Connect each vertex of a clause triangle to the corresponding literal vertex.
+\end{enumerate}
+
+Following the above steps, we generate the following graph \(G\) from our 3-SAT expression \(E\):
 

### 2017-02-07 work/work.tex
Adding 3-sat to 3-color example

@@
 Following the above steps, we generate the following graph \(G\) from our 3-SAT expression \(E\):
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.6]{images/3sat-1.png}
+\caption{\(E\) represented as a graph \(G\)}
+\end{figure}
+
+If we can properly color graph \(G\) with 3 colors, then the original 3-SAT is satisfiable and the truth values of the variables are denoted by the color they received.
+This graph is \emph{3-colorable} and the coloring is shown below.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.6]{images/3sat-2.png}
+\caption{A possible 3-coloring of \(G\)}
+\end{figure}
+
+As we see above, \emph{blue} is used for the base vertex and as part of some of the clause triangles.
+If we let \emph{yellow} denote \emph{true} and \emph{red} denote \emph{false}, we can see that our original 3-SAT is satisfied:
+
+\begin{align*}
+E &= \left( T \vee \neg T \vee \neg F \right) \wedge \left( \neg T \vee T \vee F \right) \wedge \left(T \vee T \vee F \right) \\
+&= \left( T \vee F \vee T \right) \wedge \left( F \vee T \vee F \right) \wedge \left(T \vee T \vee F \right) \\
+&= T \wedge T \wedge T \\
+&= True
+\end{align*}
+
+This shows an example of the reduction from 3-SAT to 3-Colorability. If \(G\) was unable to be colored with 3 colors, then we would know that \(E\) is unable to be satisfied.
 

### 2017-02-13 library/src/main/java/com/devshawn/coloring/library/Coloring.java
Starting visualization work

@@ -1,0 +1,309 @@
+package com.devshawn.coloring.library;
+
+import java.util.*;
+
+public class Coloring {
+
+    private int[][] graph;
+    private int vertexCount;
+    private int edgeCount;
+
+    // Apply a given heuristic to the graph
+    public ColoringResult applyHeuristic(ColoringHeuristic heuristic) {
+        ColoringResult result = new ColoringResult();
+
+        switch(heuristic) {
+            case GREEDY:
+                result = applyGreedyHeuristic();
+                break;
+            case WELSH_POWELL:
+                result = applyWelshPowell();
+                break;
+            case MAXIMAL_INDEPENDENT_SET:
+                result = applyMaximalIndependentSet();
+                break;
+            case DSATUR:
+                result = applyDSATUR();
+                break;
+        }
+
+        return result;
+    }
+
+    // Implemented heuristics
+
+    // Greedy heuristic
+    private ColoringResult applyGreedyHeuristic() {
+
+        // Make result array of vertices, assign no color to them
+        int[] result = new int[graph.length];
+        for(int i = 0; i < graph.length; i++) result[i] = -1;
+
+        // Make temporary available array to see what colors are available
+        boolean available[] = new boolean[graph.length];
+        for(int i = 0; i < graph.length; i++) available[i] = true;
+
+        // Color the graph!
+        for(int i = 0; i < graph.length; i++) {
+
+            // Check neighbors for colors
+            for(int j = 0; j < graph[i].length; j++) {
+                if(graph[i][j] == 1 && result[j] != -1) {
+                    available[result[j]] = false;
+                }
+            }
+
+            // Find smallest available color
+            int color = -1;
+            for(int j = 0; j < available.length; j++) {
+                if(available[j]) {
+                    color = j;
+                    break;
+                }
+            }
+
+            // Set this vertex to that color
+            result[i] = color;
+
+            // Reset our available colors
+            for(int j = 0; j < graph.length; j++) available[j] = true;
+        }
+
+        // Generate a ColorResult
+        return new ColoringResult(result, ColoringHeuristic.GREEDY);
+    }
+
+    // Greedy plus degree sequencing heuristic (Welsh-Powell algorithm)
+    private ColoringResult applyWelshPowell() {
+
+        // Make result array of vertices, assign no color to them
[…]

### 2017-02-13 library/src/main/java/com/devshawn/coloring/library/ColoringException.java

@@ -1,0 +1,13 @@
+package com.devshawn.coloring.library;
+
+public class ColoringException extends RuntimeException {
+
+    public static String INVALID_MATRIX = "Your adjacency matrix is invalid.";
+    public static String SELF_LOOPS = "Your graph is not a simple graph. It contains self-loops.";
+    public static String INVALID_PROBABILITY = "Probability for generating a graph must be between 0 and 1.";
+
+    public ColoringException (String message) {
+        super(message);
+    }
+
+}

### 2017-02-13 library/src/main/java/com/devshawn/coloring/library/ColoringResult.java

@@ -1,0 +1,73 @@
+package com.devshawn.coloring.library;
+
+import java.io.PrintWriter;
+
+public class ColoringResult {
+    private int[] result;
+    private int highestColor;
+    private ColoringHeuristic heuristicUsed;
+    private int[] colors;
+
+    public ColoringResult() {
+
+    }
+
+    public ColoringResult(int[] result, ColoringHeuristic heuristicUsed) {
+
+        // Find how many colors we used
+        int highestColor = -1;
+
+        for(int i = 0; i < result.length; i++) {
+            if(result[i] > highestColor) {
+                highestColor = result[i];
+            }
+        }
+
+        // Add 1 as colors start with 0
+        this.highestColor = highestColor + 1;
+
+        // Count our color usage
+        this.colors = new int[this.highestColor];
+        for(int i = 0; i < result.length; i++) {
+            colors[result[i]]++;
+        }
+
+        this.result = result;
+        this.heuristicUsed = heuristicUsed;
+    }
+
+    public int getColoringNumber() {
+        return highestColor;
+    }
+
+    public int[] getColoring() {
+        return this.result;
+    }
+
+    public String getSummary() {
+        String string = "The graph was colored with the " + heuristicUsed.getName() + " heuristic.";
+
+        string += "\r\n\tColor usage:";
+        for(int i = 0; i < colors.length; i++) {
+            string += "\r\n\tColor " + i + " => " + colors[i];
+        }
+
+        string += "\r\n\t-----------------------------------";
+
+        string += "\r\n\tVertex to color mapping:";
+        for(int i = 0; i < result.length; i++) {
+            string += "\r\n\tVertex " + i + " => color " + result[i];
+        }
+
+        string += "\r\n\t-----------------------------------";
+
+        string += "\r\n\tColors used: " + highestColor;
+
+        return string;
+    }
+
+    public void printSummary(PrintWriter output) {
+        output.println(getSummary());
+        output.flush();
+    }
+}

### 2017-02-13 library/src/main/java/com/devshawn/coloring/library/Main.java

@@ -1,0 +1,123 @@
+package com.devshawn.coloring.library;
+
+import java.io.PrintWriter;
+import java.io.StringWriter;
+import java.util.Scanner;
+
+public class Main {
+
+    static Scanner sc;
+    static PrintWriter output;
+    static StringWriter outputString;
+
+    // For testing without writing JUnit tests
+
+    public static void main(String[] args) {
+
+        sc = new Scanner(System.in);
+        output = new PrintWriter(System.out);
+
+        while(true) {
+            printOptions();
+            pickOption();
+        }
+
+    }
+
+    static void main(Scanner scanner, StringWriter myOutput) {
+        sc = scanner;
+        outputString = myOutput;
+        output = new PrintWriter(outputString);
+
+        while(true) {
+            printOptions();
+            pickOption();
+        }
+    }
+
+    private static void printOptions() {
+        output.println("Welcome to the vertex coloring library! Please pick an option below by typing its number:");
+        output.println("\r\n\t1. Enter your own graph");
+        output.println("\r\n\t2. Generate a graph with a specified number of vertices and percentage of edges");
+        output.println("\r\n\t3. Exit program\r\n");
+        output.flush();
+
+    }
+
+    private static void pickOption() {
+        int option = sc.nextInt();
+        sc.nextLine();
+
+
+        switch(option) {
+            case 1:
+                enterGraph();
+                break;
+            case 2:
+                generatePercentageGraph();
+                break;
+            case 3:
+                System.exit(0);
+            default:
+                output.println("Please choose an option.");
+                output.flush();
+                break;
+        }
+    }
+
+    private static void generatePercentageGraph() {
+        output.println("Please enter the number of vertices you want:");
+        output.flush();
+
+        int vertices = sc.nextInt();
+        sc.nextLine();
+
+        output.println("Please enter the percentage at which an edge should be placed (i.e. 30):");
+        output.flush();
+
+        double percentage = sc.nextDouble();
+        percentage = percentage / (double) 100;
+        sc.nextLine();
+
+        int[][] graph = GraphGenerator.simple(vertices, percentage);
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+
+        output.println("Edges generated: " + coloring.getEdgeCount());
+        output.flush();
+
+        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
[…]

### 2017-02-13 library/src/test/java/com/devshawn/coloring/library/ColoringHeuristicSpec.java

@@ -1,0 +1,14 @@
+package com.devshawn.coloring.library;
+
+import org.junit.Test;
+
+import static org.junit.Assert.assertEquals;
+
+public class ColoringHeuristicSpec {
+
+    @Test
+    public void getName() {
+        ColoringHeuristic heuristic = ColoringHeuristic.GREEDY;
+        assertEquals("greedy", heuristic.getName());
+    }
+}

### 2017-02-13 library/src/test/java/com/devshawn/coloring/library/ColoringResultSpec.java

@@ -1,0 +1,45 @@
+package com.devshawn.coloring.library;
+
+import org.junit.Test;
+
+import java.io.PrintWriter;
+import java.io.StringWriter;
+
+import static org.hamcrest.CoreMatchers.containsString;
+import static org.junit.Assert.assertEquals;
+import static org.junit.Assert.assertThat;
+import static org.junit.Assert.assertTrue;
+
+public class ColoringResultSpec {
+
+    @Test
+    public void getSummary() {
+        int[] results = {0};
+        ColoringResult result = new ColoringResult(results, ColoringHeuristic.GREEDY);
+
+        String expectedResult = "The graph was colored with the greedy heuristic.";
+        expectedResult += "\r\n\tColor usage:";
+        expectedResult += "\r\n\tColor 0 => 1";
+        expectedResult += "\r\n\t-----------------------------------";
+        expectedResult += "\r\n\tVertex to color mapping:";
+        expectedResult += "\r\n\tVertex 0 => color 0";
+        expectedResult += "\r\n\t-----------------------------------";
+        expectedResult += "\r\n\tColors used: 1";
+
+        assertEquals(expectedResult, result.getSummary());
+    }
+
+    @Test
+    public void printSummary() {
+        int[] results = {0};
+        ColoringResult result = new ColoringResult(results, ColoringHeuristic.GREEDY);
+        StringWriter writer = new StringWriter();
+        result.printSummary(new PrintWriter(writer));
+        assertThat(writer.toString(), containsString("Colors used: 1"));
+    }
+
+
+
+
+}
+

### 2017-02-13 library/src/test/java/com/devshawn/coloring/library/ColoringSpec.java

@@ -1,0 +1,62 @@
+package com.devshawn.coloring.library;
+
+import org.junit.Rule;
+import org.junit.Test;
+import org.junit.rules.ExpectedException;
+
+import static org.junit.Assert.assertArrayEquals;
+import static org.junit.Assert.assertEquals;
+import static org.junit.Assert.assertTrue;
+
+public class ColoringSpec {
+
+    @Rule
+    public ExpectedException thrown = ExpectedException.none();
+
+    // Tests for setting up
+
+    @Test
+    public void validGraphSetup() {
+        int[][] graph = {
+                {0, 1, 1},
+                {1, 0, 1},
+                {1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+        assertTrue(graph == coloring.getGraph());
+        assertEquals(3, coloring.getVertexCount());
+        assertEquals(3, coloring.getEdgeCount());
+
+    }
+
+    @Test
+    public void invalidGraphSet() throws ColoringException {
+        thrown.expect(ColoringException.class);
+        thrown.expectMessage(ColoringException.INVALID_MATRIX);
+
+        int[][] graph = {
+                {1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+    }
+
+    @Test
+    public void nonSimpleGraph() throws ColoringException {
+        thrown.expect(ColoringException.class);
+        thrown.expectMessage(ColoringException.SELF_LOOPS);
+
+        int[][] graph = {
+                {0, 1, 1},
+                {1, 1, 1},
+                {1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+    }
+
+}

### 2017-02-13 library/src/test/java/com/devshawn/coloring/library/DSATURSpec.java

@@ -1,0 +1,39 @@
+package com.devshawn.coloring.library;
+
+import org.junit.Test;
+
+import java.util.Arrays;
+
+import static org.junit.Assert.assertArrayEquals;
+import static org.junit.Assert.assertEquals;
+
+public class DSATURSpec {
+    @Test
+    public void smallGreedyGraph() {
+
+        // a 5-vertex graph
+
+        /*
+         * (0) ----- (2)
+         *  | \     / |
+         *  |   (4)   |
+         *  | /     \ |
+         * (1) ----- (3)
+         *
+         */
+
+        int[][] graph = {
+                {0, 1, 1, 0, 1},
+                {1, 0, 0, 1, 1},
+                {1, 0, 0, 1, 1},
+                {0, 1, 1, 0, 1},
+                {1, 1, 1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.DSATUR);
+
+        System.out.println(Arrays.toString(result.getColoring()));
+    }
+}

### 2017-02-13 library/src/test/java/com/devshawn/coloring/library/GreedySpec.java

@@ -1,0 +1,60 @@
+package com.devshawn.coloring.library;
+
+import org.junit.Test;
+
+import static org.junit.Assert.assertArrayEquals;
+import static org.junit.Assert.assertEquals;
+
+public class GreedySpec {
+
+    @Test
+    public void smallGreedyCompleteGraph() {
+
+        // a complete 3-vertex graph
+        int[][] graph = {
+                {0, 1, 1},
+                {1, 0, 1},
+                {1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
+
+        assertEquals(3, result.getColoringNumber());
+        int[] expectedColoring = {0, 1, 2};
+        assertArrayEquals(expectedColoring, result.getColoring());
+    }
+
+    @Test
+    public void smallGreedyGraph() {
+
+        // a 5-vertex graph
+
+        /*
+         * (0) ----- (2)
+         *  | \     / |
+         *  |   (4)   |
+         *  | /     \ |
+         * (1) ----- (3)
+         *
+         */
+
+        int[][] graph = {
+                {0, 1, 1, 0, 1},
+                {1, 0, 0, 1, 1},
+                {1, 0, 0, 1, 1},
+                {0, 1, 1, 0, 1},
+                {1, 1, 1, 1, 0}
+        };
+
+        Coloring coloring = new Coloring();
+        coloring.setGraph(graph);
+        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
+
+        assertEquals(3, result.getColoringNumber());
+        int[] expectedColoring = {0, 1, 1, 0, 2};
+        assertArrayEquals(expectedColoring, result.getColoring());
+    }
+
+}

### 2017-02-13 server/src/main/resources/static/components/app.routes.js

@@ -1,0 +1,7 @@
+angular.module('coloring')
+    .config(function ($stateProvider) {
+        $stateProvider.state("404", {
+            url: "*path",
+            templateUrl: "components/shared/404.html"
+        });
+    });

### 2017-02-13 server/src/main/resources/static/components/graphs/graphs.routes.js

@@ -1,0 +1,13 @@
+angular.module('coloring')
+    .config(function($stateProvider) {
+
+        var graphsState = {
+            name: "graphs",
+            url: "/graphs",
+            templateUrl: "components/graphs/graphs.html",
+            controller: "GraphsController",
+            controllerAs: "graphsController"
+        };
+
+        $stateProvider.state(graphsState);
+    });

### 2017-02-13 server/src/main/resources/static/components/graphs/graphs.service.js

@@ -1,0 +1,6 @@
+angular.module('coloring')
+    .factory('GraphService', function($resource) {
+        return new $resource('/api/graphs/:id', null, {
+            'update': { method: 'PUT'}
+        });
+    });

### 2017-02-13 server/src/main/resources/static/components/home/home.routes.js

@@ -1,0 +1,13 @@
+angular.module('coloring')
+    .config(function($stateProvider) {
+
+        var homeState = {
+            name: "home",
+            url: "/",
+            templateUrl: "components/home/home.html",
+            controller: "HomeController",
+            controllerAs: "homeController"
+        };
+
+        $stateProvider.state(homeState);
+    });

### 2017-02-13 server/src/main/resources/static/components/navbar/navbar.controller.js

@@ -1,0 +1,12 @@
+angular.module('coloring')
+    .controller('NavbarController', function ($scope, $location) {
+        var self = this;
+        self.isActive = isActive;
+
+        $scope.test = "hello";
+
+        function isActive(viewLocation) {
+            return (viewLocation == "/" && $location.url() != "/") ? false : $location.path().indexOf(viewLocation) == 0;
+        }
+
+    });

### 2017-02-13 src/main/java/com/devshawn/coloring/Coloring.java

@@ -1,309 +1,0 @@
-package com.devshawn.coloring;
-
-import java.util.*;
-
-public class Coloring {
-
-    private int[][] graph;
-    private int vertexCount;
-    private int edgeCount;
-
-    // Apply a given heuristic to the graph
-    public ColoringResult applyHeuristic(ColoringHeuristic heuristic) {
-        ColoringResult result = new ColoringResult();
-
-        switch(heuristic) {
-            case GREEDY:
-                result = applyGreedyHeuristic();
-                break;
-            case WELSH_POWELL:
-                result = applyWelshPowell();
-                break;
-            case MAXIMAL_INDEPENDENT_SET:
-                result = applyMaximalIndependentSet();
-                break;
-            case DSATUR:
-                result = applyDSATUR();
-                break;
-        }
-
-        return result;
-    }
-
-    // Implemented heuristics
-
-    // Greedy heuristic
-    private ColoringResult applyGreedyHeuristic() {
-
-        // Make result array of vertices, assign no color to them
-        int[] result = new int[graph.length];
-        for(int i = 0; i < graph.length; i++) result[i] = -1;
-
-        // Make temporary available array to see what colors are available
-        boolean available[] = new boolean[graph.length];
-        for(int i = 0; i < graph.length; i++) available[i] = true;
-
-        // Color the graph!
-        for(int i = 0; i < graph.length; i++) {
-
-            // Check neighbors for colors
-            for(int j = 0; j < graph[i].length; j++) {
-                if(graph[i][j] == 1 && result[j] != -1) {
-                    available[result[j]] = false;
-                }
-            }
-
-            // Find smallest available color
-            int color = -1;
-            for(int j = 0; j < available.length; j++) {
-                if(available[j]) {
-                    color = j;
-                    break;
-                }
-            }
-
-            // Set this vertex to that color
-            result[i] = color;
-
-            // Reset our available colors
-            for(int j = 0; j < graph.length; j++) available[j] = true;
-        }
-
-        // Generate a ColorResult
-        return new ColoringResult(result, ColoringHeuristic.GREEDY);
-    }
-
-    // Greedy plus degree sequencing heuristic (Welsh-Powell algorithm)
-    private ColoringResult applyWelshPowell() {
-
-        // Make result array of vertices, assign no color to them
-        int[] result = new int[graph.length];
[…]

### 2017-02-13 src/main/java/com/devshawn/coloring/Main.java

@@ -1,125 +1,0 @@
-package com.devshawn.coloring;
-
-import java.io.PrintStream;
-import java.io.PrintWriter;
-import java.io.StringWriter;
-import java.util.Arrays;
-import java.util.Scanner;
-
-public class Main {
-
-    static Scanner sc;
-    static PrintWriter output;
-    static StringWriter outputString;
-
-    // For testing without writing JUnit tests
-
-    public static void main(String[] args) {
-
-        sc = new Scanner(System.in);
-        output = new PrintWriter(System.out);
-
-        while(true) {
-            printOptions();
-            pickOption();
-        }
-
-    }
-
-    static void main(Scanner scanner, StringWriter myOutput) {
-        sc = scanner;
-        outputString = myOutput;
-        output = new PrintWriter(outputString);
-
-        while(true) {
-            printOptions();
-            pickOption();
-        }
-    }
-
-    private static void printOptions() {
-        output.println("Welcome to the vertex coloring library! Please pick an option below by typing its number:");
-        output.println("\r\n\t1. Enter your own graph");
-        output.println("\r\n\t2. Generate a graph with a specified number of vertices and percentage of edges");
-        output.println("\r\n\t3. Exit program\r\n");
-        output.flush();
-
-    }
-
-    private static void pickOption() {
-        int option = sc.nextInt();
-        sc.nextLine();
-
-
-        switch(option) {
-            case 1:
-                enterGraph();
-                break;
-            case 2:
-                generatePercentageGraph();
-                break;
-            case 3:
-                System.exit(0);
-            default:
-                output.println("Please choose an option.");
-                output.flush();
-                break;
-        }
-    }
-
-    private static void generatePercentageGraph() {
-        output.println("Please enter the number of vertices you want:");
-        output.flush();
-
-        int vertices = sc.nextInt();
-        sc.nextLine();
-
-        output.println("Please enter the percentage at which an edge should be placed (i.e. 30):");
-        output.flush();
-
-        double percentage = sc.nextDouble();
-        percentage = percentage / (double) 100;
-        sc.nextLine();
-
-        int[][] graph = GraphGenerator.simple(vertices, percentage);
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-
-        output.println("Edges generated: " + coloring.getEdgeCount());
-        output.flush();
-
[…]

### 2017-02-13 src/test/java/com/devshawn/coloring/ColoringResultSpec.java

@@ -1,45 +1,0 @@
-package com.devshawn.coloring;
-
-import org.junit.Test;
-
-import java.io.PrintWriter;
-import java.io.StringWriter;
-
-import static org.hamcrest.CoreMatchers.containsString;
-import static org.junit.Assert.assertEquals;
-import static org.junit.Assert.assertThat;
-import static org.junit.Assert.assertTrue;
-
-public class ColoringResultSpec {
-
-    @Test
-    public void getSummary() {
-        int[] results = {0};
-        ColoringResult result = new ColoringResult(results, ColoringHeuristic.GREEDY);
-
-        String expectedResult = "The graph was colored with the greedy heuristic.";
-        expectedResult += "\r\n\tColor usage:";
-        expectedResult += "\r\n\tColor 0 => 1";
-        expectedResult += "\r\n\t-----------------------------------";
-        expectedResult += "\r\n\tVertex to color mapping:";
-        expectedResult += "\r\n\tVertex 0 => color 0";
-        expectedResult += "\r\n\t-----------------------------------";
-        expectedResult += "\r\n\tColors used: 1";
-
-        assertEquals(expectedResult, result.getSummary());
-    }
-
-    @Test
-    public void printSummary() {
-        int[] results = {0};
-        ColoringResult result = new ColoringResult(results, ColoringHeuristic.GREEDY);
-        StringWriter writer = new StringWriter();
-        result.printSummary(new PrintWriter(writer));
-        assertThat(writer.toString(), containsString("Colors used: 1"));
-    }
-
-
-
-
-}
-

### 2017-02-13 src/test/java/com/devshawn/coloring/ColoringSpec.java

@@ -1,62 +1,0 @@
-package com.devshawn.coloring;
-
-import org.junit.Rule;
-import org.junit.Test;
-import org.junit.rules.ExpectedException;
-
-import static org.junit.Assert.assertArrayEquals;
-import static org.junit.Assert.assertEquals;
-import static org.junit.Assert.assertTrue;
-
-public class ColoringSpec {
-
-    @Rule
-    public ExpectedException thrown = ExpectedException.none();
-
-    // Tests for setting up
-
-    @Test
-    public void validGraphSetup() {
-        int[][] graph = {
-                {0, 1, 1},
-                {1, 0, 1},
-                {1, 1, 0}
-        };
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        assertTrue(graph == coloring.getGraph());
-        assertEquals(3, coloring.getVertexCount());
-        assertEquals(3, coloring.getEdgeCount());
-
-    }
-
-    @Test
-    public void invalidGraphSet() throws ColoringException {
-        thrown.expect(ColoringException.class);
-        thrown.expectMessage(ColoringException.INVALID_MATRIX);
-
-        int[][] graph = {
-                {1, 1, 0}
-        };
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-    }
-
-    @Test
-    public void nonSimpleGraph() throws ColoringException {
-        thrown.expect(ColoringException.class);
-        thrown.expectMessage(ColoringException.SELF_LOOPS);
-
-        int[][] graph = {
-                {0, 1, 1},
-                {1, 1, 1},
-                {1, 1, 0}
-        };
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-    }
-
-}

### 2017-02-13 src/test/java/com/devshawn/coloring/GreedySpec.java

@@ -1,60 +1,0 @@
-package com.devshawn.coloring;
-
-import org.junit.Test;
-
-import static org.junit.Assert.assertArrayEquals;
-import static org.junit.Assert.assertEquals;
-
-public class GreedySpec {
-
-    @Test
-    public void smallGreedyCompleteGraph() {
-
-        // a complete 3-vertex graph
-        int[][] graph = {
-                {0, 1, 1},
-                {1, 0, 1},
-                {1, 1, 0}
-        };
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
-
-        assertEquals(3, result.getColoringNumber());
-        int[] expectedColoring = {0, 1, 2};
-        assertArrayEquals(expectedColoring, result.getColoring());
-    }
-
-    @Test
-    public void smallGreedyGraph() {
-
-        // a 5-vertex graph
-
-        /*
-         * (0) ----- (2)
-         *  | \     / |
-         *  |   (4)   |
-         *  | /     \ |
-         * (1) ----- (3)
-         *
-         */
-
-        int[][] graph = {
-                {0, 1, 1, 0, 1},
-                {1, 0, 0, 1, 1},
-                {1, 0, 0, 1, 1},
-                {0, 1, 1, 0, 1},
-                {1, 1, 1, 1, 0}
-        };
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
-
-        assertEquals(3, result.getColoringNumber());
-        int[] expectedColoring = {0, 1, 1, 0, 2};
-        assertArrayEquals(expectedColoring, result.getColoring());
-    }
-
-}

### 2017-02-13 src/test/java/com/devshawn/coloring/MainSpec.java

@@ -1,77 +1,0 @@
-package com.devshawn.coloring;
-
-import org.junit.Rule;
-import org.junit.Test;
-import org.junit.contrib.java.lang.system.Assertion;
-import org.junit.contrib.java.lang.system.ExpectedSystemExit;
-
-import java.io.*;
-import java.util.Scanner;
-
-import static org.hamcrest.CoreMatchers.containsString;
-import static org.junit.Assert.assertEquals;
-import static org.junit.Assert.assertThat;
-
-public class MainSpec {
-
-    @Rule
-    public final ExpectedSystemExit exit = ExpectedSystemExit.none();
-
-    @Test
-    public void testExit() {
-        exit.expectSystemExitWithStatus(0);
-
-        String input = "3\n";
-        StringWriter output = new StringWriter();
-        Main.main(new Scanner(input), output);
-    }
-
-    @Test
-    public void generateGraph() {
-        exit.expectSystemExitWithStatus(0);
-        exit.checkAssertionAfterwards(() -> assertThat(Main.outputString.toString(), containsString("Colors used: 1")));
-
-        String input = "2\n"
-                + "1\n" // 1 vertex
-                + "0\n\n"
-                + "3\n";
-
-        StringWriter output = new StringWriter();
-        Main.main(new Scanner(input), output);
-    }
-
-    @Test
-    public void enterGraph() {
-        exit.expectSystemExitWithStatus(0);
-        exit.checkAssertionAfterwards(() -> {
-            assertThat(Main.outputString.toString(), containsString("Colors used: 2"));
-            assertThat(Main.outputString.toString(), containsString("Color 0 => 1"));
-            assertThat(Main.outputString.toString(), containsString("Color 1 => 1"));
-            assertThat(Main.outputString.toString(), containsString("Vertex 0 => color 0"));
-            assertThat(Main.outputString.toString(), containsString("Vertex 1 => color 1"));
-        });
-
-        String input = "1\n"
-                + "2\n"
-                + "01\n"
-                + "10\n\n"
-                + "3\n";
-
-        StringWriter output = new StringWriter();
-        Main.main(new Scanner(input), output);
-    }
-
-    @Test
-    public void invalidOption() {
-        exit.expectSystemExitWithStatus(0);
-        exit.checkAssertionAfterwards(() -> {
-            assertThat(Main.outputString.toString(), containsString("Please choose an option."));
-        });
-
-        String input = "4\n"
-                + "3\n";
-
[…]

### 2017-02-13 src/test/java/com/devshawn/coloring/WelshPowellSpec.java

@@ -1,72 +1,0 @@
-package com.devshawn.coloring;
-
-import org.junit.Test;
-
-import static org.junit.Assert.assertArrayEquals;
-import static org.junit.Assert.assertEquals;
-import static org.junit.Assert.assertNotEquals;
-
-public class WelshPowellSpec {
-
-    @Test
-    public void testWelshPowell() {
-
-        // a 5-vertex graph
-
-        /*
-         * (0) ----- (2)
-         *  | \     / |
-         *  |   (4)   |
-         *  | /     \ |
-         * (1) ----- (3)
-         *
-         */
-
-        int[][] graph = {
-                {0, 1, 1, 0, 1},
-                {1, 0, 0, 1, 1},
-                {1, 0, 0, 1, 1},
-                {0, 1, 1, 0, 1},
-                {1, 1, 1, 1, 0}
-        };
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.WELSH_POWELL);
-
-        assertEquals(3, result.getColoringNumber());
-        int[] expectedColoring = {1, 2, 2, 1, 0};
-        assertArrayEquals(expectedColoring, result.getColoring());
-    }
-
-    @Test
-    public void producesDifferentThanGreedy() {
-
-        // a 5-vertex graph
-
-        /*
-         * (0) ----- (2)
-         *  | \     / |
-         *  |   (4)   |
-         *  | /     \ |
-         * (1) ----- (3)
-         *
-         */
-
-        int[][] graph = {
-                {0, 1, 1, 0, 1},
-                {1, 0, 0, 1, 1},
-                {1, 0, 0, 1, 1},
-                {0, 1, 1, 0, 1},
-                {1, 1, 1, 1, 0}
-        };
-
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult greedyResult = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
-        ColoringResult welshPowellResult = coloring.applyHeuristic(ColoringHeuristic.WELSH_POWELL);
-        for(int i = 0; i < greedyResult.getColoring().length; i++) {
-            assertNotEquals(greedyResult.getColoring()[i], welshPowellResult.getColoring()[i]);
-        }
-    }
-}

### 2017-02-13 library/src/main/java/com/devshawn/coloring/library/GraphGenerator.java
Working on adding graph stuff

@@ -1,5 +1,9 @@
 package com.devshawn.coloring.library;
 
+import java.util.Scanner;
+import java.util.regex.Matcher;
+import java.util.regex.Pattern;
+
 public class GraphGenerator {
 
     // simple graph with a given number of vertices
@@ -33,5 +37,30 @@ public class GraphGenerator {
         }
 
         return true;
+    }
+
+    public static int[][] matrixStringToGraph(String str) {
+        Scanner sc = new Scanner(str);
+
+        // Find number of vertices
+        Matcher m = Pattern.compile("\r\n|\r|\n").matcher(str);
+        int vertices = 1;
+        while (m.find())
+        {
+            vertices++;
+        }
+
+        int[][] graph = new int[vertices][];
+
+        for(int i = 0; i < vertices; i++) {
+            String[] line = sc.nextLine().split("");
+            int[] numbers = new int[vertices];
+            for(int j = 0; j < line.length; j++) {
+                numbers[j] = Integer.parseInt(line[j]);
+            }
+            graph[i] = numbers;
+        }
+
+        return graph;
     }
 }

### 2017-02-13 server/src/main/resources/static/components/graphs/graphs-create.controller.js

@@ -1,0 +1,23 @@
+angular.module('coloring')
+    .controller('GraphsCreateController', function (GraphService, $state) {
+        var self = this;
+
+        self.createGraph = function() {
+            console.log("Clicked submit");
+
+            if(self.name && self.matrix) {
+                console.log("We have both");
+
+                var data = {
+                    name: self.name,
+                    matrix: self.matrix
+                };
+
+                GraphService.save(data, function(response) {
+                    console.log(response);
+                    $state.go('graphs');
+                });
+            }
+        }
+
+    });

### 2017-02-13 server/src/main/resources/static/components/graphs/graphs.controller.js

@@ -5,4 +5,10 @@ angular.module('coloring')
         GraphService.query(function(results) {
             self.graphs = results;
         });
+
+        self.deleteGraph = function(id) {
+            GraphService.delete({id: id}, function(success) {
+                console.log("Deleted");
+            });
+        }
     });

### 2017-02-13 server/src/main/resources/static/components/graphs/graphs.routes.js

@@ -9,5 +9,14 @@ angular.module('coloring')
             controllerAs: "graphsController"
         };
 
+        var graphsCreateState = {
+            name: "graphs-create",
+            url: "/graphs/create",
+            templateUrl: "components/graphs/graphs-create.html",
+            controller: "GraphsCreateController",
+            controllerAs: "graphsCreateController"
+        };
+
         $stateProvider.state(graphsState);
+        $stateProvider.state(graphsCreateState);
     });

### 2017-02-14 server/src/main/resources/static/components/graphs/graphs-view.controller.js
Adding visualization of graphs

@@ -1,0 +1,82 @@
+angular.module('coloring')
+    .controller('GraphsViewController', function (GraphService, $stateParams, $scope) {
+        var self = this;
+        console.log($stateParams.id);
+        GraphService.get({id: $stateParams.id}, function(results) {
+            self.graph = results;
+            self.showGraph = (results.vertices < 50);
+            if(self.showGraph) {
+                generateVisualization(results.matrix);
+            }
+            self.graph.matrix = arrayGraphToString(results.matrix);
+
+        });
+
+        var arrayGraphToString = function(array) {
+            var str = "";
+            for(var i = 0; i < array.length; i++) {
+                for(var j = 0; j < array.length; j++) {
+                    str += array[i][j];
+                }
+                str += "\n";
+            }
+            return str;
+        };
+
+        var generateVisualization = function(matrix) {
+            for(var i = 0; i < matrix.length; i++) {
+                console.log(i);
+                self.nodes.add([{id: i, label: i}]);
+            }
+
+            for(var i = 0; i < matrix.length; i++) {
+                for(var j = 0; j < i; j++) {
+                    if(matrix[i][j] == 1) {
+                        self.edges.add([{from: i, to: j}]);
+                    }
+                }
+            }
+        };
+
+        self.nodes = new vis.DataSet();
+        self.edges = new vis.DataSet();
+        self.network_data = {
+            nodes: self.nodes,
+            edges: self.edges
+        };
+        self.network_options = {
+            width: '100%',
+            height: '600px',
+            nodes: {
+                color: '#aaa',
+                shape: 'circle',
+                size: 10,
+                font: {
+                    size: 10,
+                    color: '#000000'
+                },
+                borderWidth: 2
+            },
+            edges: {
+                color: '#000000',
+                width: 2,
+                smooth: {
+                    enabled: false
+                }
+            },
+            physics: {
+                solver: 'repulsion',
+                repulsion: {
+                    nodeDistance: 80
+                },
+                stabilization: {
+                    enabled: true,
+                    iterations: 180, // maximum number of iteration to stabilize
+                    updateInterval: 10,
+                    onlyDynamicEdges: false,
+                    fit: true
+                }
+            }
+        };
+    });
+

### 2017-02-14 server/src/main/resources/static/components/graphs/graphs.controller.js

@@ -1,14 +1,14 @@
 angular.module('coloring')
-    .controller('GraphsController', function (GraphService) {
+    .controller('GraphsController', function (GraphService, $state) {
         var self = this;
 
         GraphService.query(function(results) {
             self.graphs = results;
         });
 
-        self.deleteGraph = function(id) {
+        self.deleteGraph = function(id, index) {
             GraphService.delete({id: id}, function(success) {
-                console.log("Deleted");
+                self.graphs.splice(index, 1);
             });
         }
     });

### 2017-02-14 server/src/main/resources/static/components/graphs/graphs.routes.js

@@ -17,6 +17,15 @@ angular.module('coloring')
             controllerAs: "graphsCreateController"
         };
 
+        var graphsViewState = {
+            name: "graphs-view",
+            url: "/graphs/:id",
+            templateUrl: "components/graphs/graphs-view.html",
+            controller: "GraphsViewController",
+            controllerAs: "graphsViewController"
+        };
+
         $stateProvider.state(graphsState);
         $stateProvider.state(graphsCreateState);
+        $stateProvider.state(graphsViewState);
     });

### 2017-02-14 server/src/main/resources/static/components/visjs/visNetwork.directive.js

@@ -1,0 +1,85 @@
+angular.module('coloring')
+    .directive('visNetwork', function () {
+        return {
+            restrict: 'EA',
+            transclude: false,
+            scope: {
+                data: '=',
+                options: '=',
+                events: '='
+            },
+            link: function (scope, element, attr) {
+                var networkEvents = [
+                    'click',
+                    'doubleClick',
+                    'oncontext',
+                    'hold',
+                    'release',
+                    'selectNode',
+                    'selectEdge',
+                    'deselectNode',
+                    'deselectEdge',
+                    'dragStart',
+                    'dragging',
+                    'dragEnd',
+                    'hoverNode',
+                    'hoverEdge',
+                    'blurNode',
+                    'blurEdge',
+                    'zoom',
+                    'showPopup',
+                    'hidePopup',
+                    'startStabilizing',
+                    'stabilizationProgress',
+                    'stabilizationIterationsDone',
+                    'stabilized',
+                    'resize',
+                    'initRedraw',
+                    'beforeDrawing',
+                    'afterDrawing',
+                    'animationFinished'
+
+                ];
+
+                var network = null;
+
+                scope.$watch('data', function () {
+                    // Sanity check
+                    if (scope.data == null) {
+                        return;
+                    }
+
+                    // If we've actually changed the data set, then recreate the graph
+                    // We can always update the data by adding more data to the existing data set
+                    if (network != null) {
+                        network.destroy();
+                    }
+
+                    network = new vis.Network(element[0], scope.data, scope.options);
+                    network.fit();
+                    network.on('stabilized', function() {
+                        //network.fit();
+                    });
+
+                    angular.forEach(scope.events, function (callback, event) {
+                        if (networkEvents.indexOf(String(event)) >= 0) {
+                            network.on(event, callback);
+                        }
+                    });
+
+                    if (scope.events != null && scope.events.onload != null &&
+                        angular.isFunction(scope.events.onload)) {
+                        scope.events.onload(network);
+                    }
+                });
+
+                scope.$watchCollection('options', function (options) {
+                    if (network == null) {
+                        return;
+                    }
+                    network.setOptions(options);
+                });
+            }
+        };
+    });
+

### 2017-02-14 library/src/main/java/com/devshawn/coloring/library/ColoringModule.java
Adding ability to color graphs from server

@@ -2,7 +2,7 @@ package com.devshawn.coloring.library;
 
 import java.util.*;
 
-public class Coloring {
+public class ColoringModule {
 
     private int[][] graph;
     private int vertexCount;

### 2017-02-14 library/src/main/java/com/devshawn/coloring/library/ColoringResult.java

@@ -44,6 +44,8 @@ public class ColoringResult {
         return this.result;
     }
 
+    public ColoringHeuristic getHeuristic() { return this.heuristicUsed; }
+
     public String getSummary() {
         String string = "The graph was colored with the " + heuristicUsed.getName() + " heuristic.";
 

### 2017-02-14 library/src/main/java/com/devshawn/coloring/library/Main.java

@@ -81,13 +81,13 @@ public class Main {
 
         int[][] graph = GraphGenerator.simple(vertices, percentage);
 
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
+        ColoringModule coloringModule = new ColoringModule();
+        coloringModule.setGraph(graph);
 
-        output.println("Edges generated: " + coloring.getEdgeCount());
+        output.println("Edges generated: " + coloringModule.getEdgeCount());
         output.flush();
 
-        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
+        ColoringResult result = coloringModule.applyHeuristic(ColoringHeuristic.GREEDY);
         result.printSummary(output);
         sc.nextLine();
     }
@@ -115,9 +115,9 @@ public class Main {
         }
 
 
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
+        ColoringModule coloringModule = new ColoringModule();
+        coloringModule.setGraph(graph);
+        ColoringResult result = coloringModule.applyHeuristic(ColoringHeuristic.GREEDY);
         result.printSummary(output);
     }
 }

### 2017-02-14 library/src/test/java/com/devshawn/coloring/library/ColoringSpec.java

@@ -23,11 +23,11 @@ public class ColoringSpec {
                 {1, 1, 0}
         };
 
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        assertTrue(graph == coloring.getGraph());
-        assertEquals(3, coloring.getVertexCount());
-        assertEquals(3, coloring.getEdgeCount());
+        ColoringModule coloringModule = new ColoringModule();
+        coloringModule.setGraph(graph);
+        assertTrue(graph == coloringModule.getGraph());
+        assertEquals(3, coloringModule.getVertexCount());
+        assertEquals(3, coloringModule.getEdgeCount());
 
     }
 
@@ -40,8 +40,8 @@ public class ColoringSpec {
                 {1, 1, 0}
         };
 
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
+        ColoringModule coloringModule = new ColoringModule();
+        coloringModule.setGraph(graph);
     }
 
     @Test
@@ -55,8 +55,8 @@ public class ColoringSpec {
                 {1, 1, 0}
         };
 
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
+        ColoringModule coloringModule = new ColoringModule();
+        coloringModule.setGraph(graph);
     }
 
 }

### 2017-02-14 library/src/test/java/com/devshawn/coloring/library/DSATURSpec.java

@@ -30,9 +30,9 @@ public class DSATURSpec {
                 {1, 1, 1, 1, 0}
         };
 
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.DSATUR);
+        ColoringModule coloringModule = new ColoringModule();
+        coloringModule.setGraph(graph);
+        ColoringResult result = coloringModule.applyHeuristic(ColoringHeuristic.DSATUR);
 
         System.out.println(Arrays.toString(result.getColoring()));
     }

### 2017-02-14 library/src/test/java/com/devshawn/coloring/library/GreedySpec.java

@@ -17,9 +17,9 @@ public class GreedySpec {
                 {1, 1, 0}
         };
 
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
+        ColoringModule coloringModule = new ColoringModule();
+        coloringModule.setGraph(graph);
+        ColoringResult result = coloringModule.applyHeuristic(ColoringHeuristic.GREEDY);
 
         assertEquals(3, result.getColoringNumber());
         int[] expectedColoring = {0, 1, 2};
@@ -48,9 +48,9 @@ public class GreedySpec {
                 {1, 1, 1, 1, 0}
         };
 
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
+        ColoringModule coloringModule = new ColoringModule();
+        coloringModule.setGraph(graph);
+        ColoringResult result = coloringModule.applyHeuristic(ColoringHeuristic.GREEDY);
 
         assertEquals(3, result.getColoringNumber());
         int[] expectedColoring = {0, 1, 1, 0, 2};

### 2017-02-14 library/src/test/java/com/devshawn/coloring/library/MaximalIndependentSetSpec.java

@@ -28,9 +28,9 @@ public class MaximalIndependentSetSpec {
                 {1, 1, 1, 1, 0}
         };
 
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.MAXIMAL_INDEPENDENT_SET);
+        ColoringModule coloringModule = new ColoringModule();
+        coloringModule.setGraph(graph);
+        ColoringResult result = coloringModule.applyHeuristic(ColoringHeuristic.MAXIMAL_INDEPENDENT_SET);
 
         assertEquals(3, result.getColoringNumber());
         int[] expectedColoring = {0, 1, 1, 0, 2};

### 2017-02-14 library/src/test/java/com/devshawn/coloring/library/WelshPowellSpec.java

@@ -30,9 +30,9 @@ public class WelshPowellSpec {
                 {1, 1, 1, 1, 0}
         };
 
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult result = coloring.applyHeuristic(ColoringHeuristic.WELSH_POWELL);
+        ColoringModule coloringModule = new ColoringModule();
+        coloringModule.setGraph(graph);
+        ColoringResult result = coloringModule.applyHeuristic(ColoringHeuristic.WELSH_POWELL);
 
         assertEquals(3, result.getColoringNumber());
         int[] expectedColoring = {1, 2, 2, 1, 0};
@@ -61,10 +61,10 @@ public class WelshPowellSpec {
                 {1, 1, 1, 1, 0}
         };
 
-        Coloring coloring = new Coloring();
-        coloring.setGraph(graph);
-        ColoringResult greedyResult = coloring.applyHeuristic(ColoringHeuristic.GREEDY);
-        ColoringResult welshPowellResult = coloring.applyHeuristic(ColoringHeuristic.WELSH_POWELL);
+        ColoringModule coloringModule = new ColoringModule();
+        coloringModule.setGraph(graph);
+        ColoringResult greedyResult = coloringModule.applyHeuristic(ColoringHeuristic.GREEDY);
+        ColoringResult welshPowellResult = coloringModule.applyHeuristic(ColoringHeuristic.WELSH_POWELL);
         for(int i = 0; i < greedyResult.getColoring().length; i++) {
             assertNotEquals(greedyResult.getColoring()[i], welshPowellResult.getColoring()[i]);
         }

### 2017-02-14 server/src/main/resources/static/components/colorings/colorings-create.controller.js

@@ -1,0 +1,36 @@
+angular.module('coloring')
+    .controller('ColoringsCreateController', function (ColoringsService, GraphService, $state, $stateParams) {
+        var self = this;
+
+        var heuristicsCallback = function(heuristics) {
+            self.heuristics = heuristics;
+        };
+
+        var graphsCallback = function(graph) {
+            self.graph = graph;
+            ColoringsService.heuristics({}, heuristicsCallback);
+        };
+
+        if($state.$current.name == "colorings-create-coloring") {
+            GraphService.get({id: $stateParams.id}, graphsCallback);
+        } else {
+
+        }
+
+        self.generateColoring = function() {
+            console.log("Clicked submit");
+
+            if(self.heuristic) {
+
+                var data = {
+                    graphId: self.graph.id,
+                    heuristic: self.heuristic
+                };
+
+                ColoringsService.save(data, function(response) {
+                    console.log(response);
+                    $state.go('colorings');
+                });
+            }
+        }
+    });

### 2017-02-14 server/src/main/resources/static/components/colorings/colorings-view.controller.js

@@ -1,0 +1,34 @@
+angular.module('coloring')
+    .controller('ColoringsViewController', function (ColoringsService, $stateParams, $scope) {
+        var self = this;
+
+        ColoringsService.get({id: $stateParams.id}, function(results) {
+            self.showGraph = (results.graph.vertices < 50);
+            if(self.showGraph) {
+                generateVisualization(results.graph.matrix, results.result.coloring);
+            }
+            self.coloring = results;
+        });
+
+        var generateVisualization = function(matrix, coloring) {
+            for(var i = 0; i < matrix.length; i++) {
+                self.nodes.add([{id: i, label: i, group: coloring[i]}]);
+            }
+
+            for(var i = 0; i < matrix.length; i++) {
+                for(var j = 0; j < i; j++) {
+                    if(matrix[i][j] == 1) {
+                        self.edges.add([{from: i, to: j}]);
+                    }
+                }
+            }
+        };
+
+        self.nodes = new vis.DataSet();
+        self.edges = new vis.DataSet();
+        self.network_data = {
+            nodes: self.nodes,
+            edges: self.edges
+        };
+    });
+

### 2017-02-14 server/src/main/resources/static/components/colorings/colorings.controller.js

@@ -1,0 +1,14 @@
+angular.module('coloring')
+    .controller('ColoringsController', function (ColoringsService, $state) {
+        var self = this;
+
+        ColoringsService.query(function(results) {
+            self.colorings = results;
+        });
+
+        self.deleteColoring = function(id, index) {
+            ColoringsService.delete({id: id}, function(success) {
+                self.colorings.splice(index, 1);
+            });
+        }
+    });

### 2017-02-14 server/src/main/resources/static/components/colorings/colorings.routes.js

@@ -1,0 +1,40 @@
+angular.module('coloring')
+    .config(function($stateProvider) {
+
+        var coloringsState = {
+            name: "colorings",
+            url: "/colorings",
+            templateUrl: "components/colorings/colorings.html",
+            controller: "ColoringsController",
+            controllerAs: "coloringsController"
+        };
+
+        var coloringsCreateState = {
+            name: "colorings-create",
+            url: "/colorings/create",
+            templateUrl: "components/colorings/colorings-create.html",
+            controller: "ColoringsCreateController",
+            controllerAs: "coloringsCreateController"
+        };
+
+        var coloringsCreateColoringState = {
+            name: "colorings-create-coloring",
+            url: "/colorings/create/:id",
+            templateUrl: "components/colorings/colorings-create-coloring.html",
+            controller: "ColoringsCreateController",
+            controllerAs: "coloringsCreateController"
+        };
+
+        var coloringsViewState = {
+            name: "colorings-view",
+            url: "/colorings/:id",
+            templateUrl: "components/colorings/colorings-view.html",
+            controller: "ColoringsViewController",
+            controllerAs: "coloringsViewController"
+        };
+
+        $stateProvider.state(coloringsState);
+        $stateProvider.state(coloringsCreateState);
+        $stateProvider.state(coloringsCreateColoringState);
+        $stateProvider.state(coloringsViewState);
+    });

### 2017-02-14 server/src/main/resources/static/components/colorings/colorings.service.js

@@ -1,0 +1,11 @@
+angular.module('coloring')
+    .factory('ColoringsService', function($resource) {
+        return new $resource('/api/colorings/:id', null, {
+            'update': { method: 'PUT'},
+            'heuristics': {
+                method: 'GET',
+                isArray: true,
+                url: '/api/colorings/heuristics'
+            }
+        });
+    });

### 2017-02-14 server/src/main/resources/static/components/graphs/graphs-view.controller.js

@@ -1,7 +1,7 @@
 angular.module('coloring')
     .controller('GraphsViewController', function (GraphService, $stateParams, $scope) {
         var self = this;
-        console.log($stateParams.id);
+
         GraphService.get({id: $stateParams.id}, function(results) {
             self.graph = results;
             self.showGraph = (results.vertices < 50);
@@ -25,7 +25,6 @@ angular.module('coloring')
 
         var generateVisualization = function(matrix) {
             for(var i = 0; i < matrix.length; i++) {
-                console.log(i);
                 self.nodes.add([{id: i, label: i}]);
             }
 
@@ -44,39 +43,6 @@ angular.module('coloring')
             nodes: self.nodes,
             edges: self.edges
         };
-        self.network_options = {
-            width: '100%',
-            height: '600px',
-            nodes: {
-                color: '#aaa',
-                shape: 'circle',
-                size: 10,
-                font: {
-                    size: 10,
-                    color: '#000000'
-                },
-                borderWidth: 2
-            },
-            edges: {
-                color: '#000000',
-                width: 2,
-                smooth: {
-                    enabled: false
-                }
-            },
-            physics: {
-                solver: 'repulsion',
-                repulsion: {
-                    nodeDistance: 80
-                },
-                stabilization: {
-                    enabled: true,
-                    iterations: 180, // maximum number of iteration to stabilize
-                    updateInterval: 10,
-                    onlyDynamicEdges: false,
-                    fit: true
-                }
-            }
-        };
+
     });
 

### 2017-02-14 server/src/main/resources/static/components/visjs/visNetwork.directive.js

@@ -43,6 +43,41 @@ angular.module('coloring')
 
                 var network = null;
 
+                var options = {
+                    width: '100%',
+                    height: '600px',
+                    nodes: {
+                        color: '#aaa',
+                        shape: 'circle',
+                        size: 10,
+                        font: {
+                            size: 10,
+                            color: '#000000'
+                        },
+                        borderWidth: 2
+                    },
+                    edges: {
+                        color: '#000000',
+                        width: 2,
+                        smooth: {
+                            enabled: false
+                        }
+                    },
+                    physics: {
+                        solver: 'repulsion',
+                        repulsion: {
+                            nodeDistance: 80
+                        },
+                        stabilization: {
+                            enabled: true,
+                            iterations: 180, // maximum number of iteration to stabilize
+                            updateInterval: 10,
+                            onlyDynamicEdges: false,
+                            fit: true
+                        }
+                    }
+                };
+
                 scope.$watch('data', function () {
                     // Sanity check
                     if (scope.data == null) {
@@ -55,7 +90,7 @@ angular.module('coloring')
                         network.destroy();
                     }
 
-                    network = new vis.Network(element[0], scope.data, scope.options);
+                    network = new vis.Network(element[0], scope.data, options);
                     network.fit();
                     network.on('stabilized', function() {
                         //network.fit();
@@ -73,12 +108,12 @@ angular.module('coloring')
                     }
                 });
 
-                scope.$watchCollection('options', function (options) {
-                    if (network == null) {
-                        return;
-                    }
-                    network.setOptions(options);
-                });
+                // scope.$watchCollection('options', function (options) {
+                //     if (network == null) {
+                //         return;
+                //     }
+                //     network.setOptions(options);
+                // });
             }
         };
     });

### 2017-02-14 library/src/main/java/com/devshawn/coloring/library/ColoringModule.java
Fixing some algorithms and working on coloring more

@@ -95,7 +95,8 @@ public class ColoringModule {
 
         // Sort our vertices by decreasing degree
         List<Integer> indexList = Arrays.asList(orderedVertices);
-        Collections.sort(indexList, (left, right) -> (degrees[indexList.indexOf(right)] - degrees[indexList.indexOf(left)]));
+        Collections.sort(indexList, (left, right) -> (degrees[indexList.indexOf(left)] - degrees[indexList.indexOf(right)]));
+        Collections.reverse(indexList);
         orderedVertices = indexList.toArray(new Integer[graph.length]);
         System.out.println(Arrays.toString(orderedVertices));
 
@@ -206,7 +207,8 @@ public class ColoringModule {
 
         // Sort our vertices by decreasing degree
         List<Integer> indexList = Arrays.asList(orderedVertices);
-        Collections.sort(indexList, (left, right) -> (degrees[indexList.indexOf(right)] - degrees[indexList.indexOf(left)]));
+        Collections.sort(indexList, (left, right) -> (degrees[indexList.indexOf(left)] - degrees[indexList.indexOf(right)]));
+        Collections.reverse(indexList);
         orderedVertices = indexList.toArray(new Integer[graph.length]);
         System.out.println(Arrays.toString(orderedVertices));
 

### 2017-02-14 server/src/main/resources/static/components/colorings/colorings-view.controller.js

@@ -3,7 +3,7 @@ angular.module('coloring')
         var self = this;
 
         ColoringsService.get({id: $stateParams.id}, function(results) {
-            self.showGraph = (results.graph.vertices < 50);
+            self.showGraph = (results.graph.vertices < 51);
             if(self.showGraph) {
                 generateVisualization(results.graph.matrix, results.result.coloring);
             }

### 2017-02-14 server/src/main/resources/static/components/graphs/graphs-create.controller.js

@@ -20,4 +20,22 @@ angular.module('coloring')
             }
         }
 
+        self.generateGraph = function() {
+            console.log(self.edges.length);
+            if(self.name2 && self.vertices && self.edges != null) {
+                console.log("A");
+
+                var data = {
+                    name: self.name2,
+                    vertices: self.vertices,
+                    edges: self.edges
+                };
+
+                GraphService.save(data, function(response) {
+                    console.log(response);
+                    $state.go('graphs');
+                });
+            }
+        }
+
     });

### 2017-02-14 server/src/main/resources/static/components/graphs/graphs-view.controller.js

@@ -4,7 +4,7 @@ angular.module('coloring')
 
         GraphService.get({id: $stateParams.id}, function(results) {
             self.graph = results;
-            self.showGraph = (results.vertices < 50);
+            self.showGraph = (results.vertices < 51);
             if(self.showGraph) {
                 generateVisualization(results.matrix);
             }

### 2017-02-14 work/work.tex
Adding refs for heuristic lettering and updating crown graph definition

@@
 
+\newcounter{heuristic} \setcounter{heuristic}{0}
+\renewcommand{\theheuristic}{\Alph{heuristic}}
+
+\newenvironment{heuristic}[1]%
+{
+\refstepcounter{heuristic}
+\subsection*{Heuristic \theheuristic}
+}%
+{}
+
 \makeatletter
@@
 \newtheorem{prop}{Proposition}
-\newtheorem*{theorem}{Theorem}
+\newtheorem{theorem}{Theorem}
 
 \theoremstyle{definition}
-\newtheorem*{definition}{Definition}
+\newtheorem{definition}{Definition}
 
@@
 
-\subsection*{Heuristic A}
+\heuristic{}\label{heuristic-greedy}
 
-Heuristic A is the greedy algorithm.
+Heuristic \ref{heuristic-greedy} is the greedy algorithm.
 
@@
 
-\subsection*{Heuristic B}
+\heuristic{}\label{heuristic-welsh}
 
-Heuristic B is the greedy algorithm with degree sequencing.
+Heuristic \ref{heuristic-welsh} is the greedy algorithm with degree sequencing.
 It orders the vertices according to the decreasing value of their degree.
@@
 
-\subsection*{Heuristic C}
+\heuristic{}\label{heuristic-mis}
 
-Heuristic C colors a graph by finding maximal independent sets of \(G\).
+Heuristic \ref{heuristic-mis} colors a graph by finding maximal independent sets of \(G\).
 
@@
 \begin{prop}
-Heuristic A and Heuristic B do not always produce an optimal solution, i.e. they do not always produce a minimum coloring of a graph.
+Heuristic \ref{heuristic-greedy} and Heuristic \ref{heuristic-welsh} do not always produce an optimal solution, i.e. they do not always produce a minimum coloring of a graph.
 \end{prop}
 
-I will show this for \emph{heuristic B}.
-It is trivial to show the same for \emph{heuristic A} as \emph{heuristic B} includes the same steps as \emph{heuristic A}.
+I will show this for \emph{heuristic \ref{heuristic-welsh}}.
+It is trivial to show the same for \emph{heuristic \ref{heuristic-greedy}} as \emph{heuristic \ref{heuristic-welsh}} includes the same steps as \emph{heuristic \ref{heuristic-greedy}}.
 I will construct a simple graph, \(G = (V, E)\), such that:
@@
 
-Let \(\chi\) refer to the coloring of \(G\) generated by \emph{heuristic B}. Let \(\chi^{*}\) be the optimal coloring of \(G\).
+Let \(\chi\) refer to the coloring of \(G\) generated by \emph{heuristic \ref{heuristic-welsh}}. Let \(\chi^{*}\) be the optimal coloring of \(G\).
 
@@
 
-By applying \emph{heuristic B}, we get the following coloring.
+By applying \emph{heuristic \ref{heuristic-welsh}}, we get the following coloring.
 The numbers indicate the ordering of vertices before applying the heuristic.
@@
 \includegraphics[scale=0.6]{images/graph-2.png}
-\caption{Coloring from applying heuristic B}
+\caption{Coloring from applying heuristic \ref{heuristic-welsh}}
 \end{figure}
@@
 \begin{prop}
-Heuristic C also does not always produce an optimal solution, i.e. it does not always produce a minimum coloring of a graph.
+Heuristic \ref{heuristic-mis} also does not always produce an optimal solution, i.e. it does not always produce a minimum coloring of a graph.
 \end{prop}
@@
 
-By applying heuristic C and finding maximal independent sets, we see that \(K\) is \emph{5-colorable}.
+By applying heuristic \ref{heuristic-mis} and finding maximal independent sets, we see that \(K\) is \emph{5-colorable}.
 
@@
 \includegraphics[scale=0.38]{images/mis-2.png}
-\caption{Coloring of \(K\) after applying heuristic C}
+\caption{Coloring of \(K\) after applying heuristic \ref{heuristic-mis}}
 \end{figure}
@@
 This graph is \emph{4-colorable} however, meaning \(\chi^{*}(K) = 4\).
-Thus, we can see that heuristic C does not always give us an optimal solution.
+Thus, we can see that heuristic \ref{heuristic-mis} does not always give us an optimal solution.
 
@@
 \begin{prop}
-Heuristic A and heuristic B have an upper bound of \(\Delta + 1\).
+Heuristic \ref{heuristic-greedy} and heuristic \ref{heuristic-welsh} have an upper bound of \(\Delta + 1\).
 \end{prop}
@@
 \begin{definition}[Crown Graph]
-A crown graph \(CR_n = (V, E)\) is an undirected graph with two sets of vertices where \(V = V_1 \cup V_2\) with an edge from \(v_i \in V_1\) to \(v_{j} \in V_2\) whenever \(i \neq j\).
+A crown graph \(CR_n = (V, E)\) is an undirected graph such that \(U = \{u_1, u_2, \dots, u_i\}\) and \(W = \{w_1, w_2, \dots, w_j\}\) where \(V = U \cup W\), \(U \cap W = \emptyset\), and \(|U| = |W|\).
+There is an edge from \(u_i\) to \(w_j\) whenever \(i \neq j\).
 A crown graph can also be viewed as a complete bipartite graph where the edges of a perfect matching have been removed.
@@
 \includegraphics[scale=0.38]{images/graph-5.png}
-\caption{Worst-case coloring of \(H\) using either heuristic A or B}
+\caption{Worst-case coloring of \(H\) using either heuristic \ref{heuristic-greedy} or \ref{heuristic-welsh}}
 \end{figure}
@@
 The proof of \emph{Brook's Theorem} can be found in \cite{lovasz}.
-Overall, heuristic A and heuristic B can produce some very undesirable results.
+Overall, heuristic \ref{heuristic-greedy} and heuristic \ref{heuristic-welsh} can produce some very undesirable results.
 In graph \(H\), at the worst case, these heuristics produce \(\chi(H) = 5\) when \(\chi^{*}(H) = 2\) as it is bipartite.
@@
 \begin{prop}
-Heuristic A and Heuristic B have a running time of $O(|V| + |E|)$.
+Heuristic \ref{heuristic-greedy} and Heuristic \ref{heuristic-welsh} have a running time of $O(|V| + |E|)$.
 \end{prop}
 
-It is possible to implement heuristic A and heuristic B efficiently and achieve a linear running time \cite{kubale}.
+It is possible to implement heuristic \ref{heuristic-greedy} and heuristic \ref{heuristic-welsh} efficiently and achieve a linear running time \cite{kubale}.
 By using an adjacency list representation of a graph \(G = (V, E)\), the greedy algorithm will look once at each vertex and each edge twice.
@@
 \begin{prop}
-Heuristic A and Heuristic B produce a feasible, proper coloring of input graph \(G = (V, E)\).
+Heuristic \ref{heuristic-greedy} and Heuristic \ref{heuristic-welsh} produce a feasible, proper coloring of input graph \(G = (V, E)\).
 \end{prop}

### 2017-02-14 library/src/main/java/com/devshawn/coloring/library/GraphGenerator.java
Adding generate adjacency matrix functions

@@ -1,5 +1,6 @@
 package com.devshawn.coloring.library;
 
+import java.util.List;
 import java.util.Scanner;
 import java.util.regex.Matcher;
 import java.util.regex.Pattern;
@@ -62,5 +63,36 @@ public class GraphGenerator {
         }
 
         return graph;
+    }
+
+    public static int[][] generateAdjacencyMatrix(int vertices, List<Tuple> edges) {
+        int[][] graph = new int[vertices][vertices];
+
+        for(Tuple edge : edges) {
+            graph[edge.to][edge.from] = 1;
+            graph[edge.from][edge.to] = 1;
+        }
+
+        return graph;
+    }
+
+    public static String adjacencyMatrixToString(int[][] graph) {
+        String str = "";
+        for(int i = 0; i < graph.length; i++) {
+            for(int j = 0; j < graph.length; j++) {
+                str += graph[i][j];
+            }
+            str += "\n";
+        }
+        return str;
+    }
+}
+
+class Tuple {
+    public final int from;
+    public final int to;
+    public Tuple(int from, int to) {
+        this.from = from;
+        this.to = to;
     }
 }

### 2017-02-14 library/src/test/java/com/devshawn/coloring/library/GraphGeneratorSpec.java

@@ -4,6 +4,11 @@ import org.junit.Rule;
 import org.junit.Test;
 import org.junit.rules.ExpectedException;
 
+import java.util.ArrayList;
+import java.util.Arrays;
+import java.util.List;
+
+import static junit.framework.TestCase.assertTrue;
 import static org.junit.Assert.assertEquals;
 
 public class GraphGeneratorSpec {
@@ -43,5 +48,51 @@ public class GraphGeneratorSpec {
         };
 
         assertEquals(true, GraphGenerator.isSymmetric(graph));
+    }
+
+    @Test
+    public void generateAdjacencyMatrix() {
+        List<Tuple> edges = new ArrayList<>();
+        edges.add(new Tuple(0, 1));
+        edges.add(new Tuple(1, 2));
+        edges.add(new Tuple(0, 2));
+        int[][] graph = GraphGenerator.generateAdjacencyMatrix(3, edges);
+        for(int i = 0; i < graph.length; i++) {
+            for(int j = 0; j < graph.length; j++) {
+                assertEquals((i == j) ? 0 : 1, graph[i][j]);
+            }
+        }
+
+        String str = GraphGenerator.adjacencyMatrixToString(graph);
+        assertTrue(str.equals("011\n101\n110\n"));
+    }
+
+    @Test
+    public void generateCrownGraph() {
+        List<Tuple> edges = new ArrayList<>();
+        edges.add(new Tuple(0, 1));
+        edges.add(new Tuple(0, 2));
+        edges.add(new Tuple(0, 5));
+        edges.add(new Tuple(0, 6));
+        edges.add(new Tuple(1, 3));
+        edges.add(new Tuple(1, 4));
+        edges.add(new Tuple(1, 7));
+        edges.add(new Tuple(2, 3));
+        edges.add(new Tuple(2, 4));
+        edges.add(new Tuple(2, 8));
+        edges.add(new Tuple(3, 5));
+        edges.add(new Tuple(3, 9));
+        edges.add(new Tuple(4, 6));
+        edges.add(new Tuple(4, 9));
+        edges.add(new Tuple(5, 7));
+        edges.add(new Tuple(5, 8));
+        edges.add(new Tuple(6, 7));
+        edges.add(new Tuple(6, 8));
+        edges.add(new Tuple(7, 9));
+        edges.add(new Tuple(8, 9));
+        int[][] graph = GraphGenerator.generateAdjacencyMatrix(10, edges);
+
[…]

### 2017-02-15 work/work.tex
Changing heuristic C to D and adding new heuristic C

@@
 
+\heuristic{}\label{heuristic-dsatur}
+
+Heuristic \ref{heuristic-dsatur} is very similar to heuristic \ref{heuristic-welsh}, but colors based on the \emph{saturation} degree of vertices, which is defined below by \cite{spinrad}.
+The original DSATUR algorithm was proposed by \cite{brelaz}.
+
+\begin{definition}[Saturation degree]
+The saturation degree of a vertex is the number of different colors used for vertices adjacency to it.
+\end{definition}
+
+\begin{algorithm}
+\caption{DSATUR algorithm}
+\begin{algorithmic}[1]
+\State Find the vertex $\in V$ with maximum degree and color it with color 1
+\While{There are uncolored vertices}
+\State Choose an uncolored vertex with maximum saturation degree
+\State If there is a tie, choose the vertex with maximum degree
+\State Color the vertex with the least possible number
+\EndWhile
+\end{algorithmic}
+\end{algorithm}
+
+\newpage
+
 \heuristic{}\label{heuristic-mis}
@@
 \end{algorithm}
-
-\newpage
 
@@
 
-\newpage
-
 \begin{prop}
@@
 American Elsevier Publishing, New York, NY.
+
+\bibitem{brelaz}
+Brélaz, D. [1979]. \emph{New methods to color the vertices of a graph}. Communications of the ACM, 22(4), pp.251-256.
 
@@
 
+\bibitem{spinrad}
+Spinrad, J. P. and Vijayan, G. [1985], \emph{Worst case analysis of a graph coloring algorithm}. Discrete Applied Mathematics, 12(1), pp.89-92.
+
 \bibitem{welsh}

### 2017-02-17 library/src/main/java/com/devshawn/coloring/library/ColoringModule.java
Adding some more fields for the database

@@ -34,6 +34,7 @@ public class ColoringModule {
 
     // Greedy heuristic
     private ColoringResult applyGreedyHeuristic() {
+        long start = System.nanoTime();
 
         // Make result array of vertices, assign no color to them
         int[] result = new int[graph.length];
@@ -69,12 +70,15 @@ public class ColoringModule {
             for(int j = 0; j < graph.length; j++) available[j] = true;
         }
 
+        long end = System.nanoTime();
+
         // Generate a ColorResult
-        return new ColoringResult(result, ColoringHeuristic.GREEDY);
+        return new ColoringResult(result, ColoringHeuristic.GREEDY, (end - start));
     }
 
     // Greedy plus degree sequencing heuristic (Welsh-Powell algorithm)
     private ColoringResult applyWelshPowell() {
+        long start = System.nanoTime();
 
         // Make result array of vertices, assign no color to them
         int[] result = new int[graph.length];
@@ -130,11 +134,13 @@ public class ColoringModule {
             for(int j = 0; j < graph.length; j++) available[j] = true;
         }
 
+        long end = System.nanoTime();
 
-        return new ColoringResult(result, ColoringHeuristic.WELSH_POWELL);
+        return new ColoringResult(result, ColoringHeuristic.WELSH_POWELL, (end - start));
     }
 
     private ColoringResult applyMaximalIndependentSet() {
+        long start = System.nanoTime();
 
         // Make result array of vertices, assign no color to them
         int[] result = new int[graph.length];
@@ -183,10 +189,13 @@ public class ColoringModule {
             color++;
         }
 
-        return new ColoringResult(result, ColoringHeuristic.MAXIMAL_INDEPENDENT_SET);
+        long end = System.nanoTime();
+
+        return new ColoringResult(result, ColoringHeuristic.MAXIMAL_INDEPENDENT_SET, (end - start));
     }
 
     private ColoringResult applyDSATUR() {
+        long start = System.nanoTime();
 
         // Make result array of vertices, assign no color to them
         int[] result = new int[graph.length];
@@ -234,7 +243,9 @@ public class ColoringModule {
             selectedVertex = getMaximumVertexFromSaturated(saturated, vertices);
         }
 
-        return new ColoringResult(result, ColoringHeuristic.DSATUR);
+        long end = System.nanoTime();
+
+        return new ColoringResult(result, ColoringHeuristic.DSATUR, (end - start));
     }
 
     private int getMaximumVertexFromSaturated(Set<Integer>[] saturated, Set<Integer> vertices) {

### 2017-02-17 library/src/main/java/com/devshawn/coloring/library/ColoringResult.java

@@ -7,12 +7,13 @@ public class ColoringResult {
     private int highestColor;
     private ColoringHeuristic heuristicUsed;
     private int[] colors;
+    private long time;
 
     public ColoringResult() {
 
     }
 
-    public ColoringResult(int[] result, ColoringHeuristic heuristicUsed) {
+    public ColoringResult(int[] result, ColoringHeuristic heuristicUsed, long time) {
 
         // Find how many colors we used
         int highestColor = -1;
@@ -34,6 +35,7 @@ public class ColoringResult {
 
         this.result = result;
         this.heuristicUsed = heuristicUsed;
+        this.time = time;
     }
 
     public int getColoringNumber() {
@@ -43,6 +45,8 @@ public class ColoringResult {
     public int[] getColoring() {
         return this.result;
     }
+
+    public long getTime() { return this.time; }
 
     public ColoringHeuristic getHeuristic() { return this.heuristicUsed; }
 

### 2017-02-17 library/src/test/java/com/devshawn/coloring/library/ColoringResultSpec.java

@@ -15,7 +15,7 @@ public class ColoringResultSpec {
     @Test
     public void getSummary() {
         int[] results = {0};
-        ColoringResult result = new ColoringResult(results, ColoringHeuristic.GREEDY);
+        ColoringResult result = new ColoringResult(results, ColoringHeuristic.GREEDY, 0);
 
         String expectedResult = "The graph was colored with the greedy heuristic.";
         expectedResult += "\r\n\tColor usage:";
@@ -32,14 +32,11 @@ public class ColoringResultSpec {
     @Test
     public void printSummary() {
         int[] results = {0};
-        ColoringResult result = new ColoringResult(results, ColoringHeuristic.GREEDY);
+        ColoringResult result = new ColoringResult(results, ColoringHeuristic.GREEDY, 0);
         StringWriter writer = new StringWriter();
         result.printSummary(new PrintWriter(writer));
         assertThat(writer.toString(), containsString("Colors used: 1"));
     }
-
-
-
 
 }
 

### 2017-02-17 server/src/main/resources/static/components/colorings/colorings-create.controller.js

@@ -29,7 +29,7 @@ angular.module('coloring')
 
                 ColoringsService.save(data, function(response) {
                     console.log(response);
-                    $state.go('colorings');
+                    $state.go('colorings-view', { 'id' : response.id });
                 });
             }
         }

### 2017-02-17 server/src/main/resources/static/components/graphs/graphs-create.controller.js

@@ -15,10 +15,10 @@ angular.module('coloring')
 
                 GraphService.save(data, function(response) {
                     console.log(response);
-                    $state.go('graphs');
+                    $state.go('graphs-view', { 'id': response.id });
                 });
             }
-        }
+        };
 
         self.generateGraph = function() {
             console.log(self.edges.length);
@@ -33,9 +33,11 @@ angular.module('coloring')
 
                 GraphService.save(data, function(response) {
                     console.log(response);
-                    $state.go('graphs');
+                    $state.go('graphs-view', { 'id': response.id });
+                }, function(error) {
+                    console.log(error);
                 });
             }
-        }
+        };
 
     });

### 2017-02-17 server/src/main/resources/static/components/graphs/graphs-view.controller.js

@@ -8,8 +8,12 @@ angular.module('coloring')
             if(self.showGraph) {
                 generateVisualization(results.matrix);
             }
-            self.graph.matrix = arrayGraphToString(results.matrix);
 
+            if(results.matrix != null) {
+                self.graph.matrix = arrayGraphToString(results.matrix);
+            } else {
+                self.graph.matrix = "The graph has too many vertices to show the matrix.";
+            }
         });
 
         var arrayGraphToString = function(array) {

### 2017-02-17 server/src/main/resources/static/components/alert/alert.directive.js
Starting server error handling

@@ -1,0 +1,73 @@
+angular.module('coloring')
+    .service('AlertService', [ '$timeout', function($timeout) {
+
+        this.delay = 5000;
+
+        this.alerts = [];
+        this.warnings = [];
+        this.notices = [];
+
+        this.setCloseDelay = function(delay) {
+            this.delay = delay;
+        };
+
+        this.removeMessage = function(msgType, message, delay) {
+            var tracked = this[msgType + 's'];
+            if (angular.isDefined(delay)) {
+                $timeout(
+                    function() {
+                        var index = tracked.indexOf(message);
+                        if (index > -1) {
+                            tracked.splice(index,1);
+                        }
+                    },
+                    this.delay
+                );
+            } else {
+                var index = tracked.indexOf(message);
+                if (index > -1) {
+                    tracked.splice(index,1);
+                }
+            }
+        };
+
+        this.Message = function(msgType, message) {
+            var key = msgType + 's';
+            this[key].push(message);
+            this.removeMessage(msgType, message, this.delay);
+        };
+
+        this.Alert = function(message) { this.Message('alert', message); };
+        this.Warning = function(message) { this.Message('warning', message); };
+        this.Notice = function(message) { this.Message('notice', message); };
+
+    } ] )
+    .directive('alertBox', [ 'AlertService', function(AlertService) {
+        return {
+            restrict: 'E',
+
+            compile: function(element, attrs) {
+                attrs.boxClass = attrs.boxClass || "alert-box";
+                attrs.alertClass = attrs.alertClass || "alert";
+                attrs.noticeClass = attrs.noticeClass || "success";
+                attrs.warningClass = attrs.warningClass || "warning";
+            },
+
+            scope: {
+                alertClass : "@",
+                warningClass : "@",
+                noticeClass : "@",
+                boxClass : "@"
+            },
+
+            controller: function($scope) {
+                $scope.alerts = AlertService.alerts;
+                $scope.warnings = AlertService.warnings;
+                $scope.notices = AlertService.notices;
+            },
+            template:  '<div ng-repeat="alert in alerts track by $index" ng-class="[boxClass, alertClass]">{{alert}}</div> \
+						<div ng-repeat="warning in warnings track by $index" ng-class="[boxClass, warningClass]">{{warning}}</div> \
+						<div ng-repeat="notice in notices track by $index" ng-class="[boxClass, noticeClass]">{{notice}}</div>'
+        };
+
+    } ] );

### 2017-02-17 server/src/main/resources/static/components/graphs/graphs-create.controller.js

@@ -1,5 +1,5 @@
 angular.module('coloring')
-    .controller('GraphsCreateController', function (GraphService, $state) {
+    .controller('GraphsCreateController', function (GraphService, $state, AlertService) {
         var self = this;
 
         self.createGraph = function() {
@@ -14,8 +14,11 @@ angular.module('coloring')
                 };
 
                 GraphService.save(data, function(response) {
-                    console.log(response);
                     $state.go('graphs-view', { 'id': response.id });
+                }, function(error) {
+                    if(error.status === 500) {
+                        AlertService.Warning(error.data.message);
+                    }
                 });
             }
         };
@@ -23,7 +26,6 @@ angular.module('coloring')
         self.generateGraph = function() {
             console.log(self.edges.length);
             if(self.name2 && self.vertices && self.edges != null) {
-                console.log("A");
 
                 var data = {
                     name: self.name2,

### 2017-02-22 work/abstract.tex
Adding abstract

\documentclass[12pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}

\begin{document}

\noindent\textbf{Presenter}: Shawn Seymour

\vspace{0.3cm}

\noindent\textbf{Project Adviser}: Peh Ng (Mathematics)

\vspace{0.3cm}

\noindent\textbf{Title}: Vertex Coloring and Applications

\vspace{0.3cm}

Consider the map of the 48 contiguous states in the USA, and suppose we want to color each state so that no two states that share a boundary have the same color. In general, we could represent every state with a \emph{vertex} and draw an \emph{edge} between two states that share a border. This problem can be modeled by a mathematical structure called a graph. A graph, denoted \(G = (V,E)\), is a set of vertices \(V\) and a set of edges \(E\). The \emph{Vertex Coloring} problem on \(G\) aims to find the minimum number of colors (the chromatic number) needed to color the vertices such that no two adjacent nodes have the same color. Vertex coloring can solve real-world problems such as finding the minimum number of time slots to schedule a final exam period so that no two courses (taken by the same student) are scheduled at the same time slot. \newline

In general, finding the chromatic number of a graph is an NP-hard problem, meaning there is no known efficient time algorithm to solve it and there will likely not be one. Hence, there is interest in finding approximation algorithms (heuristics) to find the chromatic number. In this research, we present three heuristics used to find good approximate vertex colorings in an efficient time period even though they may not give us the optimal minimum coloring of the graph. This is important as it allows us to approximately solve complex problems in a reasonable amount of time. We will present computational results comparing the efficiency (time and quality) of these heuristics.

\end{document}

### 2017-04-12 poster/poster.tex
Initial poster

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Vertex Coloring Poster
% Created by Shawn Seymour
% April 2017
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[final]{beamer}
\usepackage[scale=1.07]{beamerposter}
\usepackage{graphicx}			% allows us to import images

%-----------------------------------------------------------
% Custom commands that I use frequently
%-----------------------------------------------------------

\newcommand{\bb}[1]{\mathbb{#1}}
\newcommand{\cl}[1]{\mathcal{#1}}
\newcommand{\fA}{\mathfrak{A}}
\newcommand{\fB}{\mathfrak{B}}
\newcommand{\Tr}{{\rm Tr}}
\newtheorem{thm}{Theorem}

%-----------------------------------------------------------
% Define the column width and poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% The separation I chose is 0.024 and I want 4 columns
% Then set onecolwid to be (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be 2*onecolwid + sepwid = 0.464
%-----------------------------------------------------------

\newlength{\sepwid}
\newlength{\smallsepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\halfcolwid}
\setlength{\paperwidth}{48in}
\setlength{\paperheight}{36in}
\setlength{\sepwid}{0.024\paperwidth}
\setlength{\smallsepwid}{0.013\paperwidth}
\setlength{\halfcolwid}{0.185\paperwidth}
\setlength{\onecolwid}{0.265\paperwidth}
\setlength{\twocolwid}{0.37\paperwidth}
% \setlength{\onecolwid}{0.22\paperwidth}
% \setlength{\twocolwid}{0.464\paperwidth}
\setlength{\topmargin}{-0.5in}
\usetheme{confposter}
\usepackage{exscale}

\usepackage[no-math]{fontspec}
\usepackage{etoolbox}
\usepackage{fancyvrb}

\defaultfontfeatures{Mapping=tex-text}
\setsansfont[BoldFont={Fira Sans}]{Fira Sans Light}
\setmonofont{Fira Mono}
\newfontfamily\ExtraLight{Fira Sans ExtraLight}
\newfontfamily\Light{Fira Sans Light}
\newfontfamily\Book{Fira Sans}
\newfontfamily\Medium{Fira Sans Medium}

\AtBeginEnvironment{tabular}{\setsansfont[BoldFont={Fira Sans}, Numbers={Monospaced}]{Fira Sans Light}}

\setbeamerfont{title}{family=\Book}
\setbeamerfont{author}{family=\ExtraLight}
\setbeamerfont{date}{family=\ExtraLight}

\setbeamerfont{section title}{family=\Book}

\setbeamerfont{block title}{family=\Book}
\setbeamerfont{block title alerted}{family=\Book}

\setbeamerfont{subtitle}{family=\Light}
\setbeamerfont{frametitle}{family=\Book, series=\scshape}

\setbeamerfont{alerted text}{family=\Book}

\setbeamerfont{caption}{}
\setbeamerfont{caption name}{family=\Book}

\setbeamerfont{description item}{family=\Book}

%-----------------------------------------------------------
% The next part fixes a problem with figure numbering. Thanks Nishan!
% When including a figure in your poster, be sure that the commands are typed in the following order:
% \begin{figure}
% \includegraphics[...]{...}
% \caption{...}
% \end{figure}
% That is, put the \caption after the \includegraphics
%-----------------------------------------------------------

\usecaptiontemplate{
\small
\structure{\insertcaptionname~\insertcaptionnumber:}
\insertcaption}

%-----------------------------------------------------------
% Define colours (see beamerthemeconfposter.sty to change these colour definitions)
%-----------------------------------------------------------

\setbeamercolor{block title}{fg=ngreen,bg=white}
\setbeamercolor{block body}{fg=black,bg=white}
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70}
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10}

%-----------------------------------------------------------
% Name and authors of poster/paper/research
%-----------------------------------------------------------

\title{Vertex Coloring and Applications}
\author{Shawn Seymour}
\institute{Advisor: Dr. Peh Ng}

%-----------------------------------------------------------
% Start the poster itself
%-----------------------------------------------------------
% The \rmfamily command is used frequently throughout the poster to force a serif font to be used for the body text
% Serif font is better for small text, sans-serif font is better for headers (for readability reasons)
%-----------------------------------------------------------

\begin{document}
\begin{frame}[t]
  \begin{columns}[t]												% the [t] option aligns the column's content at the top
    \begin{column}{\sepwid}\end{column}			% empty spacer column
    \begin{column}{\onecolwid}
      \begin{block}{Abstract}
        \small
        Consider the map of the 48 contiguous states in the USA, and suppose we want to color each state so that no two states that share a boundary have the same color. In general, we could represent every state with a vertex and draw an edge between two states that share a border. This problem can be modeled by a mathematical structure called a graph. A graph, denoted G = (V, E), is a set of vertices V and a set of edges E. The Vertex Coloring problem on G aims to find the minimum number of colors (the chromatic number) needed to color the vertices such that no two adjacent vertices have the same color. Vertex coloring can solve real-world problems such as finding the minimum number of time slots to schedule a final exam period so that no two courses (taken by the same student) are scheduled at the same final exam time slot. In general, there is no known efficient time algorithm to find the chromatic number of a graph and there will likely not be one. This class of problem is known in computer science as NP-hard. Hence, there is interest in finding heuristics or approximation algorithms to find the chromatic number. In this research, we present three heuristics used to find good approximate vertex colorings even though they may not give us the optimal minimum coloring of a graph. This is important as it allows us to approximately solve complex problems in minutes rather than hours. We present computational results comparing the efficiency (time and quality) of these heuristics.

      \end{block}
      \vskip2ex
      \begin{block}{Introduction \& Background}
        \begin{itemize}
          \setlength\itemsep{0.4em}
          \item A simple graph $G = (V, E)$ is an undirected graph containing no loops or multiple edges. For our purpose, we assume all graphs are simple.
          \item An edge, denoted $(v, u) \in E$, connects vertex $v$ to vertex $u$ where $v, u \in V$.
          \item Two vertices are said to be \emph{adjacent} if they are connected by an edge.
          \item The \emph{degree} of a vertex $v \in V$, denoted $d_G(v)$, is the number of edges incident to $v$. For example, $d_{G_1}(2) = 3$.
          \item The \emph{saturation degree} of a vertex $v \in V$ is the number of different colors used for vertices adjacent to it. \newline
          \begin{figure}[h]
            \centering
            \includegraphics[width=\linewidth]{figures/example1-full.pdf}
            \caption{Example proper vertex colorings of a graph $G_1$ where $\chi(G_1) = 2$}
          \end{figure}
          \vspace{0.4cm}
          \item A \emph{proper vertex coloring} assigns colors to each vertex of $G$ such that no two adjacent vertices share the same color.
          \item The \emph{chromatic number} of $G$, denoted $\chi(G)$, is the minimum number of colors needed to properly color $G$.
          \item The \emph{vertex coloring problem} (VCP) on $G$ is to find $\chi(G)$.
          \item The VCP has been shown by \cite{sharma} to be NP-hard based on a reduction from 3-SAT, a well-known NP-hard problem.
          \item This means the VCP cannot be optimally solved in polynomial-time, so we must use heuristics (approximation algorithms) to efficiently find ``good'', but non-optimal, solutions.
          \item The above definitions can be found in \cite{bondymurty}.
        \end{itemize}

      \end{block}
  \end{column}

  \begin{column}{\sepwid}\end{column}			% empty spacer column

  \begin{column}{\twocolwid}

    \begin{columns}
      \begin{column}{\halfcolwid}
        \begin{block}{Heuristics}
          We focused on comparing three heuristics for approximating the vertex coloring problem:
          \begin{itemize}
            \setlength\itemsep{0.4em}
            \item \textbf{Greedy}: Label each vertex in $V$ (i.e. $v_1,v_2,\ldots,v_n$). Iterate through the vertices and assign the smallest available color.
            \item \textbf{Welsh-Powell} \cite{welsh}: Label each vertex in $V$ (i.e. $v_1,v_2,\ldots,v_n$) such that the vertices are sorted in decreasing order of their degree (i.e. $d_G(v_1) \geq d_G(v_2) \geq \ldots \geq d_G(v_n) $). Iterate through the vertices and assign the smallest available color.
            \item \textbf{DSATUR} \cite{brelaz}: Choose an uncolored vertex $v \in V$ with maximum saturation degree. If there is a tie, choose the vertex with maximum degree. Color $v$ with the smallest available color. Repeat until there are no uncolored vertices.
          \end{itemize}
        \end{block}
      \end{column}

      \begin{column}{\smallsepwid}\end{column}			% empty spacer column

      \begin{column}{\halfcolwid}
        \vspace{-0.7cm}
        \begin{block}{Methods}
          The three heuristics were implemented in Java and we ran simulations according to the following:
          \begin{itemize}
            \setlength\itemsep{0.4em}
            \item Graphs were generated based on the \emph{Erdős–Rényi} model. This model generates a random graph with approximately $p$ percent of edges.
            \item We focused on graphs with $90\%, 91\%, \ldots, 99\%$ edges.
            \item We ran simulations on 4 sets of graphs: 50 vertices, 100 vertices, 250 vertices, and 500 vertices.
            \item We focused on comparing the quality (number of colors used) of each heuristic. We calculated this using \emph{percent difference}. We also compared the running time of the three heuristics.
          \end{itemize}
        \end{block}
      \end{column}
    \end{columns}

    % \begin{block}{Heuristics \& Methods}
    %   We focused on comparing three heuristics for approximating the VCP:
    %   \begin{itemize}
    %     \setlength\itemsep{0.4em}
    %     \item \textbf{Greedy}: Label each vertex in $V$ (i.e. $v_1,v_2,\ldots,v_n$). Iterate through the vertices and assign the smallest available color.
    %     \item \textbf{Welsh-Powell} \cite{welsh}: Label each vertex in $V$ (i.e. $v_1,v_2,\ldots,v_n$) such that the vertices are sorted in decreasing order of their degree (i.e. $d_G(v_1) \geq d_G(v_2) \geq \ldots \geq d_G(v_n) $). Iterate through the vertices and assign the smallest available color.
    %     \item \textbf{DSATUR} \cite{brelaz}: Choose an uncolored vertex $v \in V$ with maximum saturation degree. If there is a tie, choose the vertex with maximum degree. Color $v$ with the smallest available color. Repeat until there are no uncolored vertices.
    %   \end{itemize}
    %
    %   \vspace{0.5cm}
    %
    %   We implemented these three heuristics in Java and created a web application for coloring graphs and running simulations of these heuristics on many graphs:
    %   \vspace{-1.3cm}
    %   \begin{itemize}
    %     \setlength\itemsep{0.4em}
    %     \item We generated graphs based on the \emph{Erdős–Rényi} model. This model generates a graph with approximately $p$ percent of edges.
    %     \item We initially looked at graphs with $0\%, 10\%, \ldots, 100\%$ edges.
    %     \begin{itemize}
    %       \item $0\%$ and $100\%$ are simple to color and uninteresting.
    %       \item The heuristics did not provide interesting comparisons in this range.
    %     \end{itemize}
    %     \item We then looked at graphs with $90\%, 91\%, \ldots, 99\%$ edges.
    %     \begin{itemize}
    %       \item This became interesting: large increases in number of colors used.
    %       \item We decided to focus on this range of graphs for our simulations.
    %     \end{itemize}
    %     \item We ran simulations on 4 sets of graphs: 50 vertices, 100 vertices, 250 vertices, and 500 vertices.
    %     \item We focused on comparing the quality (number of colors used) of each heuristic. We calculated this using \emph{percent difference}.
    %   \end{itemize}

    \vskip2ex
    \begin{block}{Results}
      \begin{itemize}
        \item The heuristics were compared based on the percent difference of four properties: average colors used, minimum colors used, maximum colors used, and running time. Two example comparisons are shown in Figure \ref{fig:compare-250} and Figure \ref{fig:compare-time-250}.
      \end{itemize}
      \begin{columns}
        \begin{column}{0.75\linewidth}
          \begin{figure}
            \includegraphics[width=\linewidth]{figures/comparison1-250.pdf}
            \caption{Percent difference of average colors used for 250 vertices}\label{fig:compare-250}
          \end{figure}
          \begin{figure}
            \includegraphics[width=\linewidth]{figures/comparison3-250.pdf}
            \caption{Percent difference of running time for 250 vertices}\label{fig:compare-time-250}
          \end{figure}
        \end{column}

      \end{columns}
    \end{block}

  \end{column}
  \begin{column}{\sepwid}\end{column}			% empty spacer column
  \begin{column}{\onecolwid}
    \vskip1.5ex

    \begin{figure}[h]
      \centering
      \includegraphics[width=\linewidth]{figures/comparison4.pdf}
      \caption{Percent difference of avg. colors used as vertices increase (93\% edges)}\label{fig:compare-min-250}
    \end{figure}

    \vspace{-0.2cm}

    \begin{itemize}
      \setlength\itemsep{0.4em}
      \rmfamily
      \item As the number of vertices increased, the percent difference between Greedy and Welsh-Powell as well as Greedy and DSATUR decreased. We also noticed that the percent difference between Welsh-Powell and DSATUR increased. This is shown in Figure \ref{fig:compare-min-250}.
      \item By looking at our comparisons, we saw that there is a \emph{partial order}.
      \begin{itemize}
        \item Welsh-Powell on average used $4-5\%$ less colors than Greedy.
        \item DSATUR on average used $6-7\%$ less colors than Greedy.
        \item DSATUR on average used $1-2\%$ less colors than Welsh-Powell.
      \end{itemize}
      \item This comparison stayed consistent when we looked at both minimum and maximum number of colors used.
      \item This led us to the following partial order based on colors used:
      \vspace{0.1cm}
      \begin{center}
        \textbf{Greedy $\leq$ Welsh-Powell $\leq$ DSATUR}
      \end{center}
      \item By looking at our running time comparisons, we saw:
      \begin{itemize}
        \item Welsh-Powell on average was $90-110\%$ slower than Greedy.
        \item DSATUR on average was $160-180\%$ slower than Greedy.
        \item DSATUR on average was $80-100\%$ slower than Welsh-Powell.
      \end{itemize}
    \end{itemize}

    \begin{block}{Conclusion}
      \begin{itemize}
        \item Although we were not able to compare the heuristics colorings to the optimal coloring, we were able to determine what heuristics gave the closest to optimal colors.
        \item Based on our results, we would suggest using the Welsh-Powell heuristic on most graphs.
        \begin{itemize}
          \item We decided that it would be worth 90-100\% more time to use 4-5\% less colors over Greedy. As most of these simulations ran in seconds, it is worth it.
          \item we decided it is \emph{not} worth using another 80-100\% more time to only use 1-2\% less colors if we use DSATUR over Welsh-Powell.
        \end{itemize}
      \end{itemize}
    \end{block}

    \begin{block}{Acknowledgements \& Information}
      \small{Special thanks to Peh Ng for her support, feedback, and insightful ideas throughout this project. Source code: \texttt{\footnotesize https://devshawn.com/coloring}.}
    \end{block}

    \begin{block}{References}
      \footnotesize{\rmfamily{\begin{thebibliography}{99}
      \setlength\parskip{0em}
      \setlength\itemsep{0em}

        \bibitem{bondymurty}
        Bondy, J.A. and U.S.R. Murty [1976],
        \emph{Graph theory with applications},
        American Elsevier Publishing, New York, NY.
        \bibitem{brelaz}
        Brélaz, D. [1979]. \emph{New methods to color the vertices of a graph}. Communications of the ACM, 22(4), pp.251-256.
        \bibitem{sharma}
        Sharma, P. C. and Chaudhari, N.S. [2012], \emph{A new reduction from 3-SAT to graph k-colorability for frequency assignment problem}, Int. J. Comp. Applic, 23-27.
        \bibitem{welsh}
        Welsh, D. J. and Powell, M. B. [1967], \emph{An upper bound for the chromatic number of a graph and its application to timetabling problems}, The Computer Journal, 10(1), 85-86.
      \end{thebibliography}}}
    \end{block}

  \end{column}
  \begin{column}{\sepwid}\end{column}			% empty spacer column
 \end{columns}
\end{frame}
\end{document}

### 2017-04-13 poster/poster.tex
More work on poster

@@
 \documentclass[final]{beamer}
-\usepackage[scale=1.07]{beamerposter}
+\usepackage[scale=1]{beamerposter}
 \usepackage{graphicx}			% allows us to import images
@@
       \begin{block}{Introduction \& Background}
-        \begin{itemize}
-          \setlength\itemsep{0.4em}
-          \item A simple graph $G = (V, E)$ is an undirected graph containing no loops or multiple edges. For our purpose, we assume all graphs are simple.
-          \item An edge, denoted $(v, u) \in E$, connects vertex $v$ to vertex $u$ where $v, u \in V$.
-          \item Two vertices are said to be \emph{adjacent} if they are connected by an edge.
-          \item The \emph{degree} of a vertex $v \in V$, denoted $d_G(v)$, is the number of edges incident to $v$. For example, $d_{G_1}(2) = 3$.
-          \item The \emph{saturation degree} of a vertex $v \in V$ is the number of different colors used for vertices adjacent to it. \newline
+        A simple graph $G = (V, E)$ is an undirected graph containing no loops or multiple edges.
+For our purpose, we assume all graphs are simple.
+An edge, denoted $(v, u) \in E$, connects vertex $v$ to vertex $u$ where $v, u \in V$.
+Two vertices are said to be \emph{adjacent} if they are connected by an edge.
+The \emph{degree} of a vertex $v \in V$, denoted $d_G(v)$, is the number of edges incident to $v$.
+For example, $d_{G_1}(2) = 3$.
+The \emph{saturation degree} of a vertex $v \in V$ is the number of different colors used for vertices adjacent to it.
+\newline
         \begin{figure}[h]
           \centering
-            \includegraphics[width=\linewidth]{figures/example1-full.pdf}
+          \includegraphics[width=\linewidth]{figures/huge.pdf}
           \caption{Example proper vertex colorings of a graph $G_1$ where $\chi(G_1) = 2$}
@@
         \vspace{0.4cm}
-          \item A \emph{proper vertex coloring} assigns colors to each vertex of $G$ such that no two adjacent vertices share the same color.
-          \item The \emph{chromatic number} of $G$, denoted $\chi(G)$, is the minimum number of colors needed to properly color $G$.
-          \item The \emph{vertex coloring problem} (VCP) on $G$ is to find $\chi(G)$.
-          \item The VCP has been shown by \cite{sharma} to be NP-hard based on a reduction from 3-SAT, a well-known NP-hard problem.
-          \item This means the VCP cannot be optimally solved in polynomial-time, so we must use heuristics (approximation algorithms) to efficiently find ``good'', but non-optimal, solutions.
-          \item The above definitions can be found in \cite{bondymurty}.
-        \end{itemize}
+        A \emph{proper vertex coloring} assigns colors to each vertex of $G$ such that no two adjacent vertices share the same color.
+The \emph{chromatic number} of $G$, denoted $\chi(G)$, is the minimum number of colors needed to properly color $G$.
+The \emph{vertex coloring problem} (VCP) on $G$ is to find $\chi(G)$.\newline
 
+        The VCP has been shown by \cite{sharma} to be NP-hard based on a reduction from 3-SAT, a well-known NP-hard problem.
+This means the VCP cannot be optimally solved in polynomial-time, so we must use heuristics (approximation algorithms) to efficiently find ``good'', but non-optimal, solutions.
+
+        \vspace{0.65cm}
+
+        \begin{figure}[h]
+          \centering
+          \includegraphics[width=\linewidth]{figures/compare-all.pdf}
+          \caption{Greedy, Welsh-Powell, and DSATUR vertex colorings of a graph $G_2$, respectively}\label{fig:all}
+        \end{figure}
+
+      \end{block}
+
+      \vskip1.8ex
+
+      \begin{block}{Acknowledgements \& Information}
+        \small{Special thanks to Peh Ng for her support, feedback, and insightful ideas throughout this project.
+Thanks to Engin Sungur and Humza Haider for helping with the statistical analysis and data plotting.
+Source code: \texttt{\footnotesize https://devshawn.com/coloring}.}
       \end{block}
@@
             \item Graphs were generated based on the \emph{Erdős–Rényi} model. This model generates a random graph with approximately $p$ percent of edges.
-            \item We focused on graphs with $90\%, 91\%, \ldots, 99\%$ edges.
-            \item We ran simulations on 4 sets of graphs: 50 vertices, 100 vertices, 250 vertices, and 500 vertices.
-            \item We focused on comparing the quality (number of colors used) of each heuristic.
-We calculated this using \emph{percent difference}.
-We also compared the running time of the three heuristics.
+            \item We focused on graphs with $90\%, 91\%, \ldots, 99\%$ edges. These were the most interesting.
+            \item Simulations were run on 4 sets of graphs: 50 vertices, 100 vertices, 250 vertices, and 500 vertices.
+            \item Each heuristic can produce a different amount of colors used, as shown in Figure \ref{fig:all}.
+Thus, we focused on comparing the quality (number of colors used) and the efficiency (running time) of the three heuristics.
           \end{itemize}
@@
     \end{columns}
-
-    % \begin{block}{Heuristics \& Methods}
-    %   We focused on comparing three heuristics for approximating the VCP:
-    %   \begin{itemize}
-    %     \setlength\itemsep{0.4em}
-    %     \item \textbf{Greedy}: Label each vertex in $V$ (i.e. $v_1,v_2,\ldots,v_n$). Iterate through the vertices and assign the smallest available color.
-    %     \item \textbf{Welsh-Powell} \cite{welsh}: Label each vertex in $V$ (i.e.
-$v_1,v_2,\ldots,v_n$) such that the vertices are sorted in decreasing order of their degree (i.e.
-$d_G(v_1) \geq d_G(v_2) \geq \ldots \geq d_G(v_n) $).
-Iterate through the vertices and assign the smallest available color.
-    %     \item \textbf{DSATUR} \cite{brelaz}: Choose an uncolored vertex $v \in V$ with maximum saturation degree.
-If there is a tie, choose the vertex with maximum degree.
-Color $v$ with the smallest available color.
-Repeat until there are no uncolored vertices.
-    %   \end{itemize}
-    %
-    %   \vspace{0.5cm}
-    %
-    %   We implemented these three heuristics in Java and created a web application for coloring graphs and running simulations of these heuristics on many graphs:
-    %   \vspace{-1.3cm}
-    %   \begin{itemize}
-    %     \setlength\itemsep{0.4em}
-    %     \item We generated graphs based on the \emph{Erdős–Rényi} model. This model generates a graph with approximately $p$ percent of edges.
-    %     \item We initially looked at graphs with $0\%, 10\%, \ldots, 100\%$ edges.
-    %     \begin{itemize}
-    %       \item $0\%$ and $100\%$ are simple to color and uninteresting.
-    %       \item The heuristics did not provide interesting comparisons in this range.
-    %     \end{itemize}
-    %     \item We then looked at graphs with $90\%, 91\%, \ldots, 99\%$ edges.
-    %     \begin{itemize}
-    %       \item This became interesting: large increases in number of colors used.
-    %       \item We decided to focus on this range of graphs for our simulations.
-    %     \end{itemize}
-    %     \item We ran simulations on 4 sets of graphs: 50 vertices, 100 vertices, 250 vertices, and 500 vertices.
-    %     \item We focused on comparing the quality (number of colors used) of each heuristic. We calculated this using \emph{percent difference}.
-    %   \end{itemize}
 
@@
 
-    \begin{block}{Acknowledgements \& Information}
-      \small{Special thanks to Peh Ng for her support, feedback, and insightful ideas throughout this project. Source code: \texttt{\footnotesize https://devshawn.com/coloring}.}
-    \end{block}
-
     \begin{block}{References}
-      \footnotesize{\rmfamily{\begin{thebibliography}{99}
+      %footnotesize
+      \small{\rmfamily{\begin{thebibliography}{99}
       \setlength\parskip{0em}

### 2017-04-16 poster/poster.tex
Lots of work on poster

@@
 This problem can be modeled by a mathematical structure called a graph.
-A graph, denoted G = (V, E), is a set of vertices V and a set of edges E.
-The Vertex Coloring problem on G aims to find the minimum number of colors (the chromatic number) needed to color the vertices such that no two adjacent vertices have the same color.
+A graph, denoted $G = (V, E)$, is a set of vertices $V$ and a set of edges $E$.
+The Vertex Coloring problem on $G$ aims to find the minimum number of colors (the chromatic number) needed to color the vertices such that no two adjacent vertices have the same color.
 Vertex coloring can solve real-world problems such as finding the minimum number of time slots to schedule a final exam period so that no two courses (taken by the same student) are scheduled at the same final exam time slot.
+
+        \vspace{0.4cm}
         In general, there is no known efficient time algorithm to find the chromatic number of a graph and there will likely not be one.
@@
             \item Graphs were generated based on the \emph{Erdős–Rényi} model. This model generates a random graph with approximately $p$ percent of edges.
-            \item We focused on graphs with $90\%, 91\%, \ldots, 99\%$ edges. These were the most interesting.
-            \item Simulations were run on 4 sets of graphs: 50 vertices, 100 vertices, 250 vertices, and 500 vertices.
+            \item We analyzed graphs with $10\%, 20\%, \ldots, 90\%$ edges (type). This gave us both sparse and dense graphs.
+            \item Simulations were run on graphs with 5 different amount of vertices: 50, 100, 250, 500, 1000.
             \item Each heuristic can produce a different amount of colors used, as shown in Figure \ref{fig:all}.
@@
     \begin{block}{Results}
-      \begin{itemize}
-        \item The heuristics were compared based on the percent difference of four properties: average colors used, minimum colors used, maximum colors used, and running time.
-Two example comparisons are shown in Figure \ref{fig:compare-250} and Figure \ref{fig:compare-time-250}.
-      \end{itemize}
+      We used statistical analysis to compare the number of colors used and the running time of each heuristic based on edge percentage (type) and then also looked at how this changed as the number of vertices changed.
+      \vspace{0cm}
       \begin{columns}
-        \begin{column}{0.75\linewidth}
+        \begin{column}{0.85\linewidth}
           \begin{figure}
-            \includegraphics[width=\linewidth]{figures/comparison1-250.pdf}
-            \caption{Percent difference of average colors used for 250 vertices}\label{fig:compare-250}
+            \includegraphics[width=\linewidth]{figures/plot-colors.pdf}
+            \caption{Comparing number of colors used for the three heuristics}\label{fig:compare-colors}
           \end{figure}
           \begin{figure}
-            \includegraphics[width=\linewidth]{figures/comparison3-250.pdf}
-            \caption{Percent difference of running time for 250 vertices}\label{fig:compare-time-250}
+            \includegraphics[width=\linewidth]{figures/plot-time.pdf}
+            \caption{Comparing running time of each heuristic for each edge percentage}\label{fig:compare-time}
           \end{figure}
@@
       \centering
-      \includegraphics[width=\linewidth]{figures/comparison4.pdf}
-      \caption{Percent difference of avg. colors used as vertices increase (93\% edges)}\label{fig:compare-min-250}
+      \includegraphics[width=\linewidth]{figures/plot-time-2.pdf}
+      \caption{Variance of running time of each heuristic based on edge percentage}\label{fig:compare-time-2}
     \end{figure}
@@
 
+    \rmfamily
+    \textbf{Color usage results}: \newline
+    Based on computational data of the amount of colors used, we found that there is a \emph{partial order}. This is shown in Figure \ref{fig:compare-colors}.
     \begin{itemize}
-      \setlength\itemsep{0.4em}
-      \rmfamily
-      \item As the number of vertices increased, the percent difference between Greedy and Welsh-Powell as well as Greedy and DSATUR decreased.
-We also noticed that the percent difference between Welsh-Powell and DSATUR increased.
-This is shown in Figure \ref{fig:compare-min-250}.
-      \item By looking at our comparisons, we saw that there is a \emph{partial order}.
-      \begin{itemize}
-        \item Welsh-Powell on average used $4-5\%$ less colors than Greedy.
-        \item DSATUR on average used $6-7\%$ less colors than Greedy.
-        \item DSATUR on average used $1-2\%$ less colors than Welsh-Powell.
+      \item Greedy uses on average 2.52 colors more than Welsh-Powell.
+      \item Welsh-Powell uses on average 2.79 colors more than DSATUR.
     \end{itemize}
-      \item This comparison stayed consistent when we looked at both minimum and maximum number of colors used.
-      \item This led us to the following partial order based on colors used:
+    This led us to the following partial order based on colors used:
     \vspace{0.1cm}
@@
     \end{center}
-      \item By looking at our running time comparisons, we saw:
+    As vertices increased, the partial order defined above stays consistent. DSATUR always used the least colors while Greedy always used the most.\newline
+
+    \textbf{Running time results}: \newline
+    Based on computational data of the running time, we also found a \emph{partial order}, as shown in Figure \ref{fig:compare-time}.
     \begin{itemize}
-        \item Welsh-Powell on average was $90-110\%$ slower than Greedy.
-        \item DSATUR on average was $160-180\%$ slower than Greedy.
-        \item DSATUR on average was $80-100\%$ slower than Welsh-Powell.
+      \item Welsh-Powell is significantly slower than Greedy.
+      \item DSATUR is significantly slower than Welsh-Powell.
     \end{itemize}
+    \begin{center}
+      \textbf{Greedy $\leq$ Welsh-Powell $\leq$ DSATUR}
+    \end{center}
+    As the edge percentage increases:
+    \begin{itemize}
+      \item DSATUR running time increases.
+      \item Greedy and Welsh-Powell running time increases until 50\% edges then decreases until 90\% edges.
+      \item The difference in running time between DSATUR and Greedy as well as DSATUR and Welsh-Powell grows larger and larger.
     \end{itemize}
@@
     \begin{block}{Conclusion}
-      \begin{itemize}
-        \item Although we were not able to compare the heuristics colorings to the optimal coloring, we were able to determine what heuristics gave the closest to optimal colors.
-        \item Based on our results, we would suggest using the Welsh-Powell heuristic on most graphs.
-        \begin{itemize}
-          \item We decided that it would be worth 90-100\% more time to use 4-5\% less colors over Greedy. As most of these simulations ran in seconds, it is worth it.
-          \item we decided it is \emph{not} worth using another 80-100\% more time to only use 1-2\% less colors if we use DSATUR over Welsh-Powell.
-        \end{itemize}
-      \end{itemize}
+      Although we were not able to compare the heuristics colorings to the optimal coloring, we were able to determine which heuristics gave the closest to optimal colors.
+Based on our results, we suggest using \textbf{Welsh-Powell} for most graphs.
+We believe that it is worth the extra time Welsh-Powell takes compared to Greedy but not worth the extra time DSATUR takes compared to Welsh-Powell.
     \end{block}

### 2017-04-17 poster/poster.tex
Making more changes to poster

@@
             \setlength\itemsep{0.4em}
-            \item Graphs were generated based on the \emph{Erdős–Rényi} model. This model generates a random graph with approximately $p$ percent of edges.
+            \item Graphs were generated based on the \emph{Erdős–Rényi} model \cite{erdos}. This model generates a random graph with approximately $p$ percent of edges.
             \item We analyzed graphs with $10\%, 20\%, \ldots, 90\%$ edges (type). This gave us both sparse and dense graphs.
@@
   \begin{column}{\onecolwid}
-    \vskip1.5ex
+    \vskip0.5ex
 
@@
       \centering
-      \includegraphics[width=\linewidth]{figures/plot-time-2.pdf}
+      \includegraphics[width=0.9\linewidth]{figures/plot-time-2.pdf}
       \caption{Variance of running time of each heuristic based on edge percentage}\label{fig:compare-time-2}
@@
     \textbf{Color usage results}: \newline
-    Based on computational data of the amount of colors used, we found that there is a \emph{partial order}. This is shown in Figure \ref{fig:compare-colors}.
+    Based on computational data of the amount of colors used, as shown in Figure \ref{fig:compare-colors}, we found:
     \begin{itemize}
@@
     \end{itemize}
-    This led us to the following partial order based on colors used:
+    This led us to the following partial order based on colors used, where $A \preceq B$ means heuristic $A$ uses significantly more colors than heuristic $B$.
     \vspace{0.1cm}
     \begin{center}
-      \textbf{Greedy $\leq$ Welsh-Powell $\leq$ DSATUR}
+      \textbf{Greedy $\preceq$ Welsh-Powell $\preceq$ DSATUR}
     \end{center}
-    As vertices increased, the partial order defined above stays consistent. DSATUR always used the least colors while Greedy always used the most.\newline
+    As vertices increased, the partial order defined above stays consistent. DSATUR always used the least colors while Greedy always used the most.
+
+    \vspace{0.4cm}
 
     \textbf{Running time results}: \newline
-    Based on computational data of the running time, we also found a \emph{partial order}, as shown in Figure \ref{fig:compare-time}.
+    Based on computational data of the running time, as shown in Figure \ref{fig:compare-time}, we found:
     \begin{itemize}
@@
     \end{itemize}
+    This led us to the following partial order based on running time, where $A \preceq B$ means heuristic $A$ uses significantly less time than heuristic $B$.
     \begin{center}
-      \textbf{Greedy $\leq$ Welsh-Powell $\leq$ DSATUR}
+      \textbf{Greedy $\preceq$ Welsh-Powell $\preceq$ DSATUR}
     \end{center}
@@
       Although we were not able to compare the heuristics colorings to the optimal coloring, we were able to determine which heuristics gave the closest to optimal colors.
-Based on our results, we suggest using \textbf{Welsh-Powell} for most graphs.
-We believe that it is worth the extra time Welsh-Powell takes compared to Greedy but not worth the extra time DSATUR takes compared to Welsh-Powell.
+Based on our results, we suggest using \textbf{Welsh-Powell} for most graphs because the extra time needed from Greedy to Welsh-Powell to achieve better accuracy is not as significant compared to the extra time needed from Welsh-Powell to DSATUR.
     \end{block}
@@
       %footnotesize
-      \small{\rmfamily{\begin{thebibliography}{99}
+      \footnotesize{\rmfamily{\begin{thebibliography}{99}
       \setlength\parskip{0em}
       \setlength\itemsep{0em}
-
+        \vspace{-0.3cm}
         \bibitem{bondymurty}
-        Bondy, J.A. and U.S.R. Murty [1976],
-        \emph{Graph theory with applications},
-        American Elsevier Publishing, New York, NY.
+        Bondy, J.A. and Murty, U.S.R. [1976], \emph{Graph theory with applications}, American Elsevier Publishing, New York, NY.
         \bibitem{brelaz}
         Brélaz, D. [1979]. \emph{New methods to color the vertices of a graph}. Communications of the ACM, 22(4), pp.251-256.
+        \bibitem{erdos}
+        Erdös, P. and A. Rényi [1959], \emph{On random graphs, I}, Publicationes Mathematicae (Debrecen) 6: 290-297.
         \bibitem{sharma}

### 2017-04-25 library/src/main/java/com/devshawn/coloring/library/FindGraph.java
Adding data files and some helper classes

@@ -1,0 +1,30 @@
+package com.devshawn.coloring.library;
+
+import java.util.Arrays;
+import java.util.Random;
+
+public class FindGraph {
+
+    public static void main(String[] args) {
+        ColoringResult greedy;
+        ColoringResult wp;
+        ColoringResult dsatur;
+        int[][] graph;
+
+        do {
+            ColoringModule coloringModule = new ColoringModule();
+            double prob = new Random().nextInt(11) / (double) 100;
+            graph = GraphGenerator.simple(8, prob);
+            coloringModule.setGraph(graph);
+            greedy = coloringModule.applyHeuristic(ColoringHeuristic.GREEDY);
+            wp = coloringModule.applyHeuristic(ColoringHeuristic.WELSH_POWELL);
+            dsatur = coloringModule.applyHeuristic(ColoringHeuristic.DSATUR);
+        } while (wp.getColoringNumber() == dsatur.getColoringNumber() || greedy.getColoringNumber() == wp.getColoringNumber() || greedy.getColoringNumber() == dsatur.getColoringNumber());
+
+        System.out.println(greedy.getColoringNumber());
+        System.out.println(wp.getColoringNumber());
+        System.out.println(dsatur.getColoringNumber());
+        System.out.println(Arrays.deepToString(graph));
+    }
+
+}

### 2017-04-25 library/src/main/java/com/devshawn/coloring/library/SimulationRunner.java

@@ -1,0 +1,94 @@
+package com.devshawn.coloring.library;
+
+import java.io.*;
+
+public class SimulationRunner {
+    public static void main(String[] args) throws FileNotFoundException, IOException {
+
+        final int EDGE_PERCENTAGE_START = 10;
+        final int EDGE_PERCENTAGE_END = 90;
+        final int EDGE_PERCENTAGE_INCREMENT = 10;
+        final int RUNS = 1000;
+        final int[] vertices = {50, 100, 250, 500, 1000};
+
+
+        File file = new File("full.csv");
+        file.delete();
+        PrintWriter pw = new PrintWriter(new FileWriter(file, true));
+
+        if(file.length() == 0) {
+            printHeaders(pw);
+        }
+
+        pw.flush();
+
+        // warmup
+
+        for(int run = 0; run < 1000; run++) {
+            ColoringModule coloringModule = new ColoringModule();
+            int[][] graph = GraphGenerator.simple(100, 90 / (double) 100);
+            coloringModule.setGraph(graph);
+            ColoringResult greedy = coloringModule.applyHeuristic(ColoringHeuristic.GREEDY);
+            ColoringResult wp = coloringModule.applyHeuristic(ColoringHeuristic.WELSH_POWELL);
+            ColoringResult dsatur = coloringModule.applyHeuristic(ColoringHeuristic.DSATUR);
+        }
+
+        // actual
+
+        for(int vertex : vertices) {
+            System.out.println("Starting vertex: " + vertex);
+
+            for(int edgePercentage = EDGE_PERCENTAGE_START; edgePercentage <= EDGE_PERCENTAGE_END; edgePercentage += EDGE_PERCENTAGE_INCREMENT) {
+                System.out.println("\tStarting edge percentage: " + edgePercentage);
+
+                for(int run = 0; run < RUNS; run++) {
+                    ColoringModule coloringModule = new ColoringModule();
+                    int[][] graph = GraphGenerator.simple(vertex, edgePercentage / (double) 100);
+                    coloringModule.setGraph(graph);
+                    ColoringResult greedy = coloringModule.applyHeuristic(ColoringHeuristic.GREEDY);
+                    ColoringResult wp = coloringModule.applyHeuristic(ColoringHeuristic.WELSH_POWELL);
+                    ColoringResult dsatur = coloringModule.applyHeuristic(ColoringHeuristic.DSATUR);
+
+                    printHeuristic(pw, edgePercentage, vertex, greedy);
[…]

### 2017-04-26 work/paper.tex
Starting initial paper

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{titling}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[numbib,nottoc]{tocbibind}
\usepackage{indentfirst}

\setlength{\droptitle}{-1.5cm}

\newcounter{heuristic} \setcounter{heuristic}{0}
\renewcommand{\theheuristic}{\Alph{heuristic}}

\newenvironment{heuristic}[1]%
{
\refstepcounter{heuristic}
\subsection*{Heuristic \theheuristic}
}%
{}

\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother

\title{Vertex Coloring and Applications}
\date{May 5, 2017}
\author{Shawn Seymour\\ University of Minnesota Morris}

\newtheorem{prop}{Proposition}
\newtheorem{theorem}{Theorem}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

\begin{document}

\maketitle

\setlength{\parskip}{0.3cm}

\section{Introduction}
Consider the map of the 48 contiguous states in the USA. Suppose we would like to color each state such that no two states that share a boundary have the same color. How many colors would it take to do this? This problem can be modeled as a \emph{graph}. In general, we could represent every state with a \emph{vertex} and draw an \emph{edge} between two states that share a border. A graph, denoted $G = (V, E)$, is a set of vertices $V$ and a set of edges $E$. A \emph{simple graph} is a loopless, undirected graph where no two edges connect the same pair of vertices. For our purpose, assume all graphs are simple graphs. This problem, along with many others, can be solved using vertex coloring.

\begin{figure}[H]
\centering
\includegraphics[width=0.5\linewidth]{figures/map-colors.pdf}
\caption{A proper coloring of the USA}\label{fig:map}
\end{figure}

A \emph{vertex coloring} of a graph $G$ is an assignment of colors to each vertex of a graph. A \emph{proper vertex coloring} assigns colors to a graph $G$ such that no two adjacent vertices share the same color. This can be described as a function $f : V \rightarrow S = \{1, 2, \ldots, k\}$ such that $\forall u,w \in V$, if $(u,w) \in E$, then $f(u) \neq f(w)$. Note that this constraint is the same constraint we applied to our map example. This means we could color our map according to our constraint with a proper vertex coloring.

The \emph{chromatic number}, denoted $\chi(G)$, is the minimum number of colors needed to have a proper vertex coloring of a graph $G$. The \emph{vertex coloring problem} (VCP), when given a graph $G$, is to find $\chi(G)$. By applying the vertex coloring problem to our map example, we can determine how many colors one would need to color it with regards to our constraint. As it turns out, it only takes \emph{four} colors to properly color the map of the USA. This is shown in Figure \ref{fig:map}.

\newpage

\bibliographystyle{abbrv}
\bibliography{references}

\end{document}

### 2017-05-01 work/paper.tex
More paper work

@@
 \usepackage[numbib,nottoc]{tocbibind}
-\usepackage{indentfirst}
+\usepackage{enumitem}
+\setlist[itemize]{topsep=0pt}
+\setlist[enumerate]{topsep=0pt}
+\setlength\parindent{0pt}
+\setlength{\droptitle}{-1.5cm}
 
-\setlength{\droptitle}{-1.5cm}
+\setlength{\abovedisplayskip}{0pt}
+\setlength{\belowdisplayskip}{0pt}
+\setlength{\abovedisplayshortskip}{0pt}
+\setlength{\belowdisplayshortskip}{0pt}
 
@@
 \author{Shawn Seymour\\ University of Minnesota Morris}
+
+\makeatletter
+\def\thm@space@setup{\thm@preskip=8pt
+\thm@postskip=0pt}
+\makeatother
 
@@
 \centering
-\includegraphics[width=0.5\linewidth]{figures/map-colors.pdf}
+\includegraphics[width=0.7\linewidth]{figures/map-colors.pdf}
 \caption{A proper coloring of the USA}\label{fig:map}
@@
 This means we could color our map according to our constraint with a proper vertex coloring.
+For our purpose, assume we mean proper vertex coloring when referring coloring a graph.
 
@@
 By applying the vertex coloring problem to our map example, we can determine how many colors one would need to color it with regards to our constraint.
-As it turns out, it only takes \emph{four} colors to properly color the map of the USA.
-This is shown in Figure \ref{fig:map}.
+An example coloring of our map is shown in Figure \ref{fig:map}.
+The \emph{k-Colorability problem} asks if a graph can be colored with $k$ colors.
+If a graph can be colored with $k$ colors, it is said to be \emph{k-colorable}.
+
+\section{Computational Complexity}
+Finding an optimal solution to the vertex coloring problem is not easy.
+As the VCP is an optimization problem with regards to a constraint, we know it will require heavy computation to find an optimal solution.
+In mathematics and computer science, we classify computationally complex problems based on type of problem they are and based on how many steps it takes to solve the problem with respect to its input size.
+
+\subsection{Classifying the Vertex Coloring Problem}
+We call any problem that can be answered with ``yes'' or ``no'' a \emph{decision problem}. To classify the vertex coloring problem, we must first introduce some classes, defined in \cite{moret}:
+
+\begin{itemize}
+  \item \textbf{P}: Given a decision problem, we can solve the problem in polynomial time.
+  \item \textbf{NP}: Given a decision problem, we can verify if a given solution is correct in polynomial time but cannot solve the problem in polynomial time.
+  \item \textbf{NP-hard}: Given any problem, it is at least as hard as NP problems. They do not have to be in NP, however, as they do not need to have solutions verifiable in polynomial time.
+  \item \textbf{NP-complete}: Given a decision problem, we can transform it to any other NP problem in polynomial time and still verify a given solution in polynomial time.
+\end{itemize}
+
+Take note that these classes are not exclusive: any member of P is also a member of NP.
+NP-complete problems can be thought of as the problems that are both in NP and NP-hard.
+Although it may seem like all problems that are NP are also NP-complete, this is not the case.
+There are a select few problems in NP that are not NP-complete or in P \cite{ladner1975structure}.
+
+\begin{prop}
+The vertex coloring problem is an NP-hard problem.
+\end{prop}
+
+It is NP-hard to compute the chromatic number of a graph.
+This can be proven by showing it is at least as hard as a problem in NP.
+This is typically done by a reduction (a transformation) to or from 3-SAT, which we'll examine later.
+This is proven in \cite{garey}.
+
+\begin{prop}\label{prop:k-color-np-complete}
+The k-Colorability problem, i.e. can a given graph be colored in k-colors, is NP-complete.
+\end{prop}
+
+To prove this, we need to reduce a known NP-complete problem to the vertex coloring problem in polynomial time.
+By reducing a known NP-complete problem to k-Colorability, we can conclude that k-Colorability is NP-complete as all NP-complete problems are reducible to all other NP-complete problems \cite{gareynp}.
+We will reduce the well-known 3-SAT problem first to the 3-Colorability problem.
+The 3-Colorability problem asks if a graph can be colored with 3 (or fewer) colors.
+Obviously, if a graph can be colored with 2 colors it can also be colored with 3.
+From here, the reduction from 3-Colorability to k-Colorability follows.
+
+\subsection{Reduction from 3-SAT to 3-Colorability}
+Before we prove Proposition \ref{prop:k-color-np-complete}, we must first introduce 3-SAT.
+The 3-SAT problem is an NP-complete problem that is a variation of the original NP-complete problem, boolean satisfiability (SAT).
+The 3-SAT problem is defined and proven to be NP-complete in \cite{gareynp}.
+3-SAT consists of a conjunction of a number of clauses.
+A clause is a disjunction of a given number of propositions or their negations.
+We will assume all of our satisfiability problems are in conjunctive normal form (CNF).
+3-SAT asks whether the variables can be assigned \emph{true} or \emph{false} in such a way that the whole expression evaluates to \emph{true}.
+We define SAT and 3-SAT in Definition \ref{def:sat}.
+
+\begin{definition}\label{def:sat}
+SAT is a conjunction of $m$ clauses, i.e.
+$C_1 \wedge C_2 \wedge \dots \wedge C_m$.
+Each clause is a disjunction of at most $n$ literals, i.e.
+$L_1 \vee L_2 \vee \dots \vee L_n$.
+Each literal can be a variable or negative of a variable, i.e.
+$x_1, \neg x_1, x_2, \neg x_2, \dots , x_k, \neg x_k$ where $k$ is the number of distinct literals.
+Each literal has a value of $true$ or $false$.
+3-SAT is where each clause has exactly 3 literals \((n = 3)\).
+\end{definition}
+
+To show a reduction from 3-SAT to 3-Colorability, we need to:
+
+\begin{enumerate}
+\item Given an instance of 3-SAT, construct an instance of 3-Colorability
+\item Show that if the resulting graph is 3-colorable, then the given 3-SAT problem is satisfied
+\end{enumerate}
+
+To show this reduction, we'll create an example 3-SAT problem and reduce it to 3-Colorability.
+This reduction will create a graph \(G = (V, E)\) having \((2n + 3m + 1)\) vertices and \((3n+6m)\) edges \cite{moret}.
+The process for creating the graph, described below, was taken from the proof given in \cite{sharma}.
+
+Let's introduce the following 3-SAT expression $E$:
+%
+\begin{align*}
+E = \left( x_1 \vee \neg x_2 \vee \neg x_3 \right) \wedge \left( \neg x_1 \vee x_2 \vee x_3 \right) \wedge \left( x_1 \vee x_2 \vee x_3 \right)
+\end{align*}
+%
+To create the graph of this 3-SAT problem, we:
+\begin{enumerate}
+\item Create a \emph{triangle} for each variable, i.e.
+\(x_i\).
+Each has a common vertex \(B\), called the base vertex.
+This base vertex will be colored a color, say \(p_1\), so that the other two vertices correspond to the variable and it's negation.
+\item Create a \emph{triangle} for each clause in the formula.
+\item Connect each vertex of a clause triangle to the corresponding literal vertex.
+\end{enumerate}
+
+Following the above steps, we generate the following graph $G$ from our 3-SAT expression $E$:
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.6]{images/3sat-1.png}
+\caption{\(E\) represented as a graph \(G\)}
+\end{figure}
 

### 2017-05-01 work/paper.tex
More paper work...

@@
 \refstepcounter{heuristic}
-\subsection*{Heuristic \theheuristic}
+\subsection*{Heuristic \theheuristic: #1}
+\vspace{-0.2cm}
 }%
@@
 
-\section{Computational Complexity}
+\section{Computational Complexity}\label{sec:complexity}
 Finding an optimal solution to the vertex coloring problem is not easy.
@@
 
-Following the above steps, we generate the following graph $G$ from our 3-SAT expression $E$:
+Following the above steps, we generate graph $G$, shown in Figure \ref{fig:3sat1}, from our 3-SAT expression $E$.
 
@@
 \includegraphics[scale=0.6]{images/3sat-1.png}
-\caption{\(E\) represented as a graph \(G\)}
+\caption{\(E\) represented as a graph \(G\)}\label{fig:3sat1}
 \end{figure}
+
+If we can properly color graph \(G\) with 3 colors, then the original 3-SAT is satisfiable and the truth values of the variables are denoted by the color they received.
+This graph is \emph{3-colorable} and the coloring is shown in Figure \ref{fig:3sat2}.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.6]{images/3sat-2.png}
+\caption{A possible 3-coloring of \(G\)}\label{fig:3sat2}
+\end{figure}
+
+As we see above, \emph{blue} is used for the base vertex and as part of some of the clause triangles.
+If we let \emph{yellow} denote \emph{true} and \emph{red} denote \emph{false}, we can see that our original 3-SAT is satisfied:
+%
+\begin{align*}
+E &= \left( T \vee \neg T \vee \neg F \right) \wedge \left( \neg T \vee T \vee F \right) \wedge \left(T \vee T \vee F \right) \\
+&= \left( T \vee F \vee T \right) \wedge \left( F \vee T \vee F \right) \wedge \left(T \vee T \vee F \right) \\
+&= T \wedge T \wedge T \\
+&= True
+\end{align*}
+%
+This shows an example of the reduction from 3-SAT to 3-Colorability.
+If \(G\) was unable to be colored with 3 colors, then we would know that \(E\) is unable to be satisfied.
+From here, we can reduce 3-Colorability to 4-Colorability, and then generalize that reduction and reduce 4-Colorability to k-Colorability.
+The proof and illustration of this can be found in \cite{sharma}.
+
+\section{Heuristics}
+As we have seen in section \ref{sec:complexity}, solving the vertex coloring problem requires heavy computation.
+This gives us incentive to use approximation algorithms, called heuristics, to find good, but not necessarily optimal, vertex colorings.
+We decided to analyze and compare a few of the popular heuristics for solving the VCP.
+The input for all heuristics we define is a simple graph, $G = (V, E)$.
+
+\heuristic{Greedy}\label{heuristic-greedy}
+
+Heuristic \ref{heuristic-greedy} is a simple greedy algorithm. It is defined in Algorithm \ref{alg:greedy}.
+
+\begin{algorithm}
+\caption{Greedy algorithm}\label{alg:greedy}
+\begin{algorithmic}[1]
+\State Label each vertex in $V$, i.e. $v_1, v_2, \dots, v_n$
+\For{each $v \in V$}
+\State Assign a color $p_i$ to $v_i$ using the smallest available $p_i$
+\EndFor
+\end{algorithmic}
+\end{algorithm}
+
+\heuristic{Welsh-Powell}\label{heuristic-welsh}
+
+Heuristic \ref{heuristic-welsh} is the greedy algorithm with degree sequencing.
+Before applying the greedy algorithm, it orders the vertices according to the decreasing value of their degree.
+This is also known as the Welsh-Powell algorithm, which is given in \cite{welsh}.
+It is defined in Algorithm \ref{alg:welsh}.
+
+\begin{algorithm}
+\caption{Welsh-Powell algorithm}\label{alg:welsh}
+\begin{algorithmic}[1]
+\State Label each vertex in $V$, i.e. $v_1, v_2, \dots, v_n$, such that $d_G(v_1) \geq d_G(v_2) \geq \dots \geq d_G(v_n)$
+\ForAll{$v \in V$}
+\State Assign a color $p_i$ to $v_i$ using the smallest available $p_i$
+\EndFor
+\end{algorithmic}
+\end{algorithm}
+
+\heuristic{DSATUR}\label{heuristic-dsatur}
+
+Heuristic \ref{heuristic-dsatur} is very similar to heuristic \ref{heuristic-welsh}, but colors a graoh based on the \emph{saturation} degree of vertices, which is defined in Definition \ref{def:sat} \cite{spinrad}.
+The original DSATUR algorithm was proposed by \cite{brelaz}.
+It is defined in Algorithm \ref{alg:dsatur}.
+
+\begin{definition}[Saturation degree]\label{def:sat}
+The saturation degree of a vertex is the number of different colors used for vertices adjacency to it.
+\end{definition}
+
+\begin{algorithm}
+\caption{DSATUR algorithm}\label{alg:dsatur}
+\begin{algorithmic}[1]
+\State Find the vertex $\in V$ with maximum degree and color it with color $p_1$
+\While{There are uncolored vertices}
+\State Choose an uncolored vertex with maximum saturation degree
+\State If there is a tie, choose the vertex with maximum degree
+\State Color the vertex with the least possible number
+\EndWhile
+\end{algorithmic}
+\end{algorithm}
+
+\heuristic{Maximal Independent Set}\label{heuristic-mis}
+
+Heuristic \ref{heuristic-mis} colors a graph by finding maximal independent sets of \(G\).
+A maximal independent set (MIS) is defined in Definition \ref{def:mis}.
+This is quite different than our other heuristics as it does not depend on degree.
+Ideally, we would want to find a maximum independent set, not just a maximal independent set.
+
+Finding a maximum independent set is NP-hard \cite{karp1972reducibility}, however, meaning we're including a different NP-hard in an already NP-hard problem.
+We chose to forgo this in favor of just finding a maximal independent set.
+We define the heuristic in Algorithm \ref{alg:mis}.
+
+\begin{definition}[Maximal independent set]\label{def:mis}
+An independent set is a subset $S$ of vertices in a graph such that no two vertices in the set are adjacent.
+$S$ is said to be maximal if no vertex outside the set can be added while still being an independent set.
+\end{definition}
+
+\begin{algorithm}
+\caption{Maximal independent set algorithm}\label{alg:mis}
+\begin{algorithmic}[1]
+\State $p \leftarrow 1$
+\While{$G$ is non-empty}
+\State Find a maximal independent set of $G$, i.e. $S_i$
+\State Color all vertices in $S_i$ with color $p$
+\State Let $G \leftarrow G[V \setminus S_i ] $
+\State $p \leftarrow p + 1$
+\EndWhile
+\end{algorithmic}
+\end{algorithm}
+
+\section{Initial Findings}
+Before comparing the heuristics, we first wanted to analyze the heuristics individually and look at the solutions they produce.
+
+\begin{prop}
+Heuristic \ref{heuristic-greedy} and Heuristic \ref{heuristic-welsh} produce a feasible, proper coloring of input graph \(G = (V, E)\).
+\end{prop}
+
+Let \(G = (V, E)\) be a simple graph.
+Assume that two vertices, say \(v_1, v_2 \in V\), are connected by an edge \(e_1 \in E\).
+Assume that \(v_1, v_2\) are colored the same color, say \(p_1\).
+We know that in both heuristics, we assign the smallest color \(p_i\) that is not being used by any of the neighboring vertices connected by an edge \(e_i\).
+WLOG, the heuristic would color \(v_1\) color \(p_1\).
+The next iteration, while vertex \(v_2\) is selected, the heuristic would see color \(p_1\) is assigned to a neighboring vertex \((v_1)\), and assign the next available color \(p_2\).
+Thus, we've reached a contradiction.
+
+\begin{prop}
+Heuristic \ref{heuristic-dsatur} produces a feasible, proper coloring of input graph \(G = (V, E)\).
+\end{prop}
+
+Let \(G = (V, E)\) be a simple graph.
+We color the vertex with the maximum degree color 1.
+If there are no uncolored vertices, we have properly colored $G$.
+If there are, then we find an uncolored vertex $v$ with maximum saturation degree.
+We then color $v$ with the smallest available color.
+This means we look at the set of all adjacent vertices to $v$ and find the smallest unused color in this set.
+This means we will never color an adjacent vertex with the same color, and thus, will always result in a proper coloring.
+
+\begin{prop}
+Heuristic \ref{heuristic-mis} produces a feasible, proper coloring of input graph \(G = (V, E)\).
+\end{prop}
+
+Let \(G = (V, E)\) be a simple graph.
+Let $p$, the current color, be 1.
+By definition \ref{def:mis}, we know that no two vertices in a maximal independent set are adjacent.
+Thus, when we color the vertices in the MIS with the smallest color \(p_1\) for the first iteration, we know that it is properly colored.
+We then remove the found maximal independent set and increment $p$ to 2.
+If $G$ is now empty, then $G$ is properly colored.
+If $G$ is not empty, then we repeat the process except we use color 2.
+Because we are always incrementing the color for each iteration, we know the MIS found will always be properly colored with a different color than the past MIS.
+
+Now that we know our heuristics produce feasible, proper colorings of a graph, how do we know they do not produce optimal solutions? We show they do not using counterexamples.
 

### 2017-05-02 work/paper.tex
Draft for paper

@@
 \usepackage{enumitem}
+\usepackage{multicol}
 \setlist[itemize]{topsep=0pt}
@@
 
+\begin{prop}
+Heuristic \ref{heuristic-greedy} and Heuristic \ref{heuristic-welsh} do not always produce an optimal solution, i.e. they do not always produce a minimum coloring of a graph.
+\end{prop}
+
+We will show this for \emph{heuristic \ref{heuristic-welsh}}.
+It is trivial to show the same for \emph{heuristic \ref{heuristic-greedy}} as \emph{heuristic \ref{heuristic-welsh}} includes the same steps as \emph{heuristic \ref{heuristic-greedy}}.
+We will construct a simple graph, \(G = (V, E)\), such that:
+%
+\begin{align}
+|V| &\geq 8 \\
+\chi(G) &> \chi^{*}(G)
+\end{align}
+%
+Let \(\chi(G)\) refer to the coloring of \(G\) generated by \emph{heuristic \ref{heuristic-welsh}}.
+Let \(\chi^{*}(G)\) be the optimal coloring of \(G\).
+We constructed a simple graph, \(G_1 = (V, E)\), such that \(\Delta(G_1) = 3\) and \(\delta(G_1) = 1\).
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{images/graph-1.png}
+\caption{Uncolored original graph \(G_1\)}
+\end{figure}
+
+By applying \emph{heuristic \ref{heuristic-welsh}}, we get the following coloring.
+The numbers indicate the order the vertices were colored in while applying the heuristic.
+Any vertex of the same degree got assigned arbitrarily.
+This results in \(\chi(G_1) = 3\).
+
+\begin{multicols}{2}
+  \begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.5]{images/graph-2.png}
+  \caption{Coloring of $G_1$ after applying heuristic \ref{heuristic-welsh}}
+  \end{figure}
+
+  \begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.5]{images/graph-3.png}
+  \caption{Optimal coloring of graph \(G_1\)}\label{fig:ex1-optimal}
+  \end{figure}
+\end{multicols}
+
+\begin{definition}[Bipartite graph]
+A bipartite graph is one whose vertex set can be partitioned into two subsets X and Y such that each edge has one end in X and one end in Y.
+\end{definition}
+
+We can see that $G_1$ is a bipartite graph, defined above by \cite{bondymurty}.
+Thus, \(G_1\) is 2-colorable as all bipartite graphs are 2-colorable \cite{asratian1998bipartite}.
+This means \(\chi^{*}(G_1) = 2\).
+\(G_1\) is an example graph that satisfies our given conditions \((1)\) and \((2)\).
+The optimal coloring is shown in Figure \ref{fig:ex1-optimal}.
+
+Although Heuristic \ref{heuristic-greedy} and Heuristic \ref{heuristic-welsh} do not always produce an optimal coloring for bipartite graphs, Heuristic \ref{heuristic-dsatur} always will produce an optimal coloring (2 colors).
+Heuristic \ref{heuristic-dsatur}, however, does not produce optimal colorings for all graphs.
+Both of these results are shown and proven in \cite{brelaz}.
+We chose not to create examples of these.
+
+\begin{prop}
+Heuristic \ref{heuristic-mis} also does not always produce an optimal solution, i.e. it does not always produce a minimum coloring of a graph.
+\end{prop}
+
+To show this, we will create another example that satisfies conditions \((1)\) and \((2)\) from earlier.
+We've constructed a graph \(G_2 = (V,E)\) for this example.
+This is shown Figure \ref{fig:ex2-uncolored}.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{images/mis-1.png}
+\caption{Uncolored original graph \(G_2\)}\label{fig:ex2-uncolored}
+\end{figure}
+
+By applying heuristic \ref{heuristic-mis} and finding maximal independent sets, we see that \(G_2\) is \emph{5-colorable}. This is shown Figure \ref{fig:ex2-colored}.
+
+\begin{multicols}{2}
+  \begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.5]{images/mis-2.png}
+  \caption{Coloring of \(G_2\) after applying heuristic \ref{heuristic-mis}}\label{fig:ex2-colored}
+  \end{figure}
+
+  \begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.5]{images/mis-3.png}
+  \caption{Optimal coloring of \(K\)}\label{fig:ex2-optimal}
+  \end{figure}
+\end{multicols}
+
+This is a non-optimal solution.
+We can see that the subgraph created by vertices \(\{5, 6, 7, 8\}\) is complete and thus we need at least 4 colors.
+This graph is \emph{4-colorable} however, meaning \(\chi^{*}(K) = 4\).
+Thus, we can see that heuristic \ref{heuristic-mis} does not always give us an optimal solution.
+The optimal coloring of $G_2$ is shown in Figure \ref{fig:ex2-optimal}.
+
+\begin{prop}
+Heuristic \ref{heuristic-greedy} and heuristic \ref{heuristic-welsh} have an upper bound of \(\Delta + 1\).
+\end{prop}
+
+To show this, we will create another example that satisfies conditions \((1)\) and \((2)\) from above.
+After doing some research into bipartite graphs, we learned that \emph{crown graphs} are excellent at showing how bad greedy heuristics can be, as shown and defined in \cite{kordecki}.
+
+\begin{definition}[Crown Graph]
+A crown graph \(CR_n = (V, E)\) is an undirected graph such that \(U = \{u_1, u_2, \dots, u_i\}\) and \(W = \{w_1, w_2, \dots, w_j\}\) where \(V = U \cup W\), \(U \cap W = \emptyset\), and \(|U| = |W|\).
+There is an edge from \(u_i\) to \(w_j\) whenever \(i \neq j\).
+A crown graph can also be viewed as a complete bipartite graph where the edges of a perfect matching have been removed.
+\end{definition}
+
+We constructed a simple crown graph, \(G_3 = (V, E)\). We've let \(|V| = 10\) for this example.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.38]{images/graph-4.png}
+\caption{Uncolored original graph \(G_3\)}
+\end{figure}
+
+We can see that \(\Delta(G_3) = 4\).
+We can also see \(d_{G_3}(v) = 4\) for all \(v \in V\).
+Thus, in both heuristics \ref{heuristic-greedy} and \ref{heuristic-welsh}, the algorithm would pick an order arbitrarily.
+We can show using this crown graph the worst-case scenario of these heuristics.
+
+\begin{multicols}{2}
+  \begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.38]{images/graph-5.png}
+  \caption{Worst-case coloring of \(G_3\)}\label{fig:ex3-colored}
+  \end{figure}
+
+  \begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.38]{images/graph-6.png}
+  \caption{Optimal coloring of \(H\)}\label{fig:ex3-optimal}
+  \end{figure}
+\end{multicols}
+
+\vspace{-0.3cm}
+
+As shown in Figure \ref{fig:ex3-colored}, using either heuristic with this ordering, we get \(\chi(H) = 5\).
+This gives us \(\frac{|V|}{2}\) colors.
+This is the worst-case for a crown graph as shown in \cite{johnson}, but this graph also demonstrates the upper bounds for these heuristics.
+
+Let's take a look at how this looks in general.
+Let \(P = (V, E)\) be a simple, complete graph.
+We can see that \(\chi(H) = \Delta(H) + 1\).
+This is very easy to see in a \emph{complete} graph, where all vertices are connected to every other vertex.
+This means all vertices have degree \(|V| - 1\).
+Thus, every time we color a node, a new color is needed.
+And since we have \(\Delta(P) = |V| - 1\) and \(|V|\) vertices, we will need \(\Delta(P) + 1\) colors.
+This is stated in \emph{Brooks' Theorem}.
+
+\begin{theorem}[Brooks' Theorem]
+For any connected undirected graph \(G\) with maximum degree \(\Delta\), the chromatic number of \(G\) is at most \(\Delta\) unless \(G\) is a complete graph or an odd cycle, in which case the chromatic number is \(\Delta + 1\).
+\end{theorem}
+
+The proof of Brook's Theorem can be found in \cite{lovasz}.
+Overall, heuristics can produce undesirable results compared to the optimal solution.
+In graph \(G_3\), in the worst case, some heuristics produce \(\chi(H) = 5\) when \(\chi^{*}(H) = 2\) as it is a bipartite graph.
+This is shown in Figure \ref{fig:ex3-optimal}.
+
+\vspace{-0.1cm}
+
+\section{Results}
+Our main goal was to compare these heuristics to find out which heuristic was most effective in efficiently approximating the vertex coloring problem.
+we wanted to compare both quality (number of colors used) and efficiency (running time).
+For this study, we decided to omit the maximal independent set heuristic (heuristic \ref{heuristic-mis}) as it was so different from the other 3 as well as including another NP-hard problem.
+We implemented the three heuristics in Java and ran simulations according to the following:
+\begin{itemize}
+  \setlength\itemsep{0.4em}
+  \item Graphs were generated based on the \emph{Erdős–Rényi} model \cite{erdds1959random}. This model generates a random graph with approximately $p$ percent of edges.
+  \item We analyzed graphs with $10\%, 20\%, \ldots, 90\%$ edges. We called this the type of the graph. This gave us both sparse and dense graphs.
+  \item Simulations were run on graphs with 5 different amount of vertices: 50, 100, 250, 500, 1000.
+\end{itemize}
+
+We used statistical analysis to compare the number of colors used and the running time of each heuristic based on edge percentage (type) and then also looked at how this changed as the number of vertices changed.
+
+\begin{figure}[H]
+  \includegraphics[width=\linewidth]{../poster/figures/plot-colors.pdf}
+  \caption{Comparing number of colors used for the three heuristics}\label{fig:compare-colors}
+\end{figure}
+
+Based on computational data of the amount of colors used, as shown in Figure \ref{fig:compare-colors}, we found:
+\begin{itemize}
+  \item Heuristic \ref{heuristic-greedy} (Greedy) uses on average 2.52 colors more than Heuristic \ref{heuristic-welsh} (Welsh-Powell).
+  \item Heuristic \ref{heuristic-welsh} (Welsh-Powell) uses on average 2.79 colors more than Heuristic \ref{heuristic-dsatur} (DSATUR).
+\end{itemize}
+This led us to the following partial order based on colors used, where $X \preceq Y$ means heuristic $X$ uses significantly more colors than heuristic $Y$.
+\begin{center}
+  \textbf{Greedy $\preceq$ Welsh-Powell $\preceq$ DSATUR}
+\end{center}
+As vertices increased, the partial order defined above stays consistent.
+Heuristic \ref{heuristic-dsatur} (DSATUR) always used the least colors while Heuristic \ref{heuristic-greedy} (Greedy) always used the most.
+
+Based on computational data of the running time, as shown in Figure \ref{fig:compare-time}, we found:
+\begin{itemize}
+  \item Welsh-Powell is significantly slower than Greedy.
+  \item DSATUR is significantly slower than Welsh-Powell.
+\end{itemize}
+This led us to the following partial order based on running time, where $X \preceq Y$ means heuristic $X$ uses significantly less time than heuristic $Y$.
+\begin{center}
+  \textbf{Greedy $\preceq$ Welsh-Powell $\preceq$ DSATUR}
+\end{center}
+
+\newpage
+
+As the edge percentage increases:
+\begin{itemize}
+  \item Heuristic \ref{heuristic-dsatur} (DSATUR) running time increases.
+  \item Heuristic \ref{heuristic-greedy} (Greedy) and Heuristic \ref{heuristic-welsh} (Welsh-Powell) running time increases until 50\% edges then decreases until 90\% edges.
+  \item The difference in running time between DSATUR and Greedy as well as DSATUR and Welsh-Powell grows larger and larger.
+\end{itemize}
+
+\begin{figure}[H]
+  \includegraphics[width=\linewidth]{../poster/figures/plot-time.pdf}
+  \caption{Comparing running time of each heuristic for each edge percentage}\label{fig:compare-time}
+\end{figure}
+
+\section{Conclusion}
+Although we were not able to compare the coloring of the heuristics to the optimal coloring, we were able to determine which heuristics gave the closest to the optimal solution.
+We found that Heuristic \ref{heuristic-dsatur} (DSATUR) produced the lowest number of colors.
+However, DSATUR also utilized the most time especially as the edge percentage of the graph grew.
+This is undesirable, because in large graphs, this defeats our purpose of using heuristics to efficiently approximate the VCP.
+
+Based on our overall results, we would suggest using Heuristic \ref{heuristic-welsh} (Welsh-Powell) for most graphs.
+This is because the extra time needed from Heuristic \ref{heuristic-greedy} (Greedy) to Heuristic \ref{heuristic-welsh} (Welsh-Powell) to achieve better accuracy is not as significant compared to the extra time needed from Heuristic \ref{heuristic-welsh} (Welsh-Powell) to  Heuristic \ref{heuristic-dsatur} (DSATUR).
+
 \newpage

### 2017-05-11 work/paper.tex
Making almost final changes to paper

@@
 \usepackage{multicol}
+\usepackage[numbers]{natbib}
+\bibliographystyle{abbrvnat}
+
 \setlist[itemize]{topsep=0pt}
@@
 
-\title{Vertex Coloring and Applications}
+\title{\vspace{-0.8cm}Vertex Coloring and Applications}
 \date{May 5, 2017}
-\author{Shawn Seymour\\ University of Minnesota Morris}
+\author{\Large{Shawn Seymour\vspace{0.2cm}}\\Directed Study (MATH 3993)\\University of Minnesota Morris}
 
@@
 This problem, along with many others, can be solved using vertex coloring.
+We can model the map of the 48 contiguous states as a graph.
+We can represent every state with a vertex and draw an edge between two states that share a border.
 
@@
 \includegraphics[width=0.7\linewidth]{figures/map-colors.pdf}
-\caption{A proper coloring of the USA}\label{fig:map}
+\caption{A proper vertex coloring of the 48 contiguous states of the USA}\label{fig:map}
 \end{figure}
@@
 The \emph{chromatic number}, denoted $\chi(G)$, is the minimum number of colors needed to have a proper vertex coloring of a graph $G$.
-The \emph{vertex coloring problem} (VCP), when given a graph $G$, is to find $\chi(G)$.
+The \emph{vertex coloring problem} (VCP) is to find $\chi(G)$ of any given graph $G$.
 By applying the vertex coloring problem to our map example, we can determine how many colors one would need to color it with regards to our constraint.
@@
 
+In this section, we will first discuss how to classify hard mathematical problems. Then, we will classify the VCP as a mathematically hard problem and show how this classification can be done.
+
 \subsection{Classifying the Vertex Coloring Problem}
-We call any problem that can be answered with ``yes'' or ``no'' a \emph{decision problem}. To classify the vertex coloring problem, we must first introduce some classes, defined in \cite{moret}:
+We call any problem that can be answered with ``yes'' or ``no'' a \emph{decision problem}. To classify the vertex coloring problem, we must first introduce some classes, defined by \citet{moret}:
 
@@
 Although it may seem like all problems that are NP are also NP-complete, this is not the case.
-There are a select few problems in NP that are not NP-complete or in P \cite{ladner1975structure}.
+There are a select few problems in NP that are not NP-complete or in P.
+See \citet{ladner1975structure}.
 
@@
 This is typically done by a reduction (a transformation) to or from 3-SAT, which we'll examine later.
-This is proven in \cite{garey}.
+This was proven by \citet{garey}.
 
@@
 To prove this, we need to reduce a known NP-complete problem to the vertex coloring problem in polynomial time.
-By reducing a known NP-complete problem to k-Colorability, we can conclude that k-Colorability is NP-complete as all NP-complete problems are reducible to all other NP-complete problems \cite{gareynp}.
+By reducing a known NP-complete problem to k-Colorability, we can conclude that k-Colorability is NP-complete as all NP-complete problems are reducible to all other NP-complete problems.
+See \citet{gareynp}.
 We will reduce the well-known 3-SAT problem first to the 3-Colorability problem.
@@
 Before we prove Proposition \ref{prop:k-color-np-complete}, we must first introduce 3-SAT.
-The 3-SAT problem is an NP-complete problem that is a variation of the original NP-complete problem, boolean satisfiability (SAT).
-The 3-SAT problem is defined and proven to be NP-complete in \cite{gareynp}.
-3-SAT consists of a conjunction of a number of clauses.
-A clause is a disjunction of a given number of propositions or their negations.
-We will assume all of our satisfiability problems are in conjunctive normal form (CNF).
-3-SAT asks whether the variables can be assigned \emph{true} or \emph{false} in such a way that the whole expression evaluates to \emph{true}.
-We define SAT and 3-SAT in Definition \ref{def:sat}.
-
-\begin{definition}\label{def:sat}
-SAT is a conjunction of $m$ clauses, i.e.
+The 3-SAT problem is a decision problem that is a variation of the original NP-complete problem, Boolean satisfiability (SAT).
+The 3-SAT problem was proven to be NP-complete.
+See \citet{gareynp}.
+SAT consists of a conjunction of $m$ clauses, i.e.
 $C_1 \wedge C_2 \wedge \dots \wedge C_m$.
@@
 Each literal has a value of $true$ or $false$.
-3-SAT is where each clause has exactly 3 literals \((n = 3)\).
+3-SAT has exactly 3 literals in each clause $(n = 3)$.
+We will assume all of our satisfiability problems are in conjunctive normal form (CNF).
+3-SAT asks whether the variables can be assigned \emph{true} or \emph{false} in such a way that the whole expression evaluates to \emph{true}.
+We define SAT and 3-SAT in Definition \ref{def:sat}.
+
+\begin{definition}\label{def:sat}
+SAT is a decision problem that decides the truth value of a Boolean function based on the evaluation of the clause.
+It determines whether there are truth values to each clause in a set of clauses that will return a truth value to the given Boolean function.
+3-SAT is a variation of SAT where each clause contains exactly 3 literals.
 \end{definition}
@@
 \begin{enumerate}
-\item Given an instance of 3-SAT, construct an instance of 3-Colorability
+\item Construct an instance of 3-Colorability, given an instance of 3-SAT
 \item Show that if the resulting graph is 3-colorable, then the given 3-SAT problem is satisfied
@@
 To show this reduction, we'll create an example 3-SAT problem and reduce it to 3-Colorability.
-This reduction will create a graph \(G = (V, E)\) having \((2n + 3m + 1)\) vertices and \((3n+6m)\) edges \cite{moret}.
-The process for creating the graph, described below, was taken from the proof given in \cite{sharma}.
+This reduction will create a graph \(G = (V, E)\) having \((2n + 3m + 1)\) vertices and \((3n+6m)\) edges.
+See \citet{moret}.
+The process for creating the graph, described below, was taken from the proof given by \citet{sharma}.
 
@@
 From here, we can reduce 3-Colorability to 4-Colorability, and then generalize that reduction and reduce 4-Colorability to k-Colorability.
-The proof and illustration of this can be found in \cite{sharma}.
+The proof and illustration of this can be found in \citet{sharma}.
 
 \section{Heuristics}
-As we have seen in section \ref{sec:complexity}, solving the vertex coloring problem requires heavy computation.
+As we have seen in section \ref{sec:complexity}, the vertex coloring problem is NP-hard and thus solving it to optimality for any given instance of a graph is computationally ``impossible''.
 This gives us incentive to use approximation algorithms, called heuristics, to find good, but not necessarily optimal, vertex colorings.
-We decided to analyze and compare a few of the popular heuristics for solving the VCP.
+We will analyze and compare a few of the popular heuristics for solving the VCP.
+However, in this section, we will describe the four heuristics to solve the VCP.
 The input for all heuristics we define is a simple graph, $G = (V, E)$.
+
+\newpage
 
@@
 Before applying the greedy algorithm, it orders the vertices according to the decreasing value of their degree.
-This is also known as the Welsh-Powell algorithm, which is given in \cite{welsh}.
+This is also known as the Welsh-Powell algorithm, which was defined by \citet{welsh}.
 It is defined in Algorithm \ref{alg:welsh}.
@@
 
-Heuristic \ref{heuristic-dsatur} is very similar to heuristic \ref{heuristic-welsh}, but colors a graoh based on the \emph{saturation} degree of vertices, which is defined in Definition \ref{def:sat} \cite{spinrad}.
-The original DSATUR algorithm was proposed by \cite{brelaz}.
+Heuristic \ref{heuristic-dsatur} is very similar to heuristic \ref{heuristic-welsh}, but colors a graoh based on the \emph{saturation} degree of vertices, which is defined in Definition \ref{def:saturation} \cite{spinrad}.
+The original DSATUR algorithm was proposed by \citet{brelaz}.
 It is defined in Algorithm \ref{alg:dsatur}.
 
-\begin{definition}[Saturation degree]\label{def:sat}
+\begin{definition}[Saturation degree]\label{def:saturation}
 The saturation degree of a vertex is the number of different colors used for vertices adjacency to it.
@@
 
-Finding a maximum independent set is NP-hard \cite{karp1972reducibility}, however, meaning we're including a different NP-hard in an already NP-hard problem.
+Finding a maximum independent set is NP-hard as detailed by \citet{karp1972reducibility}.
+This means we're including a different NP-hard in an already NP-hard problem.
 We chose to forgo this in favor of just finding a maximal independent set.
@@
 
-\section{Initial Findings}
+\section{Analysis of Heuristics A - D}
 Before comparing the heuristics, we first wanted to analyze the heuristics individually and look at the solutions they produce.
@@
 \begin{prop}
-Heuristic \ref{heuristic-greedy} and Heuristic \ref{heuristic-welsh} do not always produce an optimal solution, i.e. they do not always produce a minimum coloring of a graph.
+Heuristic \ref{heuristic-greedy} and Heuristic \ref{heuristic-welsh} do not produce optimal solutions, i.e. they do not always produce a minimum coloring for all graphs.
 \end{prop}
@@
 
-We can see that $G_1$ is a bipartite graph, defined above by \cite{bondymurty}.
-Thus, \(G_1\) is 2-colorable as all bipartite graphs are 2-colorable \cite{asratian1998bipartite}.
+We can see that $G_1$ is a bipartite graph.
+Thus, \(G_1\) is 2-colorable as all bipartite graphs are 2-colorable.
+See \citet{asratian1998bipartite}.
 This means \(\chi^{*}(G_1) = 2\).
@@
 Heuristic \ref{heuristic-dsatur}, however, does not produce optimal colorings for all graphs.
-Both of these results are shown and proven in \cite{brelaz}.
-We chose not to create examples of these.
+Both of these results are shown and proven by \citet{brelaz}.
 
 \begin{prop}
-Heuristic \ref{heuristic-mis} also does not always produce an optimal solution, i.e. it does not always produce a minimum coloring of a graph.
+Heuristic \ref{heuristic-dsatur} does not produce an optimal solution, i.e. it does not always produce a minimum coloring for all graphs.
+\end{prop}
+
+To show this, we've created a graph $G_2 = (V, E)$ for this example. This is shown in.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{images/dsatur-1.png}
+\caption{Uncolored original graph \(G_2\)}\label{fig:ex2-uncolored}
+\end{figure}
+
+By applying heuristic \ref{heuristic-dsatur}, we can see that $G_2$ is \emph{4-colorable}. This is shown in Figure \ref{fig:ex2-colored}.
+
+\begin{multicols}{2}
+  \begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.5]{images/dsatur-2.png}
+  \caption{Coloring of \(G_2\) after applying heuristic \ref{heuristic-dsatur}}\label{fig:ex2-colored}
+  \end{figure}
+
+  \begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.5]{images/dsatur-3.png}
+  \caption{Optimal coloring of \(G_2\)}\label{fig:ex2-optimal}
+  \end{figure}
+\end{multicols}
+
+This is a non-optimal solution as this graph is \emph{3-colorable}.
+This means \(\chi^{*}(G_2) = 3\).
+Thus, we can see that heuristic \ref{heuristic-dsatur} does not always give us an optimal solution.
+The optimal coloring of $G_2$ is shown in Figure \ref{fig:ex2-optimal}.
+
+\begin{prop}
+Heuristic \ref{heuristic-mis} does not produce an optimal solution, i.e. it does not always produce a minimum coloring for all graphs.
 \end{prop}
@@
 To show this, we will create another example that satisfies conditions \((1)\) and \((2)\) from earlier.
-We've constructed a graph \(G_2 = (V,E)\) for this example.
-This is shown Figure \ref{fig:ex2-uncolored}.
+We've constructed a graph \(G_3 = (V,E)\) for this example.
+This is shown Figure \ref{fig:ex3-uncolored}.
 
@@
 \includegraphics[scale=0.5]{images/mis-1.png}
-\caption{Uncolored original graph \(G_2\)}\label{fig:ex2-uncolored}
+\caption{Uncolored original graph \(G_3\)}\label{fig:ex3-uncolored}
 \end{figure}
 
-By applying heuristic \ref{heuristic-mis} and finding maximal independent sets, we see that \(G_2\) is \emph{5-colorable}. This is shown Figure \ref{fig:ex2-colored}.
+By applying heuristic \ref{heuristic-mis} and finding maximal independent sets, we see that \(G_3\) is \emph{5-colorable}. This is shown Figure \ref{fig:ex3-colored}.
+
+\newpage
 
@@
   \includegraphics[scale=0.5]{images/mis-2.png}
-  \caption{Coloring of \(G_2\) after applying heuristic \ref{heuristic-mis}}\label{fig:ex2-colored}
+  \caption{Coloring of \(G_3\) after applying heuristic \ref{heuristic-mis}}\label{fig:ex3-colored}
   \end{figure}
@@
   \includegraphics[scale=0.5]{images/mis-3.png}
-  \caption{Optimal coloring of \(K\)}\label{fig:ex2-optimal}
+  \caption{Optimal coloring of \(G_3\)}\label{fig:ex3-optimal}
   \end{figure}
@@
 We can see that the subgraph created by vertices \(\{5, 6, 7, 8\}\) is complete and thus we need at least 4 colors.
-This graph is \emph{4-colorable} however, meaning \(\chi^{*}(K) = 4\).
+This graph is \emph{4-colorable} however, meaning \(\chi^{*}(G_3) = 4\).
 Thus, we can see that heuristic \ref{heuristic-mis} does not always give us an optimal solution.
-The optimal coloring of $G_2$ is shown in Figure \ref{fig:ex2-optimal}.
+The optimal coloring of $G_3$ is shown in Figure \ref{fig:ex3-optimal}.
 
 \begin{prop}
-Heuristic \ref{heuristic-greedy} and heuristic \ref{heuristic-welsh} have an upper bound of \(\Delta + 1\).
+Heuristic \ref{heuristic-greedy} and heuristic \ref{heuristic-welsh} yield an upper bound of \(\Delta + 1\).
 \end{prop}
@@
 To show this, we will create another example that satisfies conditions \((1)\) and \((2)\) from above.
-After doing some research into bipartite graphs, we learned that \emph{crown graphs} are excellent at showing how bad greedy heuristics can be, as shown and defined in \cite{kordecki}.
+After doing some research into bipartite graphs, we learned that \emph{crown graphs} are excellent at showing how bad greedy heuristics can be, as shown and defined by \citet{kordecki}.
 
@@
 \includegraphics[scale=0.38]{images/graph-4.png}
-\caption{Uncolored original graph \(G_3\)}
+\caption{Uncolored original graph \(G_4\)}
 \end{figure}
 
-We can see that \(\Delta(G_3) = 4\).
-We can also see \(d_{G_3}(v) = 4\) for all \(v \in V\).
+We can see that \(\Delta(G_4) = 4\).
+We can also see \(d_{G_4}(v) = 4\) for all \(v \in V\).
 Thus, in both heuristics \ref{heuristic-greedy} and \ref{heuristic-welsh}, the algorithm would pick an order arbitrarily.
@@
   \includegraphics[scale=0.38]{images/graph-5.png}
-  \caption{Worst-case coloring of \(G_3\)}\label{fig:ex3-colored}
+  \caption{Worst-case coloring of \(G_4\)}\label{fig:ex4-colored}
   \end{figure}
@@
   \includegraphics[scale=0.38]{images/graph-6.png}
-  \caption{Optimal coloring of \(H\)}\label{fig:ex3-optimal}
+  \caption{Optimal coloring of \(G_4\)}\label{fig:ex4-optimal}
   \end{figure}
@@
 
-As shown in Figure \ref{fig:ex3-colored}, using either heuristic with this ordering, we get \(\chi(H) = 5\).
+As shown in Figure \ref{fig:ex4-colored}, using either heuristic with this ordering, we get \(\chi(G_4) = 5\).
 This gives us \(\frac{|V|}{2}\) colors.
-This is the worst-case for a crown graph as shown in \cite{johnson}, but this graph also demonstrates the upper bounds for these heuristics.
+This is the worst-case for a crown graph.
+See \citet{johnson}.
+This graph also demonstrates the upper bounds for these heuristics.
 
@@
 
-The proof of Brook's Theorem can be found in \cite{lovasz}.
+The proof of Brook's Theorem is shown by \citet{lovasz}.
 Overall, heuristics can produce undesirable results compared to the optimal solution.
-In graph \(G_3\), in the worst case, some heuristics produce \(\chi(H) = 5\) when \(\chi^{*}(H) = 2\) as it is a bipartite graph.
-This is shown in Figure \ref{fig:ex3-optimal}.
+In graph \(G_4\), in the worst case, some heuristics produce \(\chi(G_4) = 5\) when \(\chi^{*}(G_4) = 2\) as it is a bipartite graph.
+This is shown in Figure \ref{fig:ex4-optimal}.
 
@@
   \setlength\itemsep{0.4em}
-  \item Graphs were generated based on the \emph{Erdős–Rényi} model \cite{erdds1959random}. This model generates a random graph with approximately $p$ percent of edges.
+  \item Graphs were randomly generated based on the \emph{Erdős–Rényi} model, given in \citet{erdds1959random}. This model generates a random graph with approximately $p$ percent of edges.
   \item We analyzed graphs with $10\%, 20\%, \ldots, 90\%$ edges. We called this the type of the graph. This gave us both sparse and dense graphs.
@@
 
-\bibliographystyle{abbrv}
 \bibliography{references}

### 2017-05-11 work/paper.tex
Adding labels to graphs

@@
   \includegraphics[width=\linewidth]{../poster/figures/plot-colors.pdf}
-  \caption{Comparing number of colors used for the three heuristics}\label{fig:compare-colors}
+  \caption{Number of colors used for the three heuristics of graphs with 50, 100, 250, 500, \& 1000 vertices}\label{fig:compare-colors}
 \end{figure}
@@
   \includegraphics[width=\linewidth]{../poster/figures/plot-time.pdf}
-  \caption{Comparing running time of each heuristic for each edge percentage}\label{fig:compare-time}
+  \caption{Running time for each edge percentage of graphs with 50, 100, 250, 500, \& 1000 vertices}\label{fig:compare-time}
 \end{figure}
